### 2023-01-13 paper.Rmd
Initial commit

---
title: "Frozen Conflict Onset"
author: J.L.A. Krusell
date: "`r Sys.Date()`"
fontsize: 11pt
linestretch: 1.25
link-citations: true
mainfont: TeX Gyre Pagella
mathfont: TeX Gyre Pagella Math
header-includes:
    - \usepackage[font={small}]{caption}
    - \usepackage{lualatex-math}
output:
    pdf_document:
        dev: cairo_pdf
        keep_tex: true
        latex_engine: lualatex
---

```{r setup, include = F}
library(caTools)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(kableExtra)
library(knitr)
library(stargazer)

options(knitr.kable.NA = ".", knitr.kable.format = "latex",
        knitr.kable.linesep = "")
opts_chunk$set(echo = F, message = F, warning = F)

theme_set(theme_tufte(base_family = "TeX Gyre Pagella", base_size = 10))
```

## Data

To examine the potential correlates of frozen conflict onset, a
dataset at the dyadic - conflict episode level was constructed from
three different sources: the UCDP Conflict Termination Dataset, the
UCDP External Support Dataset (ESD) and the Frozen Conflict Dataset
(FCD) coded by Klosek et al. The first dataset, the UCDP Conflict
Termination dataset, provided conflict-year attributes --- including
duration, type of conflict (interstate vs intrastate), and intensity
level --- for each dyadic relation coded by UCDP from
1949 - 2020. Furthermore, it also codes the episode termination
outcomes in terms of peace agreement, ceasefire agreement, outright
victory, or low activity --- the last of which denotes when a conflict
episode drops below UCDP's threshold of 25 battle related deaths
within a single calendar year.

Data on external support was drawn from UCDP's External Support
Dataset coded from 1975 - 2017 for all intra/interstate
conflicts. This includes an undirected binary measure of whether at
least one actor within a UCDP coded dyad received external support in
a given calendar year, as well as disaggregated dichotomous indicators
for the type of support such as, for example, military support or
direct funding.

Finally, the main dependent variable of interest, the onset of a
frozen conflict, was taken from the Frozen Conflict Dataset which
covers the period from 1946 to 2011. Frozen conflict onset was coded
as the calendar year *after* which initial fighting between two actors
had ceased, which would then be followed by the state of unstable
peace which characterises frozen conflicts.

### Descriptive Statistics

```{R}
df <- readRDS("./data/merged_data.rds")
pos.df <- filter(df, lagged_fc_onset == 1)
```

In total, the final dataset comprises `r nrow(df)` episodes from
`r select(df, side_a, side_b) |> distinct() |> nrow()` unique dyadic
relations with `r nrow(pos.df)` frozen conflict onsets.[^merge] Of
those onsets, `r filter(pos.df, type_of_conflict %in% 3:4) |> nrow()` are
classified as intrastate according to UCDP with
`r filter(pos.df, end_year > 1991) |> nrow()` occurring after the
dissolution of the Soviet Union in 1991.

[^merge]: The FCD dataset lists 42 frozen conflicts in total; however,
    the ESD dataset severely restricts the available time series to
    post-1975. Furthermore, FCD takes a significantly more
    nuanced approach in coding dyadic relations as compared to the
    UCDP. For example, FCD lists separate frozen conflicts between
    Iraq and each coalition partner in the First Gulf War, while UCDP
    only records Iraq - Kuwait as a single conflict dyad.

```{r}
select(pos.df, `Side A` = side_a, `Side B` = side_b, Year = end_year) |>
    arrange(desc(Year)) |>
    kable("latex", booktabs = T, linesep = "",
          caption = "List of Frozen Conflict Onsets") |>
    kable_styling(latex_options = "striped")
```

Compared to the UCDP termination outcomes,
`r filter(pos.df, outcome %in% 1:2) |> nrow()` of the frozen conflict
episodes are coded as having ended with either a ceasefire or peace
agreement and `r filter(pos.df, outcome == 5) |> nrow()` terminated
due to low activity. This leaves `r filter(pos.df, outcome %in% 3:4) |> nrow()`
which concluded with an outright military victory for one
side. Specifically, the Ethiopia - Eritrea (EPLF), Ethiopia - Somalia,
China - N. Vietnam, and Iraq - Kuwait conflicts are all listed as
military victory in the UCDP Termination dataset.

## Analysis

The empirical analysis consists of standard logistic regressions
predicting the onset of a frozen conflict at the dyad-episode
level. More formally, let dependent variable, $y_i \in \{0, 1\}$, be a
binary indicator denoting whether dyad-episode $i$ results in a frozen
conflict. I model the outcome as a bernoulli distributed random
variable, $y_i \sim \text{Bernoulli}(\pi_i)$, where the log-odds of
onset is a linear combination of a set of covariates, $\text{logit}(\pi_i) =
\alpha + X_i' \beta$.

### Results

Figure 1 plots the regression coefficients from separate logistic
regression models predicting frozen conflict onset for each
disaggregated binary indicator of support while controlling for the
maximum intensity level during a conflict episode as well as conflict
type. The only variable with a statistically significant effect is the
indicator for at least one actor receiving funding from an external
actor during a conflict episode.

```{R fig.caption = "Coefficient estimates from separate logistic regression models for each binary indicator of disaggregated support type"}
vars <- c("ext_sup_bin", "ext_x_bin", "ext_w_bin", "ext_m_bin",
          "ext_t_bin", "ext_f_bin", "ext_l_bin")

fits <- lapply(vars, function(v) {
    fml <- sprintf("lagged_fc_onset ~ %s + factor(type_of_conflict) + max_intensity", v) |>
        formula()

    ml <- glm(fml, data = df, family = "binomial")
    ci <- confint(ml)[2,]

    data.frame(var = v, point = coef(ml)[2], lower = ci[1], upper= ci[2])
}) |> bind_rows()

plot.df <- mutate(fits, var = case_when(var == "ext_sup_bin" ~ "Aggregate Support",
                            var == "ext_x_bin" ~ "Military Support",
                                        var == "ext_w_bin" ~ "Weapons",
                                        var == "ext_m_bin" ~ "Material/Logistics",
                                        var == "ext_t_bin" ~ "Training/Expertise",
                                        var == "ext_f_bin" ~ "Funding",
                                        var == "ext_l_bin" ~ "Access to Territory"))

ggplot(plot.df, aes(point, var)) +
    geom_point() +
    geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0) +
    geom_vline(xintercept = 0, linetype = "dotted") +
    xlab("Estimate + 95% CI") +
    theme(axis.title.y = element_blank())
```

Examining only a binary indicator for external support is a crude
measurement for involvement by external actors, especially in cases
where conflict episodes may have a long duration with significant
variation in the amount of support in any given year. Thus, I also run
models using constructed versions of the external funding variable as
a raw count of how many episode years actors received funding from an
external source (`ext_f_count`), an intensity weighted average of
external funding to account for whether external funding during
critical years with the heaviest fighting may be most important
(`ext_f_wavg`), and finally, a discounted weighted average with
external funding weighted by 5-year bins such that latter years
contribute more heavily to the overall score relative to support in
the beginning of a conflict episode (`ext_f_davg`).

The results are shown in table 2. In all cases the constructed
external funding variables are statistically significant and
positively related to the probability of a frozen conflict onset.

```{R results = "asis"}
ml1 <- glm(lagged_fc_onset ~ ext_f_bin, data = df, family = "binomial")

ml2 <- glm(lagged_fc_onset ~ ext_f_bin + factor(type_of_conflict), data = df, family = "binomial")

ml3 <- glm(lagged_fc_onset ~ ext_f_bin + factor(type_of_conflict) +
               max_intensity, data = df, family = "binomial")

ml4 <- glm(lagged_fc_onset ~ ext_f_count + factor(type_of_conflict) +
               max_intensity, data = df, family = "binomial")

ml5 <- glm(lagged_fc_onset ~ ext_f_wavg + factor(type_of_conflict) +
               max_intensity, data = df, family = "binomial")

ml6 <- glm(lagged_fc_onset ~ ext_f_davg + factor(type_of_conflict) +
               max_intensity, data = df, family = "binomial")

stargazer(ml1, ml2, ml3, ml4, ml5, ml6,
          header = F)
```

To get a better understanding of the effect sizes and how external
funding changes the predicted probabilities of frozen conflict onset,
figure 4 plots the receiver operator curve (ROC) for different
predicted probability cutoffs while table 3 lists the confusion matrix
for predicted vs. actual onsets from model 5. This model was able to
correctly classify 4 out of 15 frozen conflicts with the highest
predicted probability being 0.56. In contrast, a model including only
controls with no external support only yields a maximum predicted
probability of 0.41. While the models in table 2 are quite lackluster
in their performance, there is an increase in predictive power through
the inclusion of external funding as a regression covariate.

```{R}
p <- predict(ml5, type = "response")

m <- table(p > 0.5, df$lagged_fc_onset == 1)
rownames(m) <- c("Predicted: No Onset", "Predicted Onset")
colnames(m) <- c("Actual: No Onset", "Actual: Onset")

kable(m, "latex", caption = "Confusion matrix with a 0.5 predicted probability cutoff for model 5")
```

```{R fig.caption = "ROC curve for model 5"}
invisible(colAUC(p, df$lagged_fc_onset, plotROC = T))
```

Finally, it is worth noting that the only frozen conflicts which were
correctly classified were interstate conflicts. The predicted
probabilities for intrastate frozen conflicts were extremely low,
meaning that external funding is potentially a fairly poor predictor
of whether intrastate conflicts become frozen.

```{R}
df$phat <- signif(p, 2)
filter(df, lagged_fc_onset == 1) |>
    arrange(desc(phat)) |>
    select(`Side A` = side_a, `Side B` = side_b, Year = end_year,
           `Prob. of Onset` = phat) |>
    kable("latex", booktabs = T, linesep = "",
          caption = "Predicted Probabilities from Model 5 for each frozen conflict") |>
    kable_styling(latex_options = "striped")
```

## Conclusion

Extremely tentative results point to external funding as a potential
predictor of frozen conflict onset, albeit only in the case of
interstate conflicts. These results are undermined most notably by the
continued lack of additional control variables, especially actor
specific covariates. Ideally, future models would include at a minimum
GDP per capita and level of democracy; however, this would necessitate
splitting interstate and intrastate conflicts into separate
models. Furthermore, there may also be pernicious issues lurking in
regards to time series censoring since none of the datasets align in
coverage.

Finally, although disaggregated data on the types of external support
provides the opportunity for a more nuanced empirical analysis it also
risks devolving into a p-hacking exercise without strong theoretical
reasoning guiding variable selection.

\newpage

## Appendix

```{R results = "asis"}
ml1 <- glm(lagged_fc_onset ~ ext_f_bin, data = df, family = "binomial")

ml2 <- glm(lagged_fc_onset ~ ext_f_bin + factor(type_of_conflict), data = df, family = "binomial")

ml3 <- glm(lagged_fc_onset ~ ext_f_bin + factor(type_of_conflict) +
               intensity_avg, data = df, family = "binomial")

ml4 <- glm(lagged_fc_onset ~ ext_f_count + factor(type_of_conflict) +
               intensity_avg, data = df, family = "binomial")

ml5 <- glm(lagged_fc_onset ~ ext_f_wavg + factor(type_of_conflict) +
               intensity_avg, data = df, family = "binomial")

ml6 <- glm(lagged_fc_onset ~ ext_f_davg + factor(type_of_conflict) +
               intensity_avg, data = df, family = "binomial")

stargazer(ml1, ml2, ml3, ml4, ml5, ml6,
          header = F)
```

```{R results = "asis"}
ml1 <- glm(lagged_fc_onset ~ ext_f_bin, data = df, family = "binomial")

ml2 <- glm(lagged_fc_onset ~ ext_f_bin + duration, data = df, family = "binomial")

ml3 <- glm(lagged_fc_onset ~ ext_f_bin + duration + factor(type_of_conflict),
           data = df, family = "binomial")

ml4 <- glm(lagged_fc_onset ~ ext_f_bin + duration + factor(type_of_conflict) +
               max_intensity, data = df, family = "binomial")

stargazer(ml1, ml2, ml3, ml4, header = F)
```

### 2023-01-13 R/analysis.R

@@ -1,0 +1,50 @@
+#!/usr/bin/env Rscript
+
+library(caTools)
+library(dplyr)
+library(haven)
+library(readxl)
+library(sjPlot)
+
+df <- readRDS("./data/merged_data.rds")
+
+###
+# Descriptive stats
+table(df$lagged_fc_onset)
+
+# How many observations post-Cold War
+filter(df, end_year >= 1991) |> nrow()
+filter(df, lagged_fc_onset == 1, end_year >= 1991) |> nrow()
+filter(df, lagged_fc_onset == 1, censored) |> nrow()
+
+# How many interstate vs intrastate
+filter(df, !type_of_conflict %in% 3:4) |> nrow()
+
+
+
+
+###
+# Analysis - Basic logistic models
+ml <- glm(lagged_fc_onset ~ ext_f_wavg + factor(type_of_conflict) +
+              max_intensity, data = df, family = binomial)
+summary(ml)
+
+p <- predict(ml, type = "response")
+summary(p[df$lagged_fc_onset == 1])
+table(df$lagged_fc_onset, p >= 0.25)
+
+colAUC(p, df$lagged_fc_onset, plotROC = T)
+
+###
+# Intra-state conflicts
+sub.df <- filter(df, type_of_conflict %in% 3:4)
+
+ml <- glm(lagged_fc_onset ~ ext_x_bin  + max_intensity,
+          data = sub.df, family= binomial)
+summary(ml)
+
+p <- predict(ml, type = "response")
+summary(p)
+table(sub.df$lagged_fc_onset, predict(ml, type = "response") >= 0.25)
+
+colAUC(p, sub.df$lagged_fc_onset, plotROC = T)

### 2023-01-13 R/merge.R

@@ -1,0 +1,102 @@
+#!/usr/bin/env Rscript
+
+library(dplyr)
+library(haven)
+library(readxl)
+
+# Termination dataset (1946 - 2020)
+term <- read_xlsx("./data/ucdp-term-dyad-3-2021.xlsx") |>
+    select(conflict_id, dyadep_id, dyadepisode, dyadcount, dyadterm, dyad_id, year,
+           outcome, side_a, side_b, type_of_conflict, intensity_level) |>
+    filter(year <= 2011)
+
+# External Support Dataset - Dyad-year level (1975 - 2017)
+# - Non-active dyad not included in dyadic - 11387 (Myanmar vs ULA)
+esd <- read_dta("./data/ucdp-esd-dy-181.dta") |>
+    filter(active == 1) |>
+    select(year, active, conflict_id, dyad_id, gwno_a, gwno_b, matches("ext"))
+
+df <- full_join(term, esd, by = c("conflict_id", "dyad_id", "year")) |>
+    filter(year <= 2011, !is.na(dyadep_id))
+
+# Frozen conflict dataset - 27 obs after filling in UCDP dyad_id
+load("./data/FCD/fcd.RData")
+fcd <- mutate(fcd,
+              fc_onset = 1,
+              conflictstart_y = ifelse(conflict_id == 18, 1992, conflictstart_y),
+              dyad_id = case_when(conflict_id == 1 ~ 422,
+                                  conflict_id %in% 2:5 ~ 454,
+                                  conflict_id == 6 ~ 737,
+                                  conflict_id == 7 ~ 560,
+                                  conflict_id == 10 ~ 482,
+                                  conflict_id == 13 ~ 569,
+                                  conflict_id == 14 ~ 799,
+                                  conflict_id == 16 ~ 706,
+                                  conflict_id == 18 ~ 571,
+                                  conflict_id == 22 ~ 462,
+                                  conflict_id == 24 ~ 707,
+                                  conflict_id == 26 ~ 835,
+                                  conflict_id == 28 ~ 841,
+                                  conflict_id == 29 ~ 852,
+                                  conflict_id == 30 ~ 839,
+                                  conflict_id == 31 ~ 840,
+                                  conflict_id == 32 ~ 833,
+                                  conflict_id == 33 ~ 721,
+                                  conflict_id == 34 ~ 580,
[…]

### 2023-10-10 R/conflict_candidates.R
Clean up old files

@@ -1,94 +1,0 @@
-#!/usr/bin/env Rscript
-
-library(dplyr)
-library(haven)
-library(readxl)
-library(tidyr)
-
-printf <- function(...) sprintf(...) |> print()
-
-consecutive <- function(x) {
-    if (anyNA(x))
-        warning("NA's in given vector")
-
-    cumsum(c(T, diff(x) != 1))
-}
-
-###
-# Raw Data
-#
-# Peace Agreements 1975 - 2021
-pce <- read_xlsx("./data/raw/ucdp-peace-agreements-221.xlsx") |>
-    select(conflict_id, paid, year, ended, duration, cease, inclusive, pa_type) |>
-    separate_longer_delim(conflict_id, delim = ",") |>
-    mutate(conflict_id = as.numeric(conflict_id),
-           year = as.numeric(year)) |>
-    group_by(conflict_id, year) |>
-    summarise(paid = paste0(paid, collapse = ","),
-              cease = max(cease),
-              inclusive = min(inclusive),
-              pa_type = min(pa_type))
-
-# Conflict Termination Dataset 1948 - 2019
-term <- read_excel("./data/raw/ucdp-term-acd-3-2021.xlsx") |>
-    select(conflict_id, conflictep_id, year, side_a, side_b,
-           confterm, outcome, intensity_level, incompatibility, type_of_conflict) |>
-    filter(type_of_conflict %in% 3:4)
-
-# External Support 1975 - 2017
-esd <- read_dta("./data/raw/ucdp-esd-dy-181.dta") |>
-    filter(active == 1) |>
-    group_by(conflict_id, year) |>
-    summarise(ext_sup = max(ext_sup),
-              ext_f = max(ext_f),
-              ext_m = max(ext_m))
-
-###
-# Join all datasets and select civil conflicts between 1975 & 2017
-full.df <- filter(pce, conflict_id %in% term$conflict_id) |>
-    full_join(term, by = c("conflict_id", "year")) |>
-    left_join(esd, by = c("conflict_id", "year")) |>
-    filter(year >= 1975, year <= 2019)
-
-# Find comprehensive peace agreements signed by *all* conflict actors
-df <- group_by(full.df, conflict_id) |>
-    arrange(year) |>
-    mutate(idx = consecutive(year)) |>
-    group_by(idx, .add = T) |>
-    mutate(pax = !is.na(paid) & pa_type == 1 & inclusive == 1,
[…]

### 2023-10-11 paper.qmd
Switch to quarto

---
title: When Civil Conflicts Freeze
author: J.L.A. Krusell
date: "`r Sys.Date()`"
echo: false
format:
    pdf:
        keep-tex: true
        mainfont: TeX Gyre Pagella
        mathfont: TeX Gyre Pagella
        pdf-engine: lualatex
---

```{r setup, include = F}
library(dplyr)
library(ggplot2)
library(ggthemes)
library(knitr)

options(knitr.kable.NA = ".", knitr.kable.format = "latex",
        knitr.kable.linesep = "")

theme_set(theme_tufte(base_family = "TeX Gyre Pagella", base_size = 10))
```

### 2023-10-11 lib/fc.utils/R/fc.R
Import utility package
First step to building a monorepo for the entire project

@@ -1,0 +1,8 @@
+#' fc
+#'
+#' Utility functions used both in the construction of the fc-dataset and
+#' analysis of frozen conflict onsets.
+#'
+#' @docType package
+#' @name fc
+NULL

### 2023-10-11 lib/fc.utils/R/logging.R

@@ -1,0 +1,34 @@
+#' Formatted logging
+#'
+#' Output a formatted log message. Ellipsis arguments are first passed
+#' to \code{\link{sprintf}} before being printed to console with
+#' \code{\link{print}}.
+#'
+#' @param ... \code{\link{sprintf}} style format string and arguments
+#' @param type Type of message, either INFO or WARN
+#'
+#' @examples
+#' log(type = "INFO", "Hello %s", "world")
+#'
+#' @export
+log <- function(type = c("INFO", "WARN", "ERROR"), ...) {
+    level <- match.arg(type)
+    prefix <- sprintf("[%s] [%s]", level, Sys.time())
+
+    msg <- paste(prefix, sprintf(...))
+    print(msg)
+}
+
+#' @examples
+#' info("Information message")
+#'
+#' @rdname log
+#' @export
+info <- function(...) log(type = "INFO", ...)
+
+#' @examples
+#' warn("Warning message")
+#'
+#' @rdname log
+#' @export
+warn <- function(...) log(type = "WARN", ...)

### 2023-10-11 lib/fc.utils/R/utils.R

@@ -1,0 +1,20 @@
+#' Find consecutive elements
+#'
+#' Given a NumericVector, return the grouping indices for consecutive
+#' elements.
+#'
+#' @param x NumericVector
+#'
+#' @examples
+#' consecutive(c(1900:1905, 1908:1910))
+#'
+#' @export
+consecutive <- function(x) UseMethod("consecutive", x)
+
+#' @export
+consecutive.numeric <- function(x) {
+    if (anyNA(x))
+        warning("NA's in given vector")
+
+    cumsum(c(T, diff(x) != 1))
+}

### 2023-10-11 lib/fc.utils/tests/testthat.R

@@ -1,0 +1,12 @@
+# This file is part of the standard setup for testthat.
+# It is recommended that you do not modify it.
+#
+# Where should you do additional test configuration?
+# Learn more about the roles of various files in:
+# * https://r-pkgs.org/testing-design.html#sec-tests-files-overview
+# * https://testthat.r-lib.org/articles/special-files.html
+
+library(testthat)
+library(fc)
+
+test_check("fc")

### 2023-10-11 lib/fc.utils/tests/testthat/test_utils.R

@@ -1,0 +1,9 @@
+test_that("consecutive", {
+    years <- c(1900:1903, 1905:1906)
+
+    expect_identical(consecutive(years), c(1L, 1L, 1L, 1L, 2L, 2L))
+    expect_identical(suppressWarnings(consecutive(c(1, NA, 2, 2))),
+                     c(1L, rep(NA_integer_, 3L)))
+    expect_warning(consecutive(c(1, NA, 2, 2)))
+    expect_error(suppressWarnings(consecutive(letters[1:3])))
+})

### 2023-10-11 R/conflict_candidates.R
Add dataset construction scripts

@@ -1,0 +1,94 @@
+#!/usr/bin/env Rscript
+
+library(dplyr)
+library(haven)
+library(readxl)
+library(tidyr)
+
+printf <- function(...) sprintf(...) |> print()
+
+consecutive <- function(x) {
+    if (anyNA(x))
+        warning("NA's in given vector")
+
+    cumsum(c(T, diff(x) != 1))
+}
+
+###
+# Raw Data
+#
+# Peace Agreements 1975 - 2021
+pce <- read_xlsx("./data/raw/ucdp-peace-agreements-221.xlsx") |>
+    select(conflict_id, paid, year, ended, duration, cease, inclusive, pa_type) |>
+    separate_longer_delim(conflict_id, delim = ",") |>
+    mutate(conflict_id = as.numeric(conflict_id),
+           year = as.numeric(year)) |>
+    group_by(conflict_id, year) |>
+    summarise(paid = paste0(paid, collapse = ","),
+              cease = max(cease),
+              inclusive = min(inclusive),
+              pa_type = min(pa_type))
+
+# Conflict Termination Dataset 1948 - 2019
+term <- read_excel("./data/raw/ucdp-term-acd-3-2021.xlsx") |>
+    select(conflict_id, conflictep_id, year, side_a, side_b,
+           confterm, outcome, intensity_level, incompatibility, type_of_conflict) |>
+    filter(type_of_conflict %in% 3:4)
+
+# External Support 1975 - 2017
+esd <- read_dta("./data/raw/ucdp-esd-dy-181.dta") |>
+    filter(active == 1) |>
+    group_by(conflict_id, year) |>
+    summarise(ext_sup = max(ext_sup),
+              ext_f = max(ext_f),
+              ext_m = max(ext_m))
+
+###
+# Join all datasets and select civil conflicts between 1975 & 2017
+full.df <- filter(pce, conflict_id %in% term$conflict_id) |>
+    full_join(term, by = c("conflict_id", "year")) |>
+    left_join(esd, by = c("conflict_id", "year")) |>
+    filter(year >= 1975, year <= 2019)
+
+# Find comprehensive peace agreements signed by *all* conflict actors
+df <- group_by(full.df, conflict_id) |>
+    arrange(year) |>
+    mutate(idx = consecutive(year)) |>
+    group_by(idx, .add = T) |>
+    mutate(pax = !is.na(paid) & pa_type == 1 & inclusive == 1,
[…]

### 2023-10-11 R/dyadic_candidates.R

@@ -1,0 +1,61 @@
+#!/usr/bin/env Rscript
+#
+# R script that identifies dyadic level conflict outcomes, which may
+# be frozen conflicts.
+###
+
+library(dplyr)
+library(fc)
+library(readxl)
+library(tidyr)
+
+###
+# Raw data
+#
+# Peace Agreements at the dyadic level 1975 - 2021
+pce <- read_xlsx("./data/raw/ucdp-peace-agreements-221.xlsx") |>
+    select(paid, year, dyad_id, ended, pa_date, duration, cease, inclusive, pa_type, out_iss) |>
+    separate_longer_delim(dyad_id, delim = ",") |>
+    mutate(dyad_id = as.numeric(dyad_id),
+           year = as.numeric(year),
+           end_date = as.Date(duration),
+           pa_date = as.Date(pa_date),
+           duration = end_date - pa_date) |>
+    filter(pa_type == 1, is.na(duration) | end_date > as.Date(paste0(year + 1, "-12-31")))
+
+pce_collapsed <- group_by(pce, dyad_id, year) |>
+    arrange(pa_date) |>
+    slice(n())
+
+# Conflict Terminations at the dyadic level 1948 - 2019
+term <- read_xlsx("./data/raw/ucdp-term-dyad-3-2021.xlsx") |>
+    filter(type_of_conflict %in% 3:4) |>
+    select(conflict_id, dyad_id, dyadepisode, year, side_a, side_b, outcome, intensity_level)
+
+full.df <- filter(pce_collapsed, dyad_id %in% term$dyad_id) |>
+    full_join(term, by = c("dyad_id", "year")) |>
+    filter(between(year, 1975, 2019))
+
+info("Finished with %d dyadic terminations", sum(!is.na(full.df$outcome)))
+
+write.csv(full.df, "./data/dyadic_episodes.csv", row.names = F)
+
+###
+# Filter out candidates for being frozen conflicts
+outcomes <- group_by(full.df, dyad_id) |>
+    arrange(year) |>
+    mutate(idx = consecutive(year)) |>
+    group_by(idx, .add = T) |>
+    mutate(pax = !is.na(paid),
+           next_pax = lead(pax, default = F)) |>
+    filter(!is.na(outcome) & !outcome %in% c(3, 4, 6), between(year, 1975, 2019), !pax, !next_pax)
+
+final <- group_by(outcomes, dyad_id) |>
+    filter(year == min(year)) |>
[…]

### 2023-10-11 codebook.qmd
Merge remaining files from dataset repo

---
title: Codebook v0.1
author: J.L.A. Krusell
format:
    pdf:
        header-includes: \renewcommand{\arraystretch}{1.7}
---

Introduction, coverage

only includes frozen onsets, not all outcomes

only includes first onset

### Definitions

Define a dyadic conflict episode

### Variables

| Variable      | Type    | Description                                                 |
|:--------------|:--------|:------------------------------------------------------------|
| `conflict_id` | Integer | Unique conflict-level identifier                            |
| `dyad_id`     | Integer | Unique dyadic-level identifier                              |
| `year`        | Integer | Onset year                                                  |
| `duration`    | Integer | Number of years until the conflict becomes active again     |
| `side_a`      | String  | Primary state actor in the dyadic-conflict                  |
| `side_b`      | String  | Primary non-state actor in a conflict dyad                  |
| `frozen`      | Boolean | Binary value indicating an onset of a frozen civil conflict |
| `comment`     | String  | Coding notes                                                |
| `confidence`  | Float   | Self-reported coding confidence falling within $[0.5, 1]$   |

### 2023-10-11 coding-protocol.qmd
Merge remaining files from dataset repo

---
title: Coding Protocol v0.1
author: J.L.A. Krusell
format: pdf
---

The dataset was created using a mix of programmatic and manual coding in a two-stage process. The first stage filtered out dyadic civil conflict episode outcomes according to a preset list of rules listed in detail below. This candidate list was then manually corrected in the second stage to form the final list of frozen conflict onsets.

## Coding Procedure

### Stage 1

Dyadic conflict episode outcomes were drawn from the UCDP Conflict Termination dataset v21.1. This means that the definition of a conflict episode and its corresponding termination followed the coding conventions of the UCDP dataset. Namely, dyad-years are coded as active conflicts when the number of battle related fatalities rises above a threshold of 25. Consecutive conflict years, or episodes, are terminated the first year when this number drops below the threshold.

Episode outcomes were selected as potential onsets of frozen civil conflicts when satisfying the following predicates:

- Intrastate and internationalized intrastate conflicts --- *i.e.*, when the UCDP variable $\text{type\_of\_conflict}_i \in \{ 3, 4 \}$

- Outcomes were not coded as either a one-sided military victory or actor death --- *i.e.*, when the UCDP variable $\text{outcome}_i \notin \{ 3, 4, 6 \}$

- No comprehensive peace agreement lasting at least one calendar year signed either in the termination year or the subsequent year. This corresponds to when the UCDP variable $\text{pa\_type} = 1$ while duration was calculated based on the date of the last signature.

The last predicate relied on data drawn from the UCDP Peace Agreement dataset v22.1. Due to the smaller time series in the Peace Agreement dataset, the final list was restricted to conflict outcomes between 1975 and 2019.

```{r echo = F}
candidates <- read.csv("./data/dyadic_candidates.csv")
episodes <- read.csv("./data/dyadic_episodes.csv")
```

In total, `r nrow(candidates)` candidates were identified, comprising approximately
`r sprintf("%.2f", 100 * nrow(candidates) / sum(!is.na(episodes$outcome)))` percent of all outcomes.

### Stage 2

The resulting list was then manually coded to filter out episode outcomes where the non-state actor ceased to exist as a military actor within five years of termination. This includes situations where, for example, the rebel group unilaterally renounces violence as a political strategy and transitions to becoming a political party. Or, when a rebel group becomes so weakened and/or fragmented that they effectively dissolve.

Evaluations were primarily using a mix of UCDP's own description of conflicts and conflict actors, as well as a number of different academic sources.

Since there is a degree of subjectivity in evaluating edge cases, the dataset also includes a short comment justifying the coding as well as a self-reported confidence score.

### 2023-10-12 R/sim.R
Add probit model

@@ -1,0 +1,61 @@
+#!/usr/bin/env Rscript
+
+library(bayesplot)
+library(cmdstanr)
+library(dplyr)
+library(extraDistr)
+library(MASS, include.only = "mvrnorm")
+
+set.seed(43)
+
+n <- 200
+m <- 2
+k <- 1
+
+alpha <- rnorm(1, 0, 5)
+beta <- rnorm(m, 0, 2.5)
+delta <- rnorm(k, 0, 2.5)
+
+X <- mvrnorm(n, rep(0, m), diag(rep(2, m), nrow = m, ncol = m))
+treatment <- sapply(k, \(i) rbinom(n, 1, runif(1, max = 0.75)))
+
+n_countries <- 4L
+sigma <- rhcauchy(1, 1)
+countries <- rep(rnorm(n_countries, 0, sigma), n / n_countries)
+
+theta <- pnorm(alpha + X %*% beta + treatment %*% delta + countries) |> as.vector()
+summary(theta)
+
+y <- rbinom(n, 1, theta)
+table(y)
+
+data <- list(n = n,
+             m = m,
+             k = k,
+             X = X,
+             T = treatment,
+             n_countries = n_countries,
+             country_id = as.factor(countries) |> as.integer(),
+             sample_prior = 1,
+             y = y)
+str(data)
+
+mod <- cmdstan_model("./stan/probit.stan")
+
+# Sample from prior
+data <- list(n = n, m = m, k = k, X = X, T = treatment, n_countries = n_countries,
+             country_id = rep(1, n), sample_prior = 1, y = rep(1, n))
+
+fit <- mod$sample(data = data, chains = 1)
+
+
+
+fit <- mod$sample(data = data, parallel_chains = 4, max_treedepth = 12, adapt_delta = 0.95)
+
+# Parameter plots
+fit$draws(c("alpha", "beta", "delta", "sigma")) |>
+    mcmc_recover_intervals(c(alpha, beta, delta, sigma))
+
+# Predicted probabilities
+theta_hat <- fit$draws("theta", format = "draws_matrix")
+ppc_dens_overlay(as.vector(theta), theta_hat[1:100, ])

### 2023-10-13 lib/fc.utils/R/logging.R
Rename log function

@@ -8,10 +8,10 @@
 #' @param type Type of message, either INFO or WARN
 #'
 #' @examples
-#' log(type = "INFO", "Hello %s", "world")
+#' log_msg(type = "INFO", "Hello %s", "world")
 #'
 #' @export
-log <- function(type = c("INFO", "WARN", "ERROR"), ...) {
+log_msg <- function(type = c("INFO", "WARN", "ERROR"), ...) {
     level <- match.arg(type)
     prefix <- sprintf("[%s] [%s]", level, Sys.time())
 
@@ -22,13 +22,13 @@ log <- function(type = c("INFO", "WARN", "ERROR"), ...) {
 #' @examples
 #' info("Information message")
 #'
-#' @rdname log
+#' @rdname log_msg
 #' @export
-info <- function(...) log(type = "INFO", ...)
+info <- function(...) log_msg(type = "INFO", ...)
 
 #' @examples
 #' warn("Warning message")
 #'
-#' @rdname log
+#' @rdname log_msg
 #' @export
-warn <- function(...) log(type = "WARN", ...)
+warn <- function(...) log_msg(type = "WARN", ...)

### 2023-10-19 doc/codebook.qmd
Add definitions to codebook

---
title: Codebook
author: J.L.A. Krusell
date: today
linestretch: 1.25
bibliography: ../library.bib
format:
    pdf:
        header-includes: \renewcommand{\arraystretch}{1.7}
---

# Introduction

This dataset codes the onset of frozen civil conflicts from 1975 until 2019. The unit of analysis is dyadic civil conflict episodes as defined by the UCDP/PRIO Armed Conflict Dataset [@gleditsch_armed_2002; @davies_organized_2023]. The onset of a frozen conflict signifies an episode termination where the underlying issue remained unresolved, and both actors continued to exist post-conflict with a significant threat of a renewal of violence.

For a detailed overview of the coding process see the Coding Protocol document.

# Definitions

**State-based armed conflict** is defined according to the UCDP definition as "a contested incompatibility that concerns government and/or territory where the use of armed force between two parties, of which at least one is the government of a state, results in at least 25 battle-related deaths in a calendar year" [@gleditsch_armed_2002]. Civil conflict refers to cases where the opposing side consists of at least one or more non-state actors.

**Conflict episodes** are contiguous years of active conflict as coded by the UCDP/PRIO Armed Conflict Dataset. Termination occurs the first year in which the number of battle-related deaths falls below the UCDP threshold of 25.

**Frozen conflicts** follow the definition offered by @smetana_between_2019 as being "...a protracted, post-war conflict process, characterized by the absence of stable peace between opposing sides." Crucially, this requires that conflict terminations fulfill two criteria in order to be classified as frozen: continued conflict salience and the threat of violent re-escalation. The former in this context means that the core issue underlying the initial conflict remains unresolved. More specifically, the conflict episodes cannot have terminated in a one-sided victory according to the UCDP conflict termination dataset [@kreutz_how_2010] and/or comprehensive peace agreement as recorded by the UCDP peace agreement dataset [@pettersson_organized_2019]. The latter implies that conflict actors need to have survived the initial period active conflict and have not renounced violence as a political strategy.

## Variables

| Variable      | Type    | Description                                                  |
|:--------------|:--------|:-------------------------------------------------------------|
| `conflict_id` | Integer | UCDP unique conflict-level identifier                        |
| `dyad_id`     | Integer | UCDP Unique dyadic-level identifier                          |
| `year`        | Integer | Onset year                                                   |
| `duration`    | Integer | Number of years until the conflict becomes active again      |
| `Intensity`   | Integer | Cumulative intensity?                                        |
| `side_a`      | String  | Primary state actor in the dyadic-conflict                   |
| `side_b`      | String  | Primary non-state actor in a conflict dyad                   |
| `frozen`      | Boolean | Binary value indicating the onset of a frozen civil conflict |
| `comment`     | String  | Coding notes                                                 |
| `confidence`  | Float   | Self-reported coding confidence falling within $[0.5, 1]$    |

# Notes

Only the initial episode terminating in a frozen conflict is coded by this dataset. Subsequent episodes of conflict re-escalation, which may terminate again as frozen conflicts, are not included.

It should be further noted that although this dataset uses the definition of frozen conflicts provided by @smetana_between_2019, it diverges significantly from the operationalization used in the accompanying dataset [@klosek_frozen_2021]. Notably, frozen civil conflicts are *not* restricted to only secessionist conflicts spawning *de facto* independent states. Rather, all civil conflicts as coded by the UCDP are included, inclusive of conflicts over control of government. Furthermore, in following the coding criteria of the UCDP, the threshold for termination is significantly lower. @klosek_frozen_2021 only consider cases reaching a minimum of 500 battle-related fatalities.

For those wishing to apply a more strict coding criteria, the provided cases can be further refined according to the cumulative intensity of the conflict and/or the duration of the post-conflict frozen period.

# References

::: {#refs}
:::

### 2023-10-19 R/conflict_candidates.R
Conform conflict candidates to coding protocol
Script to find conflict-level episodes which do not terminate in
either a peace agreement or one-sided victory. This was originally
written before the coding protocol was formalized. Update it in line
with the dyadic counterpart, in particular the handling of peace
agreements in according with doc/coding-protocol.qmd.

@@ -1,94 +1,65 @@
 #!/usr/bin/env Rscript
 
 library(dplyr)
+library(fc)
 library(haven)
 library(readxl)
 library(tidyr)
-
-printf <- function(...) sprintf(...) |> print()
-
-consecutive <- function(x) {
-    if (anyNA(x))
-        warning("NA's in given vector")
-
-    cumsum(c(T, diff(x) != 1))
-}
 
 ###
 # Raw Data
 #
 # Peace Agreements 1975 - 2021
+#  pa_type == 1 Full peace agreement
+#  inclusive == 1 Comprehensive peace agreement
 pce <- read_xlsx("./data/raw/ucdp-peace-agreements-221.xlsx") |>
-    select(conflict_id, paid, year, ended, duration, cease, inclusive, pa_type) |>
+    select(paid, year, conflict_id, ended, pa_date, duration, cease, inclusive, pa_type, out_iss) |>
     separate_longer_delim(conflict_id, delim = ",") |>
     mutate(conflict_id = as.numeric(conflict_id),
-           year = as.numeric(year)) |>
-    group_by(conflict_id, year) |>
-    summarise(paid = paste0(paid, collapse = ","),
-              cease = max(cease),
-              inclusive = min(inclusive),
-              pa_type = min(pa_type))
+           year = as.numeric(year),
+           end_date = as.Date(duration),
+           pa_date = as.Date(pa_date),
+           duration = end_date - pa_date) |>
+    filter(inclusive == 1, pa_type == 1,
+           is.na(duration) | end_date > as.Date(paste0(year + 1, "-12-31")))
+
+pce_collapsed <- group_by(pce, conflict_id, year) |>
+    arrange(pa_date) |>
+    slice(n())
 
 # Conflict Termination Dataset 1948 - 2019
 term <- read_excel("./data/raw/ucdp-term-acd-3-2021.xlsx") |>
-    select(conflict_id, conflictep_id, year, side_a, side_b,
-           confterm, outcome, intensity_level, incompatibility, type_of_conflict) |>
-    filter(type_of_conflict %in% 3:4)
-
-# External Support 1975 - 2017
-esd <- read_dta("./data/raw/ucdp-esd-dy-181.dta") |>
-    filter(active == 1) |>
-    group_by(conflict_id, year) |>
-    summarise(ext_sup = max(ext_sup),
-              ext_f = max(ext_f),
-              ext_m = max(ext_m))
[…]

### 2023-10-19 R/dyadic_candidates.R
Remove superfluous filter condition

@@ -48,7 +48,7 @@ outcomes <- group_by(full.df, dyad_id) |>
     group_by(idx, .add = T) |>
     mutate(pax = !is.na(paid),
            next_pax = lead(pax, default = F)) |>
-    filter(!is.na(outcome) & !outcome %in% c(3, 4, 6), between(year, 1975, 2019), !pax, !next_pax)
+    filter(!is.na(outcome) & !outcome %in% c(3, 4, 6), !pax, !next_pax)
 
 final <- group_by(outcomes, dyad_id) |>
     filter(year == min(year)) |>

### 2023-10-19 R/dataset.R
Create a conflict level dataset
Move the old merge script using the Klosek et al FCD dataset to fcd_merge.R

@@ -1,0 +1,25 @@
+#!/usr/bin/env Rscript
+#
+# Create final version of the frozen conflict dataset
+###
+
+library(dplyr)
+library(fc)
+
+adj <- read.csv("./data/dataset/adjusted_conflict_candidates.csv") |>
+    filter(Correction == 0) |>
+    select(conflict_id, year, frozen)
+
+episodes <- read.csv("./data/conflict_episodes.csv") |>
+    select(conflict_id, conflictep_id, year, gwno_a, side_a, side_b, intensity_level,
+           incompatibility, recur)
+
+cy <- left_join(episodes, adj, by = c("conflict_id", "year")) |>
+    mutate(frozen = ifelse(!is.na(frozen), 1, 0)) |>
+    group_by(conflict_id) |>
+    filter(year <= if (any(frozen == 1)) min(year[frozen == 1]) else max(year))
+
+info("Final dataset has %d conflict episodes of which %d are frozen",
+     n_distinct(cy$conflictep_id), sum(cy$frozen))
+
+saveRDS(cy, "./data/dataset/frozen_conflicts.rds")

### 2023-10-19 R/fcd_merge.R

@@ -1,0 +1,131 @@
+#!/usr/bin/env Rscript
+
+library(dplyr)
+library(haven)
+library(readxl)
+library(tidyr)
+
+printf <- function(...) sprintf(...) |> print()
+
+# Aggregation formulas for summarise, year -> dyad-level
+weighted_average <- function(x, intensity, side_b) {
+    v <- as.integer(intensity)[!is.na(x)]
+    weighted.mean(x, v / sum(v), na.rm = T)
+}
+
+# Termination dataset (1946 - 2020)
+term <- read_excel("./data/ucdp-term-dyad-3-2021.xlsx") |>
+    select(conflict_id, dyadep_id, dyadepisode, dyadcount, dyadterm, dyad_id, year,
+           outcome, side_a, side_a_id, side_b, side_b_id, type_of_conflict, intensity_level) |>
+    group_by(dyad_id, dyadepisode) |>
+    filter(max(year) >= 1975, min(year) <= 2011)
+
+# External Support Dataset - Dyad-year level (1975 - 2017)
+# - Non-active dyad not included in dyadic - 11387 (Myanmar vs ULA)
+esd <- read_dta("./data/ucdp-esd-dy-181.dta") |>
+    filter(active == 1) |>
+    select(year, active, conflict_id, dyad_id, gwno_a, gwno_b, matches("ext"))
+
+df <- full_join(term, esd, by = c("conflict_id", "dyad_id", "year")) |>
+    filter(year <= 2011, !is.na(dyadep_id)) |>
+    group_by(dyad_id, dyadepisode) |>
+    fill(gwno_a, .direction = "updown") |>
+    ungroup() |>
+    filter(!is.na(gwno_a))
+
+printf("%d dyadic episode years from %d unique dyads", nrow(df), n_distinct(df$dyad_id))
+
+# Frozen conflict dataset - 27 obs after filling in UCDP dyad_id
+load("./data/FCD/fcd.RData", envir = e <- new.env())
+refs <- read.csv("./refs/ucdp_fcd.csv")
+
+fcd <- left_join(e$fcd, refs, by = "conflict_id") |>
+    mutate(fc_onset = 1,
+           conflictstart_y = ifelse(conflict_id == 18, 1992, conflictstart_y)) |>
+    select(dyad_id, year = conflictstart_y, fc_onset) |>
+    filter(!is.na(dyad_id)) |>
+    group_by(dyad_id) |>
+    summarise(year = first(year),
[…]

### 2023-10-19 R/merge.R

@@ -1,69 +1,40 @@
 #!/usr/bin/env Rscript
 
 library(dplyr)
+library(fc)
 library(haven)
-library(readxl)
 library(tidyr)
 
-printf <- function(...) sprintf(...) |> print()
+# Full frozen conflict dataset
+ep <- readRDS("./data/dataset/frozen_conflicts.rds")
 
-# Aggregation formulas for summarise, year -> dyad-level
-weighted_average <- function(x, intensity, side_b) {
-    v <- as.integer(intensity)[!is.na(x)]
-    weighted.mean(x, v / sum(v), na.rm = T)
-}
-
-# Termination dataset (1946 - 2020)
-term <- read_excel("./data/ucdp-term-dyad-3-2021.xlsx") |>
-    select(conflict_id, dyadep_id, dyadepisode, dyadcount, dyadterm, dyad_id, year,
-           outcome, side_a, side_a_id, side_b, side_b_id, type_of_conflict, intensity_level) |>
-    group_by(dyad_id, dyadepisode) |>
-    filter(max(year) >= 1975, min(year) <= 2011)
-
-# External Support Dataset - Dyad-year level (1975 - 2017)
-# - Non-active dyad not included in dyadic - 11387 (Myanmar vs ULA)
-esd <- read_dta("./data/ucdp-esd-dy-181.dta") |>
+###
+# UCDP External support dataset - actor-year level
+esd <- read_dta("./data/raw/ucdp-esd-ay-181.dta") |>
     filter(active == 1) |>
-    select(year, active, conflict_id, dyad_id, gwno_a, gwno_b, matches("ext"))
-
-df <- full_join(term, esd, by = c("conflict_id", "dyad_id", "year")) |>
-    filter(year <= 2011, !is.na(dyadep_id)) |>
-    group_by(dyad_id, dyadepisode) |>
-    fill(gwno_a, .direction = "updown") |>
-    ungroup() |>
-    filter(!is.na(gwno_a))
-
-printf("%d dyadic episode years from %d unique dyads", nrow(df), n_distinct(df$dyad_id))
-
-# Frozen conflict dataset - 27 obs after filling in UCDP dyad_id
-load("./data/FCD/fcd.RData", envir = e <- new.env())
-refs <- read.csv("./refs/ucdp_fcd.csv")
-
-fcd <- left_join(e$fcd, refs, by = "conflict_id") |>
-    mutate(fc_onset = 1,
[…]

### 2023-10-19 R/probit.R
Add initial model script

@@ -1,0 +1,33 @@
+#!/usr/bin/env Rscript
+
+library(bayesplot)
+library(cmdstanr)
+library(dplyr)
+
+options(mc.cores = parallel::detectCores() - 1)
+
+df <- readRDS("./data/model_data.rds") |> ungroup()
+
+X <- select(df, incompatibility, cold_war, duration, max_intensity) |>
+    data.matrix()
+
+cases <- rowSums(is.na(X)) == 0
+
+treatments <- select(df, ext_sup_s_state_max_5y, ext_sup_s_rebel_max_5y) |>
+    mutate(interaction = ext_sup_s_state_max_5y * ext_sup_s_rebel_max_5y) |>
+    data.matrix()
+
+data <- list(n = sum(cases),
+             k = ncol(treatments),
+             T = treatments[cases, ],
+             m = ncol(X),
+             X = X[cases, ],
+             n_countries = n_distinct(df$gwno_a),
+             country_id = as.factor(df$gwno_a[cases]) |> as.integer(),
+             y = df$frozen[cases])
+str(data)
+
+mod <- cmdstan_model("./stan/probit.stan")
+fit <- mod$sample(data = data)
+
+fit$save_object("posteriors/probit.rds")

### 2023-10-19 paper.qmd
Merge remaining files from dataset repo
Initial partial draft of data description in manuscript

@@
 date: "`r Sys.Date()`"
+linestretch: 1.25
+bibliography: library.bib
+format: pdf
+execute:
     echo: false
-format:
-    pdf:
-        keep-tex: true
-        mainfont: TeX Gyre Pagella
-        mathfont: TeX Gyre Pagella
-        pdf-engine: lualatex
+    warning: false
 ---
@@
 ```{r setup, include = F}
+library(bayesplot)
 library(dplyr)
@@
 
-theme_set(theme_tufte(base_family = "TeX Gyre Pagella", base_size = 10))
+theme_set(theme_tufte())
 ```
+
+**Research question: to what extent does external state support to conflict actors affect the probability that a civil conflict becomes frozen?**
+
+# Empirics
+
+To empirically assess whether external state support for conflict
+actors affects the probability of a civil conflict episode terminating
+as a frozen conflict a series of bayesian probit models were run
+using original data on frozen civil conflicts. A mixed identification strategy was employed by both conditioning on relevant covariates and instrumenting external support with UN roll call votes.
+
+In what follows, I will first describe the dependent and independent variables, including operationalizations and data sources.
+I will then present the model specification and results from a conventional probit regression followed by the instrumental variable analysis.
+
+## Data
+
+### Defining Frozen Conflicts
+
+I follow the definition of frozen conflicts offered by @smetana_between_2019 as a "protracted, post-war conflict process, characterized by the absence of stable peace...[where] core issues between the opposing sides remain unresolved, the dispute is in the forefront of mutual relations, and there is a looming threat of the renewal of violence."  Accompanying this definition, @smetana_between_2019 lay forth three criteria determining whether a post-conflict period qualifies as a frozen conflict.
+First, there needs to be an initial period of active conflict followed by a cessation of large-scale violent hostilities.
+Isolated, sporadic incidences of violence may continue to occur; however, conflict actors are either unwilling or unable to launch major military offensives.
+Although not necessary, this may be characterized by the establishment of a general ceasefire and/or presence of external peacekeeping troops.
+Second, the core issue underlying the conflict needs to remain salient post-conflict.
+This entails actors maintaining their claims within a dispute.
+And, finally, there must be an ongoing threat of violent re-escalation.
+In other words, post-conflict relations should be characterized by mutual hostility and an unwillingness to abandon violence as a political strategy.
+
+These criteria distinguish frozen conflicts from related concepts such as protracted stalemates or enduring rivalries.
+In the case of the former, stalemates during a conflict can potentially endure for extended periods of time and offer conditions similar to those of a frozen conflict.
+However, they do not necessitate a cessation of active fighting.
+Stalemates can still occur during high intensity periods of a conflict where relatively few gains are made by either side.
+In the case of the latter, enduring rivalries are characterized in part by repeated onsets of militarized interstate disputes [@Goertz1992; @Goertz1993]; however, this might not entail the actual outbreak of violence.
+More importantly, enduring rivalries were originally conceptualized in terms of inter-state relations, while the definition of frozen conflicts here can be extended to the intra-state case as well.
+
+### Measuring Frozen Conflicts
+
+In order to measure the onset of frozen civil conflicts I followed the previous three criteria and coded an original dataset using dyadic intra-state conflict episodes drawn from the UCDP/PRIO Armed Conflict Dataset [@gleditsch_armed_2002; @davies_organized_2023] defined as contiguous years with at least 25 battle-related deaths in a violent contest over a government or territory between a state and non-state actor.
+To address post-conflict issue saliency, I considered an episode to be unresolved if it terminated without either a comprehensive peace treaty addressing the core underlying issues or a one-sided military victory for either the government or rebel group according to the UCDP Peace Agreement and Conflict Termination Datasets [@pettersson_organized_2019; @kreutz_how_2010].
+I also filtered out episodes where one or more actors ceased to exist within five years of the conflict termination and where the non-state actor renounces violences.
+This was done using a manual coding process with a mix of academic and news media sources for different conflicts in order to capture instances where, for example, rebel groups transition into legitimate political parties.
+
+```{r}
+dataset <- readRDS("./data/model_data.rds")
+n_frozen <- filter(dataset, frozen == 1) |> nrow()
+n_terr_frozen <- filter(dataset, frozen == 1, incompatibility == 1) |> nrow()
+```
+
+The resulting dataset consists of `r n_frozen` frozen conflicts out
+of `r nrow(dataset)` intra-state episodes, comprising approximately
+`r signif(n_frozen / nrow(dataset), 2) * 100` percent of all observations for the period 1975 to 2019.
+Of the frozen conflicts, `r n_terr_frozen` are conflicts over territory.
+(TODO: show world map and time plot)
+
+```{r}
+filter(dataset, frozen == 1) |>
+    ggplot(aes(year)) +
+    geom_bar() +
+    ylab("Frozen Civil Conflict Onsets") +
+    xlab("Year")
+```
+
+Compared to the Frozen Conflict Dataset by @klosek_frozen_2021,
+
+Compared to FCD --- significant more cases
+  - More expansive operationalization
+    - Not only looking at de facto states
+    - Captures frozen moments in civil wars over government, ex Angola
+
+  - Timing of Frozen conflicts
+    - FCD last year in which >= 500 BRD
+    - My datasets follows UCDP last year >25
+      - Captures lower level conflicts
+
+Potential Weaknesses
+  - Low-level long running insurgencies may not be "salient"
+  - Termination year inexact
+  - Potentially capturing shorter stalemates
+
+### Treatments
+
+UCDP external support dataset
+
+### Control Variables
+
+- Conflict Intensity
+- Conflict Duration
+- Incompatibility?
+- e_gdppc
+- e_pop
+- CINC
+
+### Descriptive Statistics
+
+## Observational Models
+
+\begin{align*}
+y_i^\star & = \alpha + X_i' \beta + T_i' \delta + \epsilon_i, \; \epsilon_i \sim \text{N}(0, \sigma^2) \\
+y_i & = \begin{cases}
+1 & \text{if } y_i^\star > 0 \\
+0 & \text{otherwise}
+\end{cases}
+\end{align*}
+
+### Results
+
+```{r}
+fit <- readRDS("./posteriors/probit.rds")
+
+fit$draws("delta") |> mcmc_areas() +
+    scale_y_discrete(labels = c("delta[1]" = "Ext. Sup. to Gov",
+                                "delta[2]" = "Ext. Sup. to Rebel",
+                                "delta[3]" = "Interaction"))
+```
+
+- Parameter Plot
+- Density of predicted probabilities for true success/failure
+- Marginal effects
+- Confusion Matrix + AUR/PRC plots
+
+#### Robustness Checks
+
+- External support only within the last five years of conflict
+- Frozen conflicts with a duration of only 1 year
+- Exclude low intensity conflicts?
+
+Appendix:
+
+- Proportion of episode years with external support
+- Nr of external supporters, maybe weighted by cinc?
+- Territorial conflicts only
+- FCD data
+
+## IV Model
+
+\newpage
+
+# References

### 2023-10-20 R/conflict_candidates.R
Standalone conflict level dataset
The goal is to code frozen conflicts at the dyadic level, but for now
focus on the preliminary collected data at the conflict level.

@@ -41,9 +41,6 @@ full.df <- filter(pce_collapsed, conflict_id %in% term$conflict_id) |>
 
 info("Finished with %d conflict terminations", sum(!is.na(full.df$outcome)))
 
-filter(full.df, !is.na(gwno_a)) |>
-    write.csv("./data/conflict_episodes.csv", row.names = F)
-
 ###
 # Filter out candidates for being frozen conflicts
 outcomes <- group_by(full.df, conflict_id) |>
@@ -52,7 +49,7 @@ outcomes <- group_by(full.df, conflict_id) |>
     group_by(idx, .add = T) |>
     mutate(pax = !is.na(paid),
            next_pax = lead(pax, default = F)) |>
-    filter(!is.na(outcome) & !outcome %in% c(3, 4, 6), between(year, 1975, 2019), !pax, !next_pax)
+    filter(!is.na(outcome) & !outcome %in% c(3, 4, 6), !pax, !next_pax)
 
 final <- group_by(outcomes, conflict_id) |>
     filter(year == min(year)) |>

### 2023-10-20 R/dataset.R

@@ -5,21 +5,34 @@
 
 library(dplyr)
 library(fc)
+library(readxl)
 
 adj <- read.csv("./data/dataset/adjusted_conflict_candidates.csv") |>
     filter(Correction == 0) |>
     select(conflict_id, year, frozen)
 
-episodes <- read.csv("./data/conflict_episodes.csv") |>
-    select(conflict_id, conflictep_id, year, gwno_a, side_a, side_b, intensity_level,
-           incompatibility, recur)
+###
+# Add cumulative intensity - threshold 1000 BRD
+ucdp <- readRDS("./data/raw/UcdpPrioConflict_v23_1.rds") |>
+    select(conflict_id, year, cumulative_intensity)
 
-cy <- left_join(episodes, adj, by = c("conflict_id", "year")) |>
+df <- left_join(adj, ucdp, by = c("conflict_id", "year"))
+
+###
+# Calculate duration until next active conflict episode
+episodes <- read_excel("./data/raw/ucdp-term-acd-3-2021.xlsx") |>
+    filter(type_of_conflict %in% 3:4) |>
+    select(conflict_id, conflictep_id, year, side_a, side_b, recur)
+
+full.df <- left_join(episodes, df, by = c("conflict_id", "year")) |>
     mutate(frozen = ifelse(!is.na(frozen), 1, 0)) |>
     group_by(conflict_id) |>
-    filter(year <= if (any(frozen == 1)) min(year[frozen == 1]) else max(year))
+    arrange(year) |>
+    mutate(duration = ifelse(frozen == 1, lead(year) - year, 0)) |>
+    filter(frozen == 1)
 
-info("Final dataset has %d conflict episodes of which %d are frozen",
-     n_distinct(cy$conflictep_id), sum(cy$frozen))
+info("Final dataset: %d frozen conflicts", nrow(full.df))
 
-saveRDS(cy, "./data/dataset/frozen_conflicts.rds")
+select(full.df, conflict_id, year, duration, cumulative_intensity,
+       side_a, side_b, frozen, recur) |>
+    saveRDS("./data/dataset/frozen_conflicts.rds")

### 2023-10-20 R/merge.R

@@ -3,10 +3,34 @@
 library(dplyr)
 library(fc)
 library(haven)
+library(readxl)
 library(tidyr)
 
-# Full frozen conflict dataset
-ep <- readRDS("./data/dataset/frozen_conflicts.rds")
+###
+# Construct first our full frozen + non-frozen episodes dataset
+frozen <- readRDS("./data/raw/frozen_conflicts.rds") |>
+    select(conflict_id, year, duration, frozen)
+
+term <- read_excel("./data/raw/ucdp-term-acd-3-2021.xlsx") |>
+    filter(type_of_conflict %in% 3:4) |>
+    select(conflict_id, conflictep_id, year, gwno_a = gwno_loc, side_a, side_b,
+           intensity_level, incompatibility, recur) |>
+    group_by(conflictep_id) |>
+    mutate(censored = ifelse(min(year) < 1975, 1, 0),
+           gwno_a = as.numeric(gwno_a)) |>
+    ungroup()
+
+ucdp <- readRDS("./data/raw/UcdpPrioConflict_v23_1.rds") |>
+    select(conflict_id, year, cumulative_intensity)
+
+ep <- filter(term, between(year, 1975, 2019)) |>
+    full_join(frozen, by = c("conflict_id", "year")) |>
+    left_join(ucdp, by = c("conflict_id", "year")) |>
+    mutate(frozen = ifelse(is.na(frozen), 0, 1))
+
+# Exclude subsequent episodes after onset of a frozen conflict
+reduced <- group_by(ep, conflict_id) |>
+    filter(year <= if (any(frozen == 1)) min(year[frozen == 1]) else max(year))
 
 ###
 # UCDP External support dataset - actor-year level
@@ -25,12 +49,11 @@ state_sup <- filter(esd, actor_nonstate == 0) |>
 
 full_ep.df <- full_join(rebel_sup, state_sup, by = c("conflict_id", "year"),
                     suffix = c("_rebel", "_state")) |>
-    right_join(ep, by = c("conflict_id", "year"))
+    right_join(reduced, by = c("conflict_id", "year"))
 
 ###
 # NMC - CINC data
 nmc <- read.csv("./data/raw/NMC-60-abridged.csv")
-
 
 ###
 # V-Dem - dem., gdppc, population
@@ -68,13 +91,17 @@ info("Missing CINC observation(s): %d", is.na(full.df$cinc) |> sum())
[…]

### 2023-10-20 R/merge.R
Parameterize model runs for probit regressions

@@ -96,6 +96,7 @@ final.df <- filter(full.df, !is.na(ext_sup_s_state)) |>
               year = last(year),
               side_a = first(side_a),
               side_b = first(side_b),
+              censored = max(censored),
               episode_duration = n(),
               frozen_duration = last(duration),
               recur = max(recur),
@@ -109,13 +110,13 @@ final.df <- filter(full.df, !is.na(ext_sup_s_state)) |>
               across(c(v2x_polyarchy, e_gdppc, e_pop, cinc), ~mean(.x, na.rm = T), .names = "{.col}_avg"),
 
               # At anytime during the episode was external support given?
-              across(starts_with("ext_"), ~max(.x, na.rm = T), .names = "{.col}_max"),
+              across(starts_with("ext_"), ~max(.x, na.rm = T), .names = "{.col}_bin"),
 
               # Proportion of conflict years within an episode with external support
               across(starts_with("ext_"), ~mean(.x, na.rm = T), .names = "{.col}_prop"),
 
               # Whether external support was given in the last five years before termination
-              across(starts_with("ext_"), ~max(tail(.x, 5), na.rm = T), .names = "{.col}_max_5y")) |>
+              across(starts_with("ext_"), ~max(tail(.x, 5), na.rm = T), .names = "{.col}_bin_5y")) |>
     mutate(strict_frozen = ifelse(is.na(frozen_duration) | frozen_duration > 2, frozen, 0))
 
 info("Finished with %d conflict episodes and %d frozen conflicts",

### 2023-10-20 R/models.R

@@ -1,0 +1,19 @@
+#!/usr/bin/env Rscript
+
+library(dplyr)
+library(fc)
+library(yaml)
+
+dir.create("models", showWarnings = F)
+
+treatments <- c("bin", "bin_5y", "prop")
+outcomes <- c("frozen", "strict_frozen")
+units <- c("all", "high_intensity")
+
+df <- expand.grid("outcome" = outcomes, "treatment" = treatments, "episodes" = units) |>
+    mutate(name = paste0(outcome, "_", treatment, "_", episodes))
+
+for (row in 1:nrow(df)) {
+    info("Writing model profile %s", df[row, "name"])
+    write_yaml(df[row, ], sprintf("./models/%s.yml", df[row, "name"]))
+}

### 2023-10-20 R/probit.R

@@ -1,22 +1,46 @@
 #!/usr/bin/env Rscript
 
-library(bayesplot)
 library(cmdstanr)
 library(dplyr)
+library(fc)
+library(yaml)
 
 options(mc.cores = parallel::detectCores() - 1)
 
+input <- commandArgs(trailingOnly = T)
+schema_file <- if (length(input) == 0) "models/frozen_bin_all.yml" else input
+
+stopifnot(file.exists(schema_file))
+
+schema <- read_yaml(schema_file)
+info("Running model: %s", schema$name)
+sprintf("posteriors/%s", schema$name) |> dir.create(showWarnings = F, recursive = T)
+
 df <- readRDS("./data/model_data.rds") |> ungroup()
 
-X <- select(df, incompatibility, cold_war, duration, max_intensity) |>
+###
+# Set the Unit of analysis
+if (schema$episodes != "high_intensity")
+    df <- filter(df, cumulative_intensity == 1)
+
+###
+# Select control variables - currently this is static across all models
+X <- select(df, censored, episode_duration, recur, max_intensity, incompatibility) |>
+    mutate(episode_duration = log(episode_duration)) |>
     data.matrix()
 
 cases <- rowSums(is.na(X)) == 0
 
-treatments <- select(df, ext_sup_s_state_max_5y, ext_sup_s_rebel_max_5y) |>
-    mutate(interaction = ext_sup_s_state_max_5y * ext_sup_s_rebel_max_5y) |>
-    data.matrix()
+###
+# Select the treatment option
+state_sup <- paste0("ext_sup_s_state_", schema$treatment)
+rebel_sup <- paste0("ext_sup_s_rebel_", schema$treatment)
 
+treatments <- select(df, {{ state_sup }}, {{ rebel_sup }}) |>
+    mutate(interaction = .data[[{{ state_sup }}]] * .data[[{{ rebel_sup }}]])
+
+###
+# Assemble model input
 data <- list(n = sum(cases),
              k = ncol(treatments),
              T = treatments[cases, ],
@@ -24,10 +48,15 @@ data <- list(n = sum(cases),
              X = X[cases, ],
              n_countries = n_distinct(df$gwno_a),
              country_id = as.factor(df$gwno_a[cases]) |> as.integer(),
-             y = df$frozen[cases])
-str(data)
+             y = df$strict_frozen[cases])
+
[…]

### 2023-10-23 R/models.R
Fix default model name

@@ -11,7 +11,7 @@ outcomes <- c("frozen", "strict_frozen")
 units <- c("all", "high_intensity")
 
 df <- expand.grid("outcome" = outcomes, "treatment" = treatments, "episodes" = units) |>
-    mutate(name = paste0(outcome, "_", treatment, "_", episodes))
+    mutate(name = paste0(outcome, "-", treatment, "-", episodes))
 
 for (row in 1:nrow(df)) {
     info("Writing model profile %s", df[row, "name"])

### 2023-10-23 R/probit.R
BUGFIX high intensity selection for model data

@@ -8,7 +8,7 @@ library(yaml)
 options(mc.cores = parallel::detectCores() - 1)
 
 input <- commandArgs(trailingOnly = T)
-schema_file <- if (length(input) == 0) "models/frozen_bin_all.yml" else input
+schema_file <- if (length(input) == 0) "models/frozen-bin-all.yml" else input
 
 stopifnot(file.exists(schema_file))
 
@@ -20,7 +20,7 @@ df <- readRDS("./data/model_data.rds") |> ungroup()
 
 ###
 # Set the Unit of analysis
-if (schema$episodes != "high_intensity")
+if (schema$episodes == "high_intensity")
     df <- filter(df, cumulative_intensity == 1)
 
 ###
@@ -37,18 +37,21 @@ state_sup <- paste0("ext_sup_s_state_", schema$treatment)
 rebel_sup <- paste0("ext_sup_s_rebel_", schema$treatment)
 
 treatments <- select(df, {{ state_sup }}, {{ rebel_sup }}) |>
-    mutate(interaction = .data[[{{ state_sup }}]] * .data[[{{ rebel_sup }}]])
+    mutate(interaction = .data[[{{ state_sup }}]] * .data[[{{ rebel_sup }}]]) |>
+    data.matrix()
 
 ###
 # Assemble model input
 data <- list(n = sum(cases),
              k = ncol(treatments),
              T = treatments[cases, ],
+             interaction_id = which(colnames(treatments) == "interaction"),
              m = ncol(X),
              X = X[cases, ],
              n_countries = n_distinct(df$gwno_a),
              country_id = as.factor(df$gwno_a[cases]) |> as.integer(),
              y = df$strict_frozen[cases])
+str(data)
 
 sprintf("posteriors/%s/model_input.RData", schema$name) |> save.image()
 

### 2023-10-23 paper.qmd
Add parameter and AME plots to manuscript

@@
 ---
-title: When Civil Conflicts Freeze
+title: "From Active Hostilities to Frozen Stalemates: Assessing the Role of External Support in Civil Conflict Dynamics"
 author: J.L.A. Krusell
@@
 ```{r setup, include = F}
-library(bayesplot)
 library(dplyr)
@@
 library(ggthemes)
+library(ggdist)
+library(latex2exp)
 library(knitr)
+library(patchwork)
+library(tidyr)
 
@@
 
-theme_set(theme_tufte())
+theme_set(theme_tufte(base_family = "TeX Gyre Pagella", base_size = 10))
 ```
@@
 ```{r}
-fit <- readRDS("./posteriors/probit.rds")
+#| fig.cap: Posterior estimates for \delta from a single variate probit model
+fit <- readRDS("./posteriors/frozen-bin-all/model_output.rds")
 
-fit$draws("delta") |> mcmc_areas() +
+delta <- fit$draws("delta", format = "df") |>
+    select(matches("delta")) |>
+    pivot_longer(everything(), names_to = "variable")
+
+ggplot(delta, aes(x = value, y = variable, fill = variable)) +
+    stat_halfeye(alpha = 0.8, show.legend = F) +
+    geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.2) +
+    theme(axis.title.x = element_blank(),
+          axis.title.y = element_blank()) +
     scale_y_discrete(labels = c("delta[1]" = "Ext. Sup. to Gov",
@@
 ```
+
+```{r}
+#| fig.cap: Average marginal effects for the three treatment conditions on the predicted probability of a frozen conflict onset.[^ame]
+ame <- fit$draws("ame", format = "df") |>
+    select(matches("ame")) |>
+    pivot_longer(everything(), names_to = "variable")
+
+ggplot(ame, aes(x = value, y = variable, fill = variable)) +
+    stat_eye(alpha = 0.8, show.legend = F) +
+    geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.2) +
+    theme(axis.title.x = element_blank()) +
+    labs(x = parse(text = "bold('AME '*Delta)")) +
+    scale_y_discrete(labels = c("ame[1]" = "Ext. Sup. Gov Only",
+                                "ame[2]" = "Ext. Sup. Rebel Only",
+                                "ame[3]" = "Ext. Sup. to Both")) +
+    coord_flip()
+```
+
+[^ame]: Hello World
 
@@
 # References
+
+::: {#refs}
+:::
+
+\newpage
+
+# Appendix
+
+## Robust Checks
+
+### Observational Probit Models
+
+```{R}
+#| fig-height: 7
+dirs <- list.files("posteriors", full.names = T)
+
+fits <- lapply(dirs, function(d) {
+    names <- strsplit(basename(d), "-")[[1]]
+    fit <- file.path(d, "model_output.rds") |> readRDS()
+
+    fit$summary("delta", ~quantile(.x, probs = c(0.05, 0.1, 0.5, 0.9, 0.95))) |>
+        rename(median = `50%`,
+               lower90 = `5%`,
+               upper90 = `95%`,
+               lower80 = `10%`,
+               upper80 = `90%`) |>
+        mutate(outcome = names[1],
+               treatment = names[2],
+               episodes = names[3])
+})
+
+full.df <- bind_rows(fits) |>
+    mutate(group = as.factor(outcome))
+
+fn <- function(target) {
+    sub.df <- filter(full.df, variable == target)
+    ggplot(sub.df, aes(x = treatment, y = median, color = outcome)) +
+        geom_point(size = 1.5, position = position_dodge(width = 0.25)) +
+        geom_errorbar(aes(ymin = lower90, ymax = upper90), position = position_dodge(0.25), width = 0) +
+        geom_errorbar(aes(ymin = lower80, ymax = upper80), position = position_dodge(0.25), size = 1.25, width = 0) +
+        facet_grid(~episodes) +
+        geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.2) +
+        theme(axis.title.x = element_blank(),
+              axis.title.y = element_blank())
+
+}
+
+p1 <- fn("delta[1]")
+p2 <- fn("delta[2]")
+p3 <- fn("delta[3]")
+
+#    gov_delta <- lapply(fits, \(x) x[, 1]) |> do.call(cbind, args = _)
+#reb_delta <- lapply(fits, \(x) x[, 2]) |> do.call(cbind, args = _)
+#interaction_delta <- lapply(fits, \(x) x[, 3]) |> do.call(cbind, args = _)
+#
+#colnames(gov_delta) <- 1:ncol(gov_delta)
+#p1 <- mcmc_intervals(gov_delta) + coord_flip()
+#
+#
+#colnames(reb_delta) <- 1:ncol(reb_delta)
+#p2 <- mcmc_intervals(reb_delta) + coord_flip()
+#
+#
+#colnames(interaction_delta) <- 1:ncol(interaction_delta)
+#p3 <- mcmc_intervals(interaction_delta) + coord_flip()
+
+p1 / p2 / p3
+```

### 2023-10-30 R/dataset.R
Rename fc package to fc.utils
Avoid potential clashes with CRAN packages

@@ -4,7 +4,7 @@
 ###
 
 library(dplyr)
-library(fc)
+library(fc.utils)
 library(readxl)
 
 adj <- read.csv("./data/dataset/adjusted_conflict_candidates.csv") |>

### 2023-10-30 R/merge.R

@@ -1,7 +1,7 @@
 #!/usr/bin/env Rscript
 
 library(dplyr)
-library(fc)
+library(fc.utils)
 library(haven)
 library(readxl)
 library(tidyr)

### 2023-10-30 R/models.R

@@ -1,7 +1,7 @@
 #!/usr/bin/env Rscript
 
 library(dplyr)
-library(fc)
+library(fc.utils)
 library(yaml)
 
 dir.create("models", showWarnings = F)

### 2023-10-30 R/probit.R

@@ -2,7 +2,7 @@
 
 library(cmdstanr)
 library(dplyr)
-library(fc)
+library(fc.utils)
 library(yaml)
 
 options(mc.cores = parallel::detectCores() - 1)

### 2023-10-30 lib/fc.utils/R/fc.utils.R

@@ -1,8 +1,8 @@
-#' fc
+#' fc.utils
 #'
 #' Utility functions used both in the construction of the fc-dataset and
 #' analysis of frozen conflict onsets.
 #'
 #' @docType package
-#' @name fc
+#' @name fc.utils
 NULL

### 2023-10-30 R/probit.R
Use cmdstanr data file for preserving model input
cmdstanr already saves the model input to disk as json in a temporary
directory. Move to 'posteriors/' and also separately save the variable
labels since the column names get dropped during serialization.

@@ -8,7 +8,7 @@ library(yaml)
 options(mc.cores = parallel::detectCores() - 1)
 
 input <- commandArgs(trailingOnly = T)
-schema_file <- if (length(input) == 0) "models/frozen-bin-all.yml" else input
+schema_file <- if (!exists("input") || length(input) == 0) "models/frozen-bin-all.yml" else input
 
 stopifnot(file.exists(schema_file))
 
@@ -53,13 +53,22 @@ data <- list(n = sum(cases),
              y = df$strict_frozen[cases])
 str(data)
 
-sprintf("posteriors/%s/model_input.RData", schema$name) |> save.image()
-
 mod <- cmdstan_model("./stan/probit.stan")
 fit <- mod$sample(data = data)
 
+# Treatment coefficients
 fit$summary("delta")
 
+fit$save_data_file(file.path("posteriors", schema$name), timestamp = F, random = F)
+
+# cmdstanr won't save the column names when serializing the data
+# object to json.
+variables <- c("alpha", sprintf("delta[%d]", 1:ncol(treatments)),
+                sprintf("beta[%d]", 1:ncol(X)))
+names(variables) <- c("intercept", colnames(treatments), colnames(X))
+
+file.path("posteriors", schema$name, "labels.rds") |> saveRDS(variables, file = _)
+
 # Save, save, save!
-sprintf("posteriors/%s/model_output.rds", schema$name) |>
+sprintf("posteriors/%s/fit.rds", schema$name) |>
     fit$save_object()

### 2023-10-30 paper.qmd
Rename fc package to fc.utils
Avoid potential clashes with CRAN packages
Use modelsummary for regression tables

--- 2023-10-19 paper.qmd
@@
 
+[//]: # Setup R environment
 ```{r setup, include = F}
-library(bayesplot)
 library(dplyr)
@@
 library(ggthemes)
+library(ggdist)
+library(modelsummary)
+library(kableExtra)
 library(knitr)
+library(patchwork)
+library(tidyr)
+
+# Project specific utility package
+library(fc.utils)
 
@@
 
+Descriptive statistics split between positive and negative outcomes
+
 ### Descriptive Statistics
@@
 ```{r}
-fit <- readRDS("./posteriors/probit.rds")
+#| fig.cap: Posterior estimates for \delta from a single variate probit model
+fit <- readRDS("./posteriors/frozen-bin-all/fit.rds")
 
-fit$draws("delta") |> mcmc_areas() +
+delta <- fit$draws("delta", format = "df") |>
+    select(matches("delta")) |>
+    pivot_longer(everything(), names_to = "variable")
+
+ggplot(delta, aes(x = value, y = variable, fill = variable)) +
+    stat_halfeye(alpha = 0.8, show.legend = F) +
+    geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.2) +
+    theme(axis.title.x = element_blank(),
+          axis.title.y = element_blank()) +
     scale_y_discrete(labels = c("delta[1]" = "Ext. Sup. to Gov",
@@
 # References
+
+::: {#refs}
+:::
+
+\newpage
+
+# Appendix
+
+## Robust Checks
+
+### Observational Probit Models
+
+```{R}
+models <- list.dirs("posteriors", recursive = F)
+stopifnot(length(models) != 0)
+
+fits <- lapply(models, \(dir) file.path(dir, "fit.rds") |> readRDS())
+names(fits) <- basename(models)
+```
+
+```{R}
+#| fig-height: 7
+summary.df <- lapply(fits, function(fit) {
+    post <- post_summarise(fit, "delta", rename = T, model_info = T,
+                           probs = c(0.025, 0.17, 0.5, 0.83, 0.975)) |>
+        rename(median = `50%`,
+               lower95 = `2.5%`,
+               upper95 = `97.5%`,
+               lower66 = `17%`,
+               upper66 = `83%`)
+}) |> bind_rows()
+
+summary.df$variable <- sub("_(prop|bin.*)$", "", summary.df$variable)
+
+ggplot(summary.df, aes(treatment_aggregation, median, color = outcome)) +
+    geom_point(size = 1.5, position = position_dodge(width = 0.25)) +
+    geom_errorbar(aes(ymin = lower95, ymax = upper95), position = position_dodge(0.25), width = 0) +
+    geom_errorbar(aes(ymin = lower66, ymax = upper66), position = position_dodge(0.25), linewidth = 1.25, width = 0) +
+    geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.2) +
+    facet_grid(variable ~ episodes)
+
+```
+
+```{R}
+models <- list.dirs("posteriors", recursive = F)
+stopifnot(length(models) != 0)
+
+fits <- lapply(models, \(dir) file.path(dir, "fit.rds") |> readRDS())
+
+frozen <- fits[grepl("/frozen", models)]
+strict_frozen <- fits[grepl("/strict_frozen", models)]
+
+modelsummary(frozen, pars = c("alpha", "delta", "beta"),
+             statistic = "conf.int", title = "Frozen Conflicts") |>
+    add_header_above(c(" " = 1, "All Episodes" = 3, "High Intensity Only" = 3)) |>
+    kable_styling(font_size = 8, full_width = F) |>
+    landscape()
+```
+
+```{R}
+modelsummary(strict_frozen, pars = c("alpha", "delta", "beta"),
+             statistic = "conf.int", title = "Frozen Conflicts Lasting >2 Years") |>
+    add_header_above(c(" " = 1, "All Episodes" = 3, "High Intensity Only" = 3)) |>
+    kable_styling(font_size = 8, full_width = F) |>
+    landscape()
+```

### 2023-10-30 lib/fc.utils/R/modelsummary.R

@@ -1,0 +1,60 @@
+#' @export
+model_info <- function(s) UseMethod("model_info", s)
+
+#' @export
+model_info.character <- function(s) {
+    v <- strsplit(s, "-")[[1]]
+    if (length(v) != 3 || any(nchar(v) == 0))
+        stop(sprintf("Misformed model string? %s", s))
+
+    list(outcome = v[1], treatment_aggregation = v[2], episodes = v[3])
+}
+
+#' @importFrom parameters model_parameters
+#' @export
+model_parameters.CmdStanMCMC <- function(model, pars = c(), ...) {
+    # Note, modelsummary is calling `parameters` with do.call with
+    # args as `list(model = model, ...)`. If the first function
+    # argument for our method is called anything besides `model`, it
+    # will silently fail because modelsummary decided to silence all
+    # error messages...
+    draws <- model$draws(pars, format = "df")
+
+    # This is a hack, but we need to recover the original variable names
+    # in the right order
+    labels <- dirname(model$data_file()) |> file.path("labels.rds") |> readRDS()
+
+    args <- append(list(.x = draws), labels)
+    draws <- do.call(posterior::rename_variables, args)
+
+    parameters::parameters(draws, ...)
+}
+
+#' @export
+post_summarise <- function(x, ...) UseMethod("post_summarise", x)
+
+#' @export
+post_summarise.CmdStanFit <- function(x, pars, model_info = F, rename = F, probs = c(0.05, 0.5, 0.95)) {
+    draws <- x$draws(pars)
+
+    if (isTRUE(rename)) {
+        f <- dirname(x$data_file()) |> file.path("labels.rds")
+        if (!file.exists(f))
+            stop("No labels found for model")
+
+        labels <- readRDS(f)
+        args <- append(list(.x = draws), labels[labels %in% posterior::variables(draws)])
+        draws <- do.call(posterior::rename_variables, args)
+    }
+
+    df <- posterior::summarise_draws(draws, ~stats::quantile(.x, probs))
+
+    if (isTRUE(model_info)) {
+        # This is also a horrible hack
+        model <- dirname(x$data_file()) |> basename()
+        li <- model_info(model)
+        df <- cbind(df, li)
+    }
+
+    return(df)
+}

### 2023-10-31 paper.qmd
Avoid kable for docx output support

@@
 library(modelsummary)
-library(kableExtra)
 library(knitr)
@@
 ```{R}
-models <- list.dirs("posteriors", recursive = F)
-stopifnot(length(models) != 0)
+regtable <- function(fits, ...) {
+    names(fits) <- NULL
+    modelsummary(fits, pars = c("delta", "beta", "alpha"), statistic = "conf.int", ...)
+}
 
-fits <- lapply(models, \(dir) file.path(dir, "fit.rds") |> readRDS())
-
-frozen <- fits[grepl("/frozen", models)]
-strict_frozen <- fits[grepl("/strict_frozen", models)]
-
-modelsummary(frozen, pars = c("alpha", "delta", "beta"),
-             statistic = "conf.int", title = "Frozen Conflicts") |>
-    add_header_above(c(" " = 1, "All Episodes" = 3, "High Intensity Only" = 3)) |>
-    kable_styling(font_size = 8, full_width = F) |>
-    landscape()
+m <- c("frozen-bin-all", "strict_frozen-bin-all",
+       "frozen-bin-high_intensity", "strict_frozen-bin-high_intensity")
+regtable(fits[m], title = "Binary Treatments")
 ```
@@
 ```{R}
-modelsummary(strict_frozen, pars = c("alpha", "delta", "beta"),
-             statistic = "conf.int", title = "Frozen Conflicts Lasting >2 Years") |>
-    add_header_above(c(" " = 1, "All Episodes" = 3, "High Intensity Only" = 3)) |>
-    kable_styling(font_size = 8, full_width = F) |>
-    landscape()
+m <- c("frozen-bin_5y-all", "strict_frozen-bin_5y-all",
+       "frozen-bin_5y-high_intensity", "strict_frozen-bin_5y-high_intensity")
+regtable(fits[m], title = "Binary Treatments - Last 5 Years")
 ```
+
+```{R}
+m <- c("frozen-prop-all", "strict_frozen-prop-all",
+       "frozen-prop-high_intensity", "strict_frozen-prop-high_intensity")
+regtable(fits[m], title = "Proportion Treatments")
+```

### 2023-11-08 doc/coding-protocol.qmd
Move dataset docs
Manuscript notes

--- 2023-10-11 coding-protocol.qmd
@@
 ---
+
+TODO: manual coding, specific criteria for determining when an actor ceases to exist
+TODO: how to handle actor fragmentation? Ex: Sendero Luminoso
+TODO: Be specific about sources that will be used, reference list?
 
@@
 ```{r echo = F}
-candidates <- read.csv("./data/dyadic_candidates.csv")
-episodes <- read.csv("./data/dyadic_episodes.csv")
+candidates <- read.csv("../data/dyadic_candidates.csv")
+episodes <- read.csv("../data/dyadic_episodes.csv")
 ```

### 2023-11-08 paper.qmd
Manuscript notes

@@
 library(dplyr)
+library(flextable)
 library(ggplot2)
@@
 
-# Empirics
+**Motivation**
 
-To empirically assess whether external state support for conflict
-actors affects the probability of a civil conflict episode terminating
-as a frozen conflict a series of bayesian probit models were run
-using original data on frozen civil conflicts. A mixed identification strategy was employed by both conditioning on relevant covariates and instrumenting external support with UN roll call votes.
+Contribution: (*conceptual* and *empirical*)
 
-In what follows, I will first describe the dependent and independent variables, including operationalizations and data sources.
-I will then present the model specification and results from a conventional probit regression followed by the instrumental variable analysis.
+- *Original Data*, conceptual/descriptive claim about prevalence of frozen conflicts
+- Novel empirical analysis, provide a more nuanced analysis of *conflict termination* and contributes to our understanding of how external support can shape the dynamics of conflicts
 
-## Data
+The paper will argue that there exists both a theoretical and empirical gap.
+Studies on conflict termination have generally considered military victory and peace agreements as the only outcome.
+I further disaggregate the notion of termination to cover also the situation where conflicts are not conclusively termination and instead become "frozen."
 
-### Defining Frozen Conflicts
+In doing so, I show that frozen conflicts are more prevalent than thought and conceptually can and should cover a greater range of cases than just /de facto/ states.
+I also distinguish the concept of frozen conflicts from the related notions of stalemates and protracted conflicts (and maybe intractable rivalries?).
+
+I provide a novel quantitative analysis on the likelihood of a conflict becoming frozen.
+Specifically, I examine the effect of support by external state actors to belligerents in the course of a conflict episode.
+This will contribute both to our understanding of different types of conflict termination as well as the ways in which external support can shape the dynamics of conflicts.
+
+- Potential shortcoming: what I am not doing is taking into account the preferences of external supporters and motivation for offering support.
+Nor am I distinguishing between different external support types.
+
+**Why should we care about frozen conflicts?**
+
+# Defining Frozen Conflicts
 
@@
 More importantly, enduring rivalries were originally conceptualized in terms of inter-state relations, while the definition of frozen conflicts here can be extended to the intra-state case as well.
+
+# Theory
+
+*The fundamental question*: why would actors accept a frozen conflict over a comprehensive peace agreement?
+
+This mirrors the "why fight" question underlying bargaining theory where fighting is assumed to be costly such that with perfect information rational actors should prefer a "bargained solution." Is there a similar parallel to frozen conflicts?
+
+What is the cost of perpetuating a frozen conflict?
+How does this play in with the traditional, bargaining theory approach to conflict termination.
+Namely, actors become willing to terminate a conflict as information is revealed through fighting and the payoff from victory diminishes, based on the subjective probability of winning and the accumulated costs both past and future.
+
+Is there a situation in which rational actors would prefer to leave a conflict unresolved with the future threat of violent re-escalation?
+Clearly, the rebel side faces a credible commitment dilemma in that they can never guarantee that the government won't renege in the future.
+
+Literature on ceasefires (ex Clayton and Stitcher (2021)) emphasize the strategic logic underlying actors' decisions to accept or reject a ceasefire.
+More specifically, actors may be willing to enter into a ceasefire either as a signaling mechanism to demonstrate an intent to peacefully resolve the conflict or as a delaying mechanism to allow them to temporarily pause fighting in order to rebuild their strength.
+The decision to accept a frozen conflict, despite the ongoing costs, may be similar to this latter mechanism.
+Conflict actors may not be willing to completely abandon their war aims; however, due to a variety of factors are unable to win an outright victory.
+By letting the conflict become frozen, either through a ceasefire, partial peace agreement, or a stalemate, actors can temporarily halt the ongoing costs of fighting in hopes of a future favourable shift in the balance of power.
+This probably requires actors to have long time horizons and be willing to burden the costs of not fully resolving the conflict.
+Specifically, letting a conflict become frozen means that there is always a threat of re-escalation so actors need to continue to commit to military spending, while suffer ongoing economic costs from lost territory.
+
+"...Leeds (1999) and Walter (2002) suggest that once negotiation begins because of a hurting stalemate, any warring group would prefer to have those negotiations be successful rather than to return to their group members without a settlement." --- from Park (2015)
+
+This is a potential theoretical contribution to the bargaining literature in that I could argue a comprehensive peace agreement does not necessarily always result from a hurting stalemate.
+Letting conflicts freeze could be one way that actors ease the costs of a hurting stalemate without being forced to commit to a settlement.
+This could be especially attractive in scenarios where there is such a high degree of mutual distrust, with few international guarantees for a negotiated settlement, where credible commitment problems could lead to a frozen conflict.
+Especially when actors receive external support, which a) is more likely to lead to the type of stalemate scenario where neither side can outright win militarily and b) allow actors to sustain the military threat required for a frozen conflict to perpetuate.
+
+Note, extensive literature arguing that external actors also increase the number of veto players which decrease the likelihood of a negotiated settlement (ex Cunningham 2006).
+
+- Differentiating the theoretical mechanisms for the different treatment conditions: no external support, both sides receive external support, government only, rebel only
+
+- States tend to be fairly conservative about sovereignty claims (see Hutchinson f.ex.). The best that separatists can hope for is a *de facto* state.
+
+- Big difference compared to ceasefires, frozen conflicts are not /a priori/ agreed upon by belligerents.
+Even if an actor prefers a conflict to become frozen, there is no guarantee that the opposing side won't attempt a re-escalation at any time.
+
+  - Even if actors prefer a frozen conflict over a comprehensive peace agreement; why not codify the status quo and formalized a ceasefire?
+(I think the key here is to emphasize that actors **cannot** win militarily, in other words they are stalemated into a frozen conflicts and refuse to give up their war aims.
+The question then is not, why do conflict actors prefer frozen conflicts, but rather in the face of a frozen conflict, why do conflict actors refuse a negotiated settlement?)
+
+# Empirics
+
+To empirically assess whether external state support for conflict actors affects the probability of a civil conflict episode terminating as a frozen conflict a series of bayesian probit models were run using original data on frozen civil conflicts.
+A mixed identification strategy was employed by both conditioning on relevant covariates and instrumenting external support with UN roll call votes.
+
+In what follows, I will first present the model specification and results from a probit regression predicting the onset of a frozen conflict.
+I will then decrease the dependent and independent variables, including operationalizations and data sources.
+
+## Probit Model
+
+TODO: eventually I would like to move towards a dyadic analysis, rather than conflict level
+
+- Unit of analysis: conflict episodes, ie consecutive years of active fighting (above UCDP threshold of 25 battle deaths)
+- Binary outcome, binary treatment, while conditioning on a number of relevant covariates
+
+\begin{align*}
+y_i^\star & = \alpha + X_i' \beta + T_i' \delta + \epsilon_i, \; \epsilon_i \sim \text{N}(0, \sigma^2) \\
+y_i & = \begin{cases}
+1 & \text{if } y_i^\star > 0 \\
+0 & \text{otherwise}
+\end{cases}
+\end{align*}
+
+- Estimate with Stan, weakly informative priors, $\delta \sim \text{Normal}(1, 1)$
+
+## Data
 
@@
 
-UCDP external support dataset
+UCDP external support dataset, binary variable for whether an external
+state actor offered either fungible or non-fungible support at least
+once to either the rebels or government side in a conflict episode.
+
+- Include interaction term, three treatment conditions:
+
+1. No support (control group)
+2. Gov. support only
+3. Reb. support only
+4. Both support
 
@@
 
-- Conflict Intensity
-- Conflict Duration
-- Incompatibility?
-- e_gdppc
-- e_pop
-- CINC
+- Episode Intensity
+- Episode Duration
+- Incompatibility? (ie separatist conflict over territory)
+- Recurrent episode
+- Right censored
 
@@
 
-\begin{align*}
-y_i^\star & = \alpha + X_i' \beta + T_i' \delta + \epsilon_i, \; \epsilon_i \sim \text{N}(0, \sigma^2) \\
-y_i & = \begin{cases}
-1 & \text{if } y_i^\star > 0 \\
-0 & \text{otherwise}
-\end{cases}
-\end{align*}
+### Results
 
-### Results
+Figure 1 shows the parameter estimates for $\delta$.
 
@@
 ```
+
+Figure 2 shows the average marginal effects, *i.e.* the change in the posterior predicted probability between each treatment condition and the control condition averaged across the sample.
+
+\begin{equation}
+\text{AME}_j = \sum_{i=1}^N \Phi(\alpha + X_i' \beta + \delta_i) - \Phi(\alpha + X_i' \beta)
+\end{equation}
 
@@
 
-## Robust Checks
+## Robustness Checks
 
-### Observational Probit Models
+Different combinations of the following...
+
+- Aggregation of the treatments: binary treatment, binary treatment in the last 5 years of the conflict episode, and proportional treatment.
+The latter is calculated as the proportion of conflict-years in an episode where one side received external support
+
+- All conflict episodes and only high intensity episodes --- try to distinguish between low-level insurgencies and actual civil wars
+
+- Normal definition for frozen conflicts and a strict definition that requires a duration of at least 2 years in order to exclude temporary one-year pauses in fighting that may be more the result of UCDP's coding decision rather than representing actual frozen conflict.
 

### 2023-11-28 slides/gues3s-23-11-27.qmd
guess3s presentation

---
title: "Frozen Civil Conflicts"
subtitle: "The Role of External State Support"
author:
    - name: J.L.A. Krusell
institute: "Department of Political Science"
format:
    revealjs:
        theme: solarized
        footer: "Department of Political Science --- University of Gothenburg"
suppress-bibliography: true
---

```{r setup, include F}
library(dplyr)
library(fc.utils)
library(ggdist)
library(ggplot2)
library(ggthemes)
library(tidyr)

theme_set(theme_solarized())
```

## Defining Frozen Conflicts

"...protracted, post-war conflict process, characterized by the absence of stable peace...[where] core issues between the opposing sides remain unresolved, the dispute is in the forefront of mutual relations, and there is a looming threat of the renewal of violence." [@smetana_between_2019]

## Research Question {auto-animate=true}

Why do some civil conflicts become frozen?

## Research Question {auto-antimate=true}

To what extent does external state support affect the likelihood of a civil conflict becoming frozen?

. . .

::: {.absolute bottom=50%}
$\text{External State Support} \rightarrow \text{Frozen Civil Conflict Onset}$
:::

## Methods

Two strategies: bayesian probit model and instrumental variable analysis

. . .

- Unit of Analysis: Civil conflict episodes [@gleditsch_armed_2002]

- Outcome: Frozen conflict onset

- Treatment:
  1. External state support for *rebels*
  2. External state support for *government*
  3. Interaction

## Measuring Frozen Conflicts

Civil conflict episodes were coded as becoming frozen conflicts if the following conditions were satisfied:

- Episode did **not** terminate with either a one-sided military victory or comprehensive peace agreement

- Non-state actor continued to exist for at least 5 years following episode termination without renouncing political violence

## Measuring Frozen Conflicts

248 civil conflict episodes spanning 1975-2017, of which 69 became frozen

```{r}
dataset <- readRDS("../data/model_data.rds")

filter(dataset, frozen == 1) |>
ggplot(aes(year)) +
    geom_bar() +
    xlab("Year") +
    ylab("Frozen Civil Conflict Onsets")
```

## Measuring Frozen Conflicts {.smaller}

```{r}
conditions <- dataset |>
    mutate(treatment = case_when(ext_sup_s_rebel_bin == 1 & ext_sup_s_state_bin == 1 ~ "Ext. Sup. to Both",
                                 ext_sup_s_rebel_bin == 1 & ext_sup_s_state_bin == 0 ~ "Ext. Sup. Rebel Only",
                                 ext_sup_s_rebel_bin == 0 & ext_sup_s_state_bin == 1 ~ "Ext. Sup. Gov Only",
                                 ext_sup_s_rebel_bin == 0 & ext_sup_s_state_bin == 0 ~ "Ext. Sup. to Neither"))

m <- table(conditions$frozen, conditions$treatment)
rownames(m) <- c("Not Frozen", "Frozen")
knitr::kable(m)
```

## Additional Covariates

:::{.incremental}
- Conflict dynamics: duration, intensity, recurrence
- Incompatibility
- Ongoing intra/inter-state conflict
- Cold war
- Censoring
:::

## Probit Model

\begin{align*}
y^\star_i & = \alpha + X_i' \beta + T_i' \delta_{k_i} + \gamma_{c_i} + \epsilon_i, \: \epsilon_i \sim \text{N}(0, 1) \\
y_i & = \begin{cases}
1 & \text{if } y_i^\star > 0 \\
0
\end{cases}
\end{align*}

. . .

:::: {.columns}

::: {.column width="30%"}
\begin{align*}
\alpha & \sim \text{N}(0, 2.5) \\
\beta & \sim \text{N}(0, 1) \\
\end{align*}
:::

::: {.column width="40%"}
\begin{align*}
\delta_j & \sim \text{N}(\mu_j, \tau_j) \\
\mu_j & \sim \text{N}(0, 1) \\
\tau_j & \sim \text{N}_{+}(0, 1)
\end{align*}
:::

::: {.column width="30%"}
\begin{align*}
\gamma_c & \sim \text{N}(0, \sigma) \\
\sigma & \sim \text{N}_{+}(0, 1)
\end{align*}
:::

:::

## Parameter Estimates

```{r}

fit <- readRDS("../posteriors/frozen-bin-all/fit.rds")

delta <- fit$draws("delta", format = "df") |>
    select(matches("delta")) |>
    pivot_longer(everything(), names_to = "variable")

ll <- extract_digits(delta$variable)
delta$variable <- sprintf("delta[%s]", sapply(ll, `[`, 1))
delta$group <- ifelse(sapply(ll, `[`, 2) == 1, "Government Incompatibility",
                      "Territorial Incompatibility")

ggplot(delta, aes(x = value, y = variable, fill = variable)) +
    stat_halfeye(alpha = 0.8, show.legend = F) +
    geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.2) +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank()) +
    scale_y_discrete(labels = c("delta[1]" = "Ext. Sup. to Gov",
                                "delta[2]" = "Ext. Sup. to Rebel",
                                "delta[3]" = "Interaction")) +
    facet_grid(. ~ group) +
    scale_fill_solarized()

```

## Average Marginal Effects {.smaller}

Three treatment conditions:

#. External state support for the **government only**
#. External state support for the **rebels only**
#. External state support for **both sides**

\begin{equation}
\text{AME}_{jk} = \frac{1}{N_k} \sum_{i=1}^{N_k} \Phi(\alpha + X_i' \beta + W_j' \delta_k + \gamma_{c_i}) - \Phi\left(\alpha + X_i' \beta + \gamma_{c_i}\right)
\end{equation}

## Average Marginal Effects

```{r}
ame <- fit$draws("ame", format = "df") |>
    select(matches("ame")) |>
    pivot_longer(everything(), names_to = "variable")

ll <- extract_digits(ame$variable)
ame$variable <- sprintf("ame[%s]", sapply(ll, `[`, 1))
ame$group <- ifelse(sapply(ll, `[`, 2) == 1, "Government Incompatibility",
                    "Territorial Incompatibility")

ggplot(ame, aes(x = value, y = variable, fill = variable)) +
    stat_eye(alpha = 0.8, show.legend = F) +
    geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.2) +
    theme(axis.title.x = element_blank()) +
    labs(x = parse(text = "bold('AME '*Delta)")) +
    scale_y_discrete(labels = c("ame[1]" = "Ext. Sup. Gov Only",
                                "ame[2]" = "Ext. Sup. Rebel Only",
                                "ame[3]" = "Ext. Sup. to Both")) +
    coord_flip() +
    facet_grid(. ~ group) +
    scale_fill_solarized()

```

# Part II

## Instrument

UN General Assembly voting 10 years prior to a conflict onset

. . .

**Pros**:

- No measurement error, unlikely to have a direct effect on civil conflict

. . .

**Cons**:

- Cannot exclude possibility of additional systematic confounders

## I.V. Model

\begin{align*}
T^\star_i & = \alpha_1 + X_i' \beta_1 + Z_i' \lambda + \upsilon_{c_i} + \eta_i \\
y^\star_i & = \alpha_2 + X_i' \beta_2 + T_i' \delta_{k_i} + \gamma_{c_i} + \epsilon_i \\
\eta_i, \epsilon_i & \sim \text{MVN}(0, \Sigma)
\end{align*}

### 2023-11-30 slides/peer-presentation-23-11-30.qmd
Polisci peer presentation

--- 2023-11-28 slides/gues3s-23-11-27.qmd
@@
         footer: "Department of Political Science --- University of Gothenburg"
+        title-slide-attributes:
+            data-background-image: ../assets/LO_GUeng_CMYK_flag.png
+            data-background-size: 10%
+            data-background-position: 2% 0%
 suppress-bibliography: true
@@
 
+## Introduction {.smaller}
+
+*To what extent does external state support affect the likelihood of a civil conflict becoming frozen?*
+
+::: {.incremental}
+
+- Previous research on the link between external support and conflict termination [@mason1996civil; @balch2008third; @karlen2017legacy; @sawyer2017role]
+
+    - Generally assumes two possible outcomes: military victory or peace agreement
+
+    - Very little scholarly work on frozen conflicts
+
+- **My contribution**:
+
+    - Theoretical, extending bargaining theory to the domain of frozen conflicts
+    - Empirical, novel dataset to examine relationship between external support and frozen civil conflict onset
+
+:::
+
 ## Defining Frozen Conflicts
@@
 
-## Research Question {auto-animate=true}
+# Theory
 
-Why do some civil conflicts become frozen?
+## Government Contestation
 
-## Research Question {auto-antimate=true}
+::: {.incremental}
 
-To what extent does external state support affect the likelihood of a civil conflict becoming frozen?
+- External support shifts the balance of power
 
-. . .
+    - Often assumed that the rebel side is the weaker party in civil conflicts
 
-::: {.absolute bottom=50%}
-$\text{External State Support} \rightarrow \text{Frozen Civil Conflict Onset}$
+- Introduces additional veto players that complicates peace negotiations
+
 :::
+
+## Territorial Contestation {.smaller}
+
+::: {.incremental}
+
+**For the rebels**:
+
+- Opportunity to consolidate power, and establish a *de facto* state free from credible committments issues arising from negotiated solutions
+
+    - External support should be decisive in aiding rebels to seize territory
+
+**For the government**:
+
+- Future prospect of forcible reintegration
+
+    - External support extends time horizon and offsets accumulated costs
+
+:::
+
+# Empirics
 
@@
 
-- Outcome: Frozen conflict onset
+- Binary outcome: Frozen conflict onset
 
-- Treatment:
+- Binary treatments:
   1. External state support for *rebels*
@@
 
-## Measuring Frozen Conflicts
+## Measuring Frozen Conflicts {.smaller}
 
@@
 
+:::: {.columns}
+
+::: {.column}
 ```{r}
+#| fig-height: 7
 dataset <- readRDS("../data/model_data.rds")
+frozen <- filter(dataset, frozen == 1) |>
+    mutate(country = sub("Government of ", "", side_a),
+           country = case_when(country == "Russia (Soviet Union)" ~ "Russia",
+                               country == "United Kingdom" ~ "UK",
+                               country == "Congo" ~ "Republic of Congo",
+                               country == "Myanmar (Burma)" ~ "Myanmar",
+                               country == "Bosnia-Herzegovina" ~ "Bosnia and Herzegovina",
+                               country == "Serbia (Yugoslavia)" ~ "Serbia",
+                               country == "DR Congo (Zaire)" ~ "Democratic Republic of the Congo",
+                               T ~ country))
 
+sub.df <- group_by(frozen, country) |> count(frozen)
+
+world <- map_data("world") |>
+    left_join(sub.df, by = c("region" = "country")) |>
+    mutate(n = ifelse(is.na(n), 0, 1) )
+
+ggplot(world, aes(long, lat, map_id = region, fill= n, group = group)) +
+    geom_map(map = world) +
+    theme(legend.position = "none") +
+    scale_color_solarized()
+```
+:::
+
+::: {.column}
+```{r}
+#| fig-height: 7
 filter(dataset, frozen == 1) |>
@@
 ```
+:::
+::::
 
@@
 
-## Probit Model
+# Probit Analysis
+
+## Probit Model {auto-animate=true}
 
@@
 
-. . .
+## Probit Model {auto-animate=true}
+
+\begin{align*}
+y^\star_i & = \alpha + X_i' \beta + T_i' \delta_{k_i} + \gamma_{c_i} + \epsilon_i, \: \epsilon_i \sim \text{N}(0, 1) \\
+y_i & = \begin{cases}
+1 & \text{if } y_i^\star > 0 \\
+0
+\end{cases}
+\end{align*}
+
+\begin{align*}
+\alpha & \sim \text{N}(0, 2.5) \\
+\beta & \sim \text{N}(0, 1) \\
+\end{align*}
+
+## Probit Model {auto-animate=true}
+
+\begin{align*}
+y^\star_i & = \alpha + X_i' \beta + T_i' \delta_{k_i} + \gamma_{c_i} + \epsilon_i, \: \epsilon_i \sim \text{N}(0, 1) \\
+y_i & = \begin{cases}
+1 & \text{if } y_i^\star > 0 \\
+0
+\end{cases}
+\end{align*}
+
+:::: {.columns}
+
+::: {.column}
+\begin{align*}
+\alpha & \sim \text{N}(0, 2.5) \\
+\beta & \sim \text{N}(0, 1) \\
+\end{align*}
+:::
+
+::: {.column}
+\begin{align*}
+\delta_j & \sim \text{N}(\mu_j, \tau_j) \\
+\mu_j & \sim \text{N}(0, 1) \\
+\tau_j & \sim \text{N}_{+}(0, 1)
+\end{align*}
+:::
+
+::::
+
+## Probit Model {auto-animate=true}
+
+\begin{align*}
+y^\star_i & = \alpha + X_i' \beta + T_i' \delta_{k_i} + \gamma_{c_i} + \epsilon_i, \: \epsilon_i \sim \text{N}(0, 1) \\
+y_i & = \begin{cases}
+1 & \text{if } y_i^\star > 0 \\
+0
+\end{cases}
+\end{align*}
 
@@
 
-:::
+::::
 
@@
 
+where $W_j$ is the corresponding treatment vector for treatment condition $j$.
+
 ## Average Marginal Effects
@@
 
-# Part II
+# I.V. Analysis
 
-## Instrument
+## Instrument {.smaller .nostretch}
 
+:::: {.columns}
+
+::: {.column}
 UN General Assembly voting 10 years prior to a conflict onset
 
-. . .
+- Captures foreign policy homophily/heterophily
+:::
+
+::: {.column}
+![](../assets/diagram.png)
+:::
+
+::::
+
+## Instrument {.smaller auto-animate=true}
+
+- Multiple approaches to measurement:
+
+    - Average S-score (Signorino and Ritter 1999):
+
+       \begin{align*}
+        S_{ab} = 1 - \cfrac{\sum | Y_{av} - Y_{bv} |}{V}
+        \end{align*}
+
+        where $Y_{av}$ denotes the recorded vote for country $a$ on resolution $v$ for $v = 1, \ldots V$ within the target time period.
+Then, for a subset of countries, $\mathcal{F}$, relevant to country $a$,
+
+        \begin{align*}
+        Z_a = \frac{1}{|\mathcal{F}|} \sum_{i \in \mathcal{F}} S_{ai}
+        \end{align*}
+
+## Instrument {.smaller auto-animate=true}
+
+- Multiple approaches to measurement:
+
+    - IRT latent scores (Bailey et al 2017):
+
+        \begin{align*}
+        P(Y_{itv} = k) = \Phi(\gamma_{kv} - \beta_v \theta_{it}) - \Phi(\gamma_{k-1,v} - \beta_v \theta_{it})
+        \end{align*}
+
+        where $i = 1, \ldots, N$ now indexes the country. Then,
+
+        \begin{align*}
+        Z_a = \frac{1}{|\mathcal{F}| + |\mathcal{G}|} \sum_{i \in \mathcal{F}, t \in \mathcal{G}} \lVert \theta_{at} - \theta_{it} \rVert
+        \end{align*}
+
+        for the restricted time period $\mathcal{G}$.
+
+## Instrument {.smaller}
+
+:::: {.columns}
+
+::: {.column width="60%"}
+
+UN General Assembly voting
 
@@
 
-- No measurement error, unlikely to have a direct effect on civil conflict
+- Unlikely to have a direct effect on civil conflict
 
-. . .
+:::
+
+::: {.column width="40%"}
+```{r}
+#| fig-height: 4
+#| fig-width: 4
+library(dagitty)
+library(ggdag)
+
+dag <- dagitty("dag {
+Z [pos=\"0,0.5\"];
+T [pos=\"0,0\"];
+Y [pos=\"1,0\"];
+U [pos=\"0.5,-0.5\"];
+Z -> T; T -> Y; U -> Y; U -> T
+}")
+
+ggdag(dag) +
+    theme(panel.grid.major = element_blank(),
+          panel.grid.minor = element_blank(),
+          axis.text = element_blank(),
+          axis.title = element_blank(),
+          axis.ticks = element_blank(),
+          axis.line = element_blank(),
+          panel.background = element_blank())
+```
+:::
+
+::::
+
+## Instrument {.smaller}
+
+:::: {.columns}
+
+::: {.column width="60%"}
+UN General Assembly voting
+
+**Pros**:
+
+- Unlikely to have a direct effect on civil conflict
 
@@
 - Cannot exclude possibility of additional systematic confounders
+- Difficult to instrumentize *both* treatments
+:::
+
+::: {.column width="40%"}
+```{r}
+#| fig-height: 4
+#| fig-width: 4
+library(dagitty)
+library(ggdag)
+
+dag <- dagitty("dag {
+W [pos=\"0.5,1\"];
+Z [pos=\"0,0.5\"];
+T [pos=\"0,0\"];
+Y [pos=\"1,0\"];
+U [pos=\"0.5,-0.5\"];
+Z -> T; T -> Y; U -> Y; U -> T; W -> Z; W -> Y
+}")
+
+ggdag(dag) +
+    theme(panel.grid.major = element_blank(),
+          panel.grid.minor = element_blank(),
+          axis.text = element_blank(),
+          axis.title = element_blank(),
+          axis.ticks = element_blank(),
+          axis.line = element_blank(),
+          panel.background = element_blank())
+```
+:::
+
+::::
 
@@
 \begin{align*}
-T^\star_i & = \alpha_1 + X_i' \beta_1 + Z_i' \lambda + \upsilon_{c_i} + \eta_i \\
-y^\star_i & = \alpha_2 + X_i' \beta_2 + T_i' \delta_{k_i} + \gamma_{c_i} + \epsilon_i \\
+T^\star_i & = \alpha_1 + X_i' \beta_1 + Z_i' \lambda + \gamma_{1c_i} + \eta_i \\
+y^\star_i & = \alpha_2 + X_i' \beta_2 + T_i \delta + \gamma_{2c_i} + \epsilon_i \\
 \eta_i, \epsilon_i & \sim \text{MVN}(0, \Sigma)
 \end{align*}
+
+Assumptions:
+
+1. $\lambda \neq 0$
+2. $Z \perp\!\!\!\perp Y \mid T$
+3. $Z \perp\!\!\!\perp \epsilon \mid T, X$
+
+# Conclusion
+
+. . .
+
+Should the instrument be completely abandoned?

### 2023-11-30 R/merge.R
Create binary indicator for high intensity

@@ -127,11 +127,7 @@ final.df <- filter(full.df, !is.na(ext_sup_s_state)) |>
               ongoing_intrastate = max(ongoing_intrastate) > 0,
               ongoing_conflict = ongoing_interstate | ongoing_intrastate,
               cumulative_intensity = max(cumulative_intensity),
-              max_intensity = max(intensity_level, na.rm = T),
-              avg_intensity = mean(intensity_level, na.rm = T),
-              wavg_intensity = weighted.mean(intensity_level, 1:n(), na.rm = T),
-
-              across(c(v2x_polyarchy, e_gdppc, e_pop, cinc), ~mean(.x, na.rm = T), .names = "{.col}_avg"),
+              high_intensity = max(intensity_level) == 2,
 
               # At anytime during the episode was external support given?
               across(starts_with("ext_"), ~max(.x, na.rm = T), .names = "{.col}_bin"),

### 2023-11-30 R/probit.R

@@ -25,9 +25,9 @@ if (schema$episodes == "high_intensity")
 
 ###
 # Select control variables - currently this is static across all models
-X <- select(df, censored, episode_duration, recur, max_intensity, incompatibility,
+X <- select(df, censored, episode_duration, recur, high_intensity, incompatibility,
             cold_war, ongoing_intrastate, ongoing_interstate) |>
-    mutate(episode_duration = log(episode_duration)) |>
+    mutate(episode_duration = log(episode_duration) |> scale()) |>
     data.matrix()
 
 cases <- rowSums(is.na(X)) == 0

### 2023-11-30 R/probit.R
Split hierarchical and non-hierarchical versions of probit model

@@ -37,6 +37,8 @@ cases <- rowSums(is.na(X)) == 0
 state_sup <- paste0("ext_sup_s_state_", schema$treatment)
 rebel_sup <- paste0("ext_sup_s_rebel_", schema$treatment)
 
+info("Treatment variables: %s and %s", state_sup, rebel_sup)
+
 treatments <- select(df, {{ state_sup }}, {{ rebel_sup }}) |>
     mutate(interaction = .data[[{{ state_sup }}]] * .data[[{{ rebel_sup }}]]) |>
     data.matrix()
@@ -56,8 +58,8 @@ stan_data <- list(N = sum(cases),
                   y = df$strict_frozen[cases])
 str(stan_data)
 
-mod <- cmdstan_model("./stan/probit.stan")
-fit <- mod$sample(data = stan_data, adapt_delta = 0.95)
+mod <- cmdstan_model("./stan/hierarchical_probit.stan")
+fit <- mod$sample(data = stan_data, sig_figs = 3, adapt_delta = 0.95)
 
 # Treatment coefficients
 fit$summary("delta")

### 2023-11-30 R/sbc.R

@@ -14,13 +14,15 @@ rank_statistic <- function(draws, true_value) {
     sum(draws < true_value)
 }
 
-n <- 300
-m <- 2
+n <- 248
+m <- 8
 k <- 3
 
 parameters <- c("alpha", "beta", "gamma", "sigma", "delta", "mu", "tau")
 
-X <- mvrnorm(n, rep(0, m), diag(rep(2, m), nrow = m, ncol = m))
+#X <- mvrnorm(n, rep(0, m), diag(rep(2, m), nrow = m, ncol = m)) |>
+#    apply(2, scale)
+X <- sample(0:1, n * m, replace = T) |> matrix(nrow = n, ncol = m)
 
 n_countries <- 3
 country_id <- sample(1:n_countries, n, replace = T)
@@ -28,7 +30,9 @@ country_id <- sample(1:n_countries, n, replace = T)
 n_conflict_types <- 2
 conflict_id <- sample(1:n_conflict_types, n, replace = T)
 
-treatments <- sample(0:1, n*3, replace = T) |> matrix(nrow = n, ncol = 3)
+treatments <- sample(0:1, n*2, replace = T) |> matrix(nrow = n, ncol = 2)
+treatments <- cbind(treatments, treatments[, 1] * treatments[, 2])
+
 
 data <- list(n = n,
              k = k,
@@ -53,14 +57,14 @@ y_sim <- sim_data$draws("y_sim", format = "matrix")
 # For each simulated outcome, run the model and compute the rank
 # statistic for each parameter between the posterior draws and the
 # true (ie simulated) value
-mod <- cmdstan_model("./stan/probit.stan")
+mod <- cmdstan_model("./stan/hierarchical_probit.stan")
 ranks <- mclapply(1:nrow(y_sim), function(i) {
     stan_data <- data
     stan_data$y <- as.vector(y_sim[i, ])
     stan_data$interaction_id <- 1
 
-    fit <- mod$sample(data = stan_data, iter_sampling = 500, chains = 2, parallel_chains = 2,
-                      adapt_delta = 0.95, thin = 2, refresh = 0)
+    fit <- mod$sample(data = stan_data, sig_figs = 3, iter_sampling = 500, chains = 2,
+                      parallel_chains = 2, adapt_delta = 0.95, thin = 2, refresh = 0)
     mapply(rank_statistic, fit$draws(parameters, format = "data.frame"), pv[i, ])
 })
 

### 2023-11-30 paper.qmd
Update manuscript plots for varying treatment effects

@@
 
+ll <- extract_digits(delta$variable)
+delta$variable <- sprintf("delta[%s]", sapply(ll, `[`, 1))
+delta$group <- sapply(ll, `[`, 2) |> as.factor()
+
 ggplot(delta, aes(x = value, y = variable, fill = variable)) +
-    stat_halfeye(alpha = 0.8, show.legend = F) +
+    stat_halfeye(alpha = 0.8) +
     geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.2) +
@@
                                 "delta[2]" = "Ext. Sup. to Rebel",
-                                "delta[3]" = "Interaction"))
+                                "delta[3]" = "Interaction")) +
+    facet_grid(. ~ group)
+
 ```
@@
 
+ll <- extract_digits(ame$variable)
+ame$variable <- sprintf("ame[%s]", sapply(ll, `[`, 1))
+ame$group <- sapply(ll, `[`, 2) |> as.factor()
+
 ggplot(ame, aes(x = value, y = variable, fill = variable)) +
@@
                                 "ame[3]" = "Ext. Sup. to Both")) +
-    coord_flip()
+    coord_flip() +
+    facet_grid(. ~ group)
 ```
@@
 
-```{R}
+```{R eval = F}
 models <- list.dirs("posteriors", recursive = F)
@@
 
-```{R}
+```{R eval = F}
 #| fig-height: 7
@@
 
-```{R}
+```{R eval = F}
 regtable <- function(fits, ...) {
@@
 
-```{R}
+```{R eval = F}
 m <- c("frozen-bin_5y-all", "strict_frozen-bin_5y-all",
@@
 
-```{R}
+```{R  eval = F}
 m <- c("frozen-prop-all", "strict_frozen-prop-all",

### 2023-11-30 R/dyadic_candidates.R

@@ -5,7 +5,7 @@
 ###
 
 library(dplyr)
-library(fc)
+library(fc.utils)
 library(readxl)
 library(tidyr)
 

### 2023-11-30 lib/fc.utils/R/utils.R

@@ -1,3 +1,10 @@
+#' @export
+extract_digits <- function(x) {
+    m <- gregexpr("\\d", x)
+    regmatches(x, m)
+}
+
+
 #' Find consecutive elements
 #'
 #' Given a NumericVector, return the grouping indices for consecutive

### 2023-11-30 paper.qmd
Manuscript notes

@@
 
+- Frozen conflicts as a suboptimal solution to the contest between an incumbent and challenger.
+
 # Empirics
@@
 ## Probit Model
+
+I model the onset of a frozen civil conflict using a bayesian probit regression where the unit analysis is civil conflict episodes, *i.e.*, contiguous years of active fighting within a conflict.
+The outcome variable, $y \in \{ 1, 2 \}$, is a binary indicator denoting the
 
@@
 
+## List of Frozen Conflicts
+
+## Simulation Based Calibration
+
 ## Robustness Checks

### 2023-12-12 R/merge.R
Avoid the mess with json schemas and simply generate input datasets

@@ -149,4 +149,4 @@ info("%d high intensity episodes, with %d frozen onsets and %d >2 onsets",
      sum(high_intensity$frozen),
      sum(high_intensity$strict_frozen))
 
-saveRDS(final.df, "./data/model_data.rds")
+saveRDS(final.df, "./data/merged_data.rds")

### 2023-12-12 R/models.R

@@ -2,18 +2,60 @@
 
 library(dplyr)
 library(fc.utils)
-library(yaml)
 
-dir.create("models", showWarnings = F)
+dir.create("data/models", showWarnings = F)
+df <- readRDS("./data/merged_data.rds") |> ungroup()
 
+###
+# Create a design matrix with all possible analysis combinations
 treatments <- c("bin", "bin_5y", "prop")
 outcomes <- c("frozen", "strict_frozen")
 units <- c("all", "high_intensity")
+censored <- c("no_censored", "with_censored")
 
-df <- expand.grid("outcome" = outcomes, "treatment" = treatments, "episodes" = units) |>
-    mutate(name = paste0(outcome, "-", treatment, "-", episodes))
+design <- expand.grid("outcome" = outcomes, "treatment" = treatments,
+                      "episodes" = units, "censored" = censored) |>
+    mutate(name = sprintf("%s-%s-%s-%s", outcome, treatment, episodes, censored))
 
-for (row in 1:nrow(df)) {
-    info("Writing model profile %s", df[row, "name"])
-    write_yaml(df[row, ], sprintf("./models/%s.yml", df[row, "name"]))
+for (row in 1:nrow(design)) {
+    info("Model %s", design[row, "name"])
+
+    # Start off by setting the unit of analysis
+    sub.df <- if (design[row, "episodes"] == "high_intensity")
+        filter(df, cumulative_intensity == 1)
+    else
+        df
+
+    # Filter out censored observations if necessary, ie conflicts
+    # beginning before 1975
+    if (design[row, "censored"] == "no_censored")
+        sub.df <- filter(sub.df, censored == 0)
+
+    # Select control variables - this is mostly static across models
+    X <- select(sub.df, censored, episode_duration, recur, high_intensity, incompatibility,
+                cold_war, ongoing_intrastate, ongoing_interstate)
+
+    # If we've dropped censored episodes, no need for binary control
+    # var
+    if (design[row, "censored"] == "no_censored")
+        X <- select(X, -censored)
+
+    X <- mutate(X, episode_duration = log(episode_duration) |> scale()) |>
+        data.matrix()
+
+    cases <- rowSums(is.na(X)) == 0
+
+    # Select the treatment option
+    state_sup <- paste0("ext_sup_s_state_", design[row, "treatment"])
[…]

### 2023-12-12 R/probit.R

@@ -3,77 +3,41 @@
 library(cmdstanr)
 library(dplyr)
 library(fc.utils)
-library(yaml)
+library(tools)
 
 options(mc.cores = parallel::detectCores() - 1)
 
 input <- commandArgs(trailingOnly = T)
-schema_file <- if (!exists("input") || length(input) == 0) "models/frozen-bin-all.yml" else input
+default <- "data/models/frozen-bin-all-with_censored.RData"
+model_data <- if (!exists("input") || length(input) == 0) default else input
 
-stopifnot(file.exists(schema_file))
+stopifnot(file.exists(model_data))
 
-schema <- read_yaml(schema_file)
-info("Running model: %s", schema$name)
-sprintf("posteriors/%s", schema$name) |> dir.create(showWarnings = F, recursive = T)
+load(model_data)
+model_name <- basename(model_data) |> file_path_sans_ext()
 
-df <- readRDS("./data/model_data.rds") |> ungroup()
+info("Running model: %s", model_name)
+sprintf("posteriors/%s", model_name) |> dir.create(showWarnings = F, recursive = T)
 
-###
-# Set the Unit of analysis
-if (schema$episodes == "high_intensity")
-    df <- filter(df, cumulative_intensity == 1)
-
-###
-# Select control variables - currently this is static across all models
-X <- select(df, censored, episode_duration, recur, high_intensity, incompatibility,
-            cold_war, ongoing_intrastate, ongoing_interstate) |>
-    mutate(episode_duration = log(episode_duration) |> scale()) |>
-    data.matrix()
-
-cases <- rowSums(is.na(X)) == 0
-
-###
-# Select the treatment option
-state_sup <- paste0("ext_sup_s_state_", schema$treatment)
-rebel_sup <- paste0("ext_sup_s_rebel_", schema$treatment)
-
-info("Treatment variables: %s and %s", state_sup, rebel_sup)
-
-treatments <- select(df, {{ state_sup }}, {{ rebel_sup }}) |>
-    mutate(interaction = .data[[{{ state_sup }}]] * .data[[{{ rebel_sup }}]]) |>
-    data.matrix()
-
-###
-# Assemble model input
 stan_data <- list(N = sum(cases),
                   K = ncol(treatments),
                   T = treatments[cases, ],
                   interaction_id = which(colnames(treatments) == "interaction"),
                   M = ncol(X),
                   X = X[cases, ],
[…]

### 2023-12-12 R/probit.R
BUGFIX: set the outcome variable depending on model
All of the models were previously being run using the strict
operationalization of frozen conflict. Fix so that we run both the
strict and normal versions of the frozen variable.

@@ -55,7 +55,7 @@ stan_data <- list(N = sum(cases),
                   country_id = as.factor(df$gwno_a[cases]) |> as.integer(),
                   n_contest_types = 2,
                   contest_id = df$incompatibility[cases] + 1,
-                  y = df$strict_frozen[cases])
+                  y = df[[schema$outcome]][cases])
 str(stan_data)
 
 mod <- cmdstan_model("./stan/hierarchical_probit.stan")

### 2023-12-12 paper.qmd
WIP: update skeleton notes in manuscript

@@
 ---
-title: "From Active Hostilities to Frozen Stalemates: Assessing the Role of External Support in Civil Conflict Dynamics"
+title: "Frozen Civil Conflicts"
+subtitle: "The Role of External State Support"
 author: J.L.A. Krusell
@@
     warning: false
+abstract: |
+    Frozen conflicts are among some of the most pernicious, difficult to
+    resolve conflicts. Surprisingly, little scholarly attention has been
+    paid to the factors influencing their emergence. This paper addresses
+    one possible dimension, namely the effect of external state support to
+    conflict actors on the probability of an ongoing civil conflict
+    becoming frozen. I use an original dataset coding frozen conflicts as
+    an outcome among UCDP civil conflict episodes from 1975 to 2019 in
+    order to construct a series of bayesian, hierarchical probit models
+    motivated by an application of bargaining theory to the domain of
+    frozen conflicts. Results indicate that directed support to rebels in
+    the absence of similar support to the government side increases the
+    probability that a civil conflict episode terminates as a frozen
+    conflict. All other treatment conditions --- specifically, directed
+    support provided only to the government and directed supported given
+    to both sides --- are found to be inconsequential. Taken together,
+    these findings help provide a more nuanced understanding of conflict
+    termination beyond the common assumption that conflicts either
+    conclude through a negotiated settlement or a one-sided military
+    victory. Furthermore, this paper also helps shed light on the
+    conditions under which actors may prefer a sub-optimal solution,
+    allowing a conflict to transition into a state of unstable peace with
+    an ever present threat of violent re-escalation, rather than a
+    conclusive termination.
 ---
@@
 
+# Introduction
+
 **Research question: to what extent does external state support to conflict actors affect the probability that a civil conflict becomes frozen?**
@@
 - *Original Data*, conceptual/descriptive claim about prevalence of frozen conflicts
-- Novel empirical analysis, provide a more nuanced analysis of *conflict termination* and contributes to our understanding of how external support can shape the dynamics of conflicts
+- Novel empirical analysis, provides a more nuanced analysis of *conflict termination* and contributes to our understanding of how external support can shape the dynamics of conflicts
 
@@
 In doing so, I show that frozen conflicts are more prevalent than thought and conceptually can and should cover a greater range of cases than just /de facto/ states.
-I also distinguish the concept of frozen conflicts from the related notions of stalemates and protracted conflicts (and maybe intractable rivalries?).
+I also distinguish the concept of frozen conflicts from the related notions of stalemates and intractable rivalries.
 
 I provide a novel quantitative analysis on the likelihood of a conflict becoming frozen.
-Specifically, I examine the effect of support by external state actors to belligerents in the course of a conflict episode.
+Specifically, I examine the effect of support by external state actors to belligerents in the course of a civil conflict episode.
 This will contribute both to our understanding of different types of conflict termination as well as the ways in which external support can shape the dynamics of conflicts.
@@
 **Why should we care about frozen conflicts?**
+
+# Previous Literature
+
+There exists extensive work on the impact of external third party interventions on civil conflict outcomes [@regan_third-party_2002; @balch-lindsay_third-party_2008; @gent_going_2008].
+This literature has emphasised the importance of targeted interventions, either military or economic, in favor of the opposition as increasing the probability of either a negotiated settlement or rebel military victory.
+The opposition is generally assumed to be the weaker side at the onset of fighting such that biased interventions in favor of the opposition can result in a much larger shift in the balance of power, allowing the opposition to credibly challenge the government.
+Subsequent research has further disaggregated the types of support that are offered to conflict belligerents, arguing in contrast that fungible support in particular actually decreases the likelihood of a negotiated settlement and increases conflict duration by contributing to information uncertainty over the true strength of the opposition [@sawyer_role_2017; @karlen_escalate_2023].
+Furthermore, external support has been found to also increase the likelihood of conflict recurrence by possibly creating an expectation of renewed assistance which may cause actors to be less risk averse and willing to militarily relitigate past conflicts [@karlen_legacy_2017].
+
+The common thread across these works is a general assumption that
+conflicts terminate either with a one-sided military victory or a negotiated settlement.
+Frozen conflicts, however, represent a qualitatively different type of conflict termination.
+Unlike a one-sided military victory, there is no clear victor when a conflict freezes, rather the conflict stalemates with both sides either unwilling or unable to continue large-scale hostilities.
+And, compared to a negotiated settlement, although frozen conflicts can be preceded by a ceasefire or partial settlement, they crucially do not encompass comprehensive peace agreements which would otherwise conclusively terminate a conflict.
+Thus, frozen conflicts represent a unique outcome; one that has previously not been examined, especially in regards to international dynamics such as external state support.
 
@@
 
-*The fundamental question*: why would actors accept a frozen conflict over a comprehensive peace agreement?
+**Fundamental question**: why would actors accept a frozen conflict over a comprehensive peace agreement?
 
-This mirrors the "why fight" question underlying bargaining theory where fighting is assumed to be costly such that with perfect information rational actors should prefer a "bargained solution." Is there a similar parallel to frozen conflicts?
+When a conflict freezes, actors elide the ongoing costs of fighting; however, there remains an ever-present threat of violent re-escalation.
+Thus, actors continue to bear the cost of militarization, lest there is an unfavourable shift in the balance of power enabling the opposing side to re-launch the conflict.[^recurrence] The question then is, why bear the cost of an ongoing frozen conflict, as well as the uncertainty of re-escalation which ensures that war spoils remain transient?
+A comprehensive peace agreement, especially with international guarantees, solidifies the current status quo and allows actors to translate war spoils to permanent gains.
 
-What is the cost of perpetuating a frozen conflict?
-How does this play in with the traditional, bargaining theory approach to conflict termination.
-Namely, actors become willing to terminate a conflict as information is revealed through fighting and the payoff from victory diminishes, based on the subjective probability of winning and the accumulated costs both past and future.
+[^recurrence]: Conflict recurrence can still be a threat even in the case of a conclusive termination, *i.e.*, one-sided military victory or comprehensive peace agreement.
+However, the uncertainty around future violence in a frozen conflict is higher since there is neither a *de jure* agreement settling the underlying issue or the payoff of a total victory where the victor can dictate the post-conflict status quo.
 
-Is there a situation in which rational actors would prefer to leave a conflict unresolved with the future threat of violent re-escalation?
-Clearly, the rebel side faces a credible commitment dilemma in that they can never guarantee that the government won't renege in the future.
+Three possible reasons for why actors may prefer to let a civil conflict freeze:
 
-Literature on ceasefires (ex Clayton and Stitcher (2021)) emphasize the strategic logic underlying actors' decisions to accept or reject a ceasefire.
-More specifically, actors may be willing to enter into a ceasefire either as a signaling mechanism to demonstrate an intent to peacefully resolve the conflict or as a delaying mechanism to allow them to temporarily pause fighting in order to rebuild their strength.
-The decision to accept a frozen conflict, despite the ongoing costs, may be similar to this latter mechanism.
-Conflict actors may not be willing to completely abandon their war aims; however, due to a variety of factors are unable to win an outright victory.
-By letting the conflict become frozen, either through a ceasefire, partial peace agreement, or a stalemate, actors can temporarily halt the ongoing costs of fighting in hopes of a future favourable shift in the balance of power.
-This probably requires actors to have long time horizons and be willing to burden the costs of not fully resolving the conflict.
-Specifically, letting a conflict become frozen means that there is always a threat of re-escalation so actors need to continue to commit to military spending, while suffer ongoing economic costs from lost territory.
+1.
+**Issue Indivisibility**: Issue indivisibility is often times downplayed in bargaining theory --- it is often assumed that there exists a possible division of contested goods even in the most intractable conflicts --- however, it could play an important role in frozen conflicts.
+Particularly from the perspective of the government when a negotiated settlement may involve the loss of territory to a rebel group.
+There exists strong norms around territorial integrity as well as domestic pressure which make it difficult for states to acquiesce to secession.
+Furthermore, in highly fractionalized states, the government has a strong incentive to refuse a negotiated settlement in order to deter future sovereignty claims by other groups.
 
-"...Leeds (1999) and Walter (2002) suggest that once negotiation begins because of a hurting stalemate, any warring group would prefer to have those negotiations be successful rather than to return to their group members without a settlement." --- from Park (2015)
+2.
+**Credible Committments**: For separatist conflicts it is often difficult for the government to credibly commit to an agreement granting greater autonomy to a separatist region.
+For the rebels, *de facto* independence as a half-way measure, achieved through a frozen conflict, can be the best possible outcome given that the international community tends to be fairly conservative regarding sovereignty claims.
+Even in the case of conflicts where the core contestation is over control of the central government, there exists a significant amount of literature pointing to the difficulties of guaranteeing the integrity of the incumbent, especially in authoritarian contexts were there is a fear of retribution.
 
-This is a potential theoretical contribution to the bargaining literature in that I could argue a comprehensive peace agreement does not necessarily always result from a hurting stalemate.
-Letting conflicts freeze could be one way that actors ease the costs of a hurting stalemate without being forced to commit to a settlement.
-This could be especially attractive in scenarios where there is such a high degree of mutual distrust, with few international guarantees for a negotiated settlement, where credible commitment problems could lead to a frozen conflict.
-Especially when actors receive external support, which a) is more likely to lead to the type of stalemate scenario where neither side can outright win militarily and b) allow actors to sustain the military threat required for a frozen conflict to perpetuate.
+3.
+**Time Horizons**: Probably the most significant mechanism concerns actors with forward-looking time horizons.
+Frozen conflicts, much like ceasefires which are often times involved in creating the conditions for a frozen conflict, can be a delaying tactic if an actor believes that future shifts in the distribution of power will be favourable to their cause.
+This can be especially important for separatist conflicts where maintaining sovereignty claims over a territory allows for an easy pretext for future escalations.
+International norms against wars of territorial conquest make relitigating a separatist war where the rebels have achieved a *de jure* independent state significantly more difficult, as compared to a conflict that remains a civil war within the eyes of the international community.
 
-Note, extensive literature arguing that external actors also increase the number of veto players which decrease the likelihood of a negotiated settlement (ex Cunningham 2006).
+How does external support affect this calculus?
 
-- Differentiating the theoretical mechanisms for the different treatment conditions: no external support, both sides receive external support, government only, rebel only
+1.
+Balance of power.
+This mirrors the theoretical arguments made in the previous literature on external support and conflict outcomes.
+Essentially, support to a belligerent increases their fighting capacity, and especially in the case of a rebel group that is severely outmatched by a government, increases the likelihood of a military stalemate.
+This does not mean that the conflict will automatically become frozen, but it is an important first step.
+Furthermore, the involvement of third party actors can complicate peace negotiations since interests may not also be aligned between the donor and recipient, or even between donor states.
 
-- States tend to be fairly conservative about sovereignty claims (see Hutchinson f.ex.). The best that separatists can hope for is a *de facto* state.
-
-- Big difference compared to ceasefires, frozen conflicts are not /a priori/ agreed upon by belligerents.
-Even if an actor prefers a conflict to become frozen, there is no guarantee that the opposing side won't attempt a re-escalation at any time.
-
-  - Even if actors prefer a frozen conflict over a comprehensive peace agreement; why not codify the status quo and formalized a ceasefire?
-(I think the key here is to emphasize that actors **cannot** win militarily, in other words they are stalemated into a frozen conflicts and refuse to give up their war aims.
-The question then is not, why do conflict actors prefer frozen conflicts, but rather in the face of a frozen conflict, why do conflict actors refuse a negotiated settlement?)
-
-- Frozen conflicts as a suboptimal solution to the contest between an incumbent and challenger.
+2.
+External support may also create an expectation that support will continue in the future such that an actor can offset the costs of an ongoing frozen conflict.
+For example, the post-Soviet separatist states are entirely reliant on Russian support, which allows them to maintain a state of militarization that deters re-escalation by the incumbent state.
 
@@
 ```{r}
+frozen <- filter(dataset, frozen == 1) |>
+    mutate(country = sub("Government of ", "", side_a),
+           country = case_when(country == "Russia (Soviet Union)" ~ "Russia",
+                               country == "United Kingdom" ~ "UK",
+                               country == "Congo" ~ "Republic of Congo",
+                               country == "Myanmar (Burma)" ~ "Myanmar",
+                               country == "Bosnia-Herzegovina" ~ "Bosnia and Herzegovina",
+                               country == "Serbia (Yugoslavia)" ~ "Serbia",
+                               country == "DR Congo (Zaire)" ~ "Democratic Republic of the Congo",
+                               T ~ country))
+
+sub.df <- group_by(frozen, country) |> count(frozen)
+
+world <- map_data("world") |>
+    left_join(sub.df, by = c("region" = "country")) |>
+    mutate(n = ifelse(is.na(n), 0, 1) )
+
+ggplot(world, aes(long, lat, map_id = region, fill= n, group = group)) +
+    geom_map(map = world) +
+    theme(legend.position = "none") +
+    scale_color_solarized()
+```
+
+```{r}
 filter(dataset, frozen == 1) |>
@@
     geom_bar() +
-    ylab("Frozen Civil Conflict Onsets") +
-    xlab("Year")
+    xlab("Year") +
+    ylab("Frozen Civil Conflict Onsets")
 ```
@@
 
+```{r}
+conditions <- dataset |>
+    mutate(treatment = case_when(ext_sup_s_rebel_bin == 1 & ext_sup_s_state_bin == 1 ~ "Ext. Sup. to Both",
+                                 ext_sup_s_rebel_bin == 1 & ext_sup_s_state_bin == 0 ~ "Ext. Sup. Rebel Only",
+                                 ext_sup_s_rebel_bin == 0 & ext_sup_s_state_bin == 1 ~ "Ext. Sup. Gov Only",
+                                 ext_sup_s_rebel_bin == 0 & ext_sup_s_state_bin == 0 ~ "Ext. Sup. to Neither"))
+
+m <- table(conditions$frozen, conditions$treatment)
+rownames(m) <- c("Not Frozen", "Frozen")
+knitr::kable(m)
+```
+
 ### Control Variables
@@
 delta$variable <- sprintf("delta[%s]", sapply(ll, `[`, 1))
-delta$group <- sapply(ll, `[`, 2) |> as.factor()
+delta$group <- ifelse(sapply(ll, `[`, 2) == 1, "Government Incompatibility",
+                      "Territorial Incompatibility")
 
 ggplot(delta, aes(x = value, y = variable, fill = variable)) +
-    stat_halfeye(alpha = 0.8) +
+    stat_halfeye(alpha = 0.8, show.legend = F) +
     geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.2) +
@@
                                 "delta[3]" = "Interaction")) +
-    facet_grid(. ~ group)
-
+    facet_grid(. ~ group) +
+    scale_fill_solarized()
 ```
@@
 #| fig.cap: Average marginal effects for the three treatment conditions on the predicted probability of a frozen conflict onset.[^ame]
+
 ame <- fit$draws("ame", format = "df") |>
@@
 ame$variable <- sprintf("ame[%s]", sapply(ll, `[`, 1))
-ame$group <- sapply(ll, `[`, 2) |> as.factor()
+ame$group <- ifelse(sapply(ll, `[`, 2) == 1, "Government Incompatibility",
+                    "Territorial Incompatibility")
 
@@
     coord_flip() +
-    facet_grid(. ~ group)
+    facet_grid(. ~ group) +
+    scale_fill_solarized()
 ```
@@
 
+```{r}
+filter(dataset, frozen == 1) |>
+    arrange(side_a, year) |>
+    select(Government = side_a, `Rebel Group(s)` = side_b, Year = year) |>
+    mutate(Government = sub("Government of", "", Government) |> trimws()) |>
+    kable(longtable = T)
+```
+
 ## Simulation Based Calibration
@@
 names(fits) <- basename(models)
+
+ll <- lapply(fits, function(fit) {
+    fit$summary("ame[2,2]", pzero, quantile2, median)
+}) |> bind_rows()
+
+ggplot(ll, aes(pzero, median)) +
+    geom_point() +
+    geom_errorbar(aes(ymin = q5, ymax = q95))
+
 ```

### 2023-12-13 R/sim.R
Align simulation script with dataset

@@ -11,9 +11,9 @@ library(dplyr)
 
 options(mc.cores = parallel::detectCores() - 1)
 
-n <- 300
+n <- 248
+m <- 8
 k <- 3
-m <- 2
 
 n_countries <- 2
 n_contest_types <- 2

### 2023-12-14 paper.qmd
Notes

@@
 I will then decrease the dependent and independent variables, including operationalizations and data sources.
+
+## Skeleton
+
+Bayesian probit model
+
+Let $y_i \in \{ 0, 1 \}$ be abinary indicator
+
+- Unit of analysis - civil conflict episodes
+
+  - Contiguous years of active fighting according to the UCDP definition
+
+  - Justification --- not interested in duration or predicting
+    __when__ a conflict terminates
+
+- X a matrix of control covariates
+
+- T a binary treatment matrix consisting of 3 variables:
+
+    1. External state support for rebels
+    2. External state support for the government
+    3. Interaction between previous two indicators to capture the
+       differential effect of both sides receiving external state
+       support.
+
+- Let $y^\star$ be a latent dependent variable such that ... regression equation ... and $P(y_i = 1) = \Phi(...)$ where $\Phi(\cdot)$ is the standard normal CDF.
+
+    Then $\delta_{k_i}$ is the vector of group varying treatment effects. $\gamma_{c_i}$ is a country specific intercept to capture heterogeneity
+
+I place weakly information priors with hierarchical priors on $\delta$ and $\gamma$ in order to partially pool the estimators.
+
+... prior equations ...
 

### 2024-01-27 R/merge.R
Add additional case to design matrix
Add a robustness check to also run our probit models without censored
episodes that are recurrent episodes.

@@ -17,6 +17,7 @@ term <- read_excel("./data/raw/ucdp-term-acd-3-2021.xlsx") |>
            intensity_level, incompatibility, recur) |>
     group_by(conflictep_id) |>
     mutate(censored = ifelse(min(year) < 1975, 1, 0),
+           recur = max(recur),
            gwno_a = as.numeric(gwno_a)) |>
     ungroup()
 

### 2024-01-27 R/models.R

@@ -11,7 +11,7 @@ df <- readRDS("./data/merged_data.rds") |> ungroup()
 treatments <- c("bin", "bin_5y", "prop")
 outcomes <- c("frozen", "strict_frozen")
 units <- c("all", "high_intensity")
-censored <- c("no_censored", "with_censored")
+censored <- c("no_censored", "no_recurring_censored", "with_censored")
 
 design <- expand.grid("outcome" = outcomes, "treatment" = treatments,
                       "episodes" = units, "censored" = censored) |>
@@ -28,8 +28,11 @@ for (row in 1:nrow(design)) {
 
     # Filter out censored observations if necessary, ie conflicts
     # beginning before 1975
-    if (design[row, "censored"] == "no_censored")
+    if (design[row, "censored"] == "no_censored") {
         sub.df <- filter(sub.df, censored == 0)
+    } else if (design[row, "censored"] == "no_recurring_censored") {
+        sub.df <- filter(sub.df, !(censored == 1 | recur == 1))
+    }
 
     # Select control variables - this is mostly static across models
     X <- select(sub.df, censored, episode_duration, recur, high_intensity, incompatibility,
@@ -37,7 +40,7 @@ for (row in 1:nrow(design)) {
 
     # If we've dropped censored episodes, no need for binary control
     # var
-    if (design[row, "censored"] == "no_censored")
+    if (design[row, "censored"] != "with_censored")
         X <- select(X, -censored)
 
     X <- mutate(X, episode_duration = log(episode_duration) |> scale()) |>

### 2024-01-27 R/models.R
Use integer identifiers for each model run
The string names for each model are becoming too long as different
design features are added. Instead, just use a short integer
identifier and save the design matrix as a mapping table.

@@ -15,10 +15,13 @@ censored <- c("no_censored", "no_recurring_censored", "with_censored")
 
 design <- expand.grid("outcome" = outcomes, "treatment" = treatments,
                       "episodes" = units, "censored" = censored) |>
-    mutate(name = sprintf("%s-%s-%s-%s", outcome, treatment, episodes, censored))
+    mutate(id = sprintf("%04d", row_number()))
+
+# Save the design matrix
+saveRDS(design, "./data/models/design.rds")
 
 for (row in 1:nrow(design)) {
-    info("Model %s", design[row, "name"])
+    info("Model %s", design[row, "id"])
 
     # Start off by setting the unit of analysis
     sub.df <- if (design[row, "episodes"] == "high_intensity")
@@ -59,6 +62,6 @@ for (row in 1:nrow(design)) {
     # Set the outcome variable
     y <- sub.df[[design[row, "outcome"]]]
 
-    f <- sprintf("./data/models/%s.RData", design[row, "name"])
+    f <- sprintf("./data/models/data_%s.RData", design[row, "id"])
     save(sub.df, cases, treatments, X, y, file = f)
 }

### 2024-01-27 R/probit.R

@@ -8,7 +8,7 @@ library(tools)
 options(mc.cores = parallel::detectCores() - 1)
 
 input <- commandArgs(trailingOnly = T)
-default <- "data/models/frozen-bin-all-with_censored.RData"
+default <- "data/models/data_001.RData"
 model_data <- if (!exists("input") || length(input) == 0) default else input
 
 stopifnot(file.exists(model_data))

### 2024-01-28 R/probit.R
Update Makefile to use model RDatas

@@ -8,7 +8,7 @@ library(tools)
 options(mc.cores = parallel::detectCores() - 1)
 
 input <- commandArgs(trailingOnly = T)
-default <- "data/models/data_001.RData"
+default <- "data/models/data_0001.RData"
 model_data <- if (!exists("input") || length(input) == 0) default else input
 
 stopifnot(file.exists(model_data))

### 2024-01-29 R/probit.R
Don't parallelize stan runs
Avoid clashing with Makefile jobs.

@@ -5,7 +5,7 @@ library(dplyr)
 library(fc.utils)
 library(tools)
 
-options(mc.cores = parallel::detectCores() - 1)
+options(mc.cores = 1)
 
 input <- commandArgs(trailingOnly = T)
 default <- "data/models/data_0001.RData"

### 2024-01-29 R/sbc.R
BUGFIX: harmonize variable names b/w stan files

@@ -33,11 +33,10 @@ conflict_id <- sample(1:n_conflict_types, n, replace = T)
 treatments <- sample(0:1, n*2, replace = T) |> matrix(nrow = n, ncol = 2)
 treatments <- cbind(treatments, treatments[, 1] * treatments[, 2])
 
-
-data <- list(n = n,
-             k = k,
+data <- list(N = n,
+             K = k,
              T = treatments,
-             m = m,
+             M = m,
              X = X,
              n_countries = n_countries,
              country_id = country_id,

### 2024-01-29 paper.qmd
Replace json-based model runs in manuscript

@@
 theme_set(theme_tufte(base_family = "TeX Gyre Pagella", base_size = 10))
+
+design <- readRDS("./data/models/design.rds")
 ```
@@
 # Theory
+
+Why do external states intervene in civil conflicts?
+To further their national interests.
+Interveners face constraints, strategic decisions over limited resources, as well as from other interveners (ex: Anderson 2019)
 
@@
 #| fig.cap: Posterior estimates for \delta from a single variate probit model
-fit <- readRDS("./posteriors/frozen-bin-all/fit.rds")
+main_model <- filter(design, outcome == "frozen", treatment == "bin", episodes == "all",
+                     censored == "with_censored") |> pull(id)
+
+fit <- sprintf("posteriors/data_%s/fit.rds", main_model) |>  readRDS()
 
@@
 #| fig.cap: Average marginal effects for the three treatment conditions on the predicted probability of a frozen conflict onset.[^ame]
-
 ame <- fit$draws("ame", format = "df") |>

### 2024-01-29 paper.qmd
Aggregate all the robustness checks

@@
 
-```{R eval = F}
+```{R}
+pzero <- function(x) mean(x > 0)
+
 models <- list.dirs("posteriors", recursive = F)
@@
 
-ll <- lapply(fits, function(fit) {
-    fit$summary("ame[2,2]", pzero, quantile2, median)
+model_ames <- lapply(fits, function(fit) {
+    fit$summary("ame", pzero, posterior::quantile2, median)
 }) |> bind_rows()
 
-ggplot(ll, aes(pzero, median)) +
-    geom_point() +
-    geom_errorbar(aes(ymin = q5, ymax = q95))
+ll <- extract_digits(model_ames$variable)
+ame_keys <- sprintf("ame[%s]", sapply(ll, `[`, 1))
+
+model_ames$variable <- case_when(ame_keys == "ame[1]" ~ "Ext. Sup. Gov Only",
+                                 ame_keys == "ame[2]" ~ "Ext. Sup. Rebel Only",
+                                 ame_keys == "ame[3]" ~ "Ext. Sup. to Both")
+model_ames$group <- ifelse(sapply(ll, `[`, 2) == 1, "Government Incompatibility",
+                           "Territorial Incompatibility")
+
+ggplot(model_ames, aes(pzero, median)) +
+    geom_point(alpha = 0.2) +
+    geom_errorbar(aes(ymin = q5, ymax = q95), alpha = 0.25) +
+    geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.2) +
+    facet_grid(variable ~ group) +
+    labs(y = parse(text = "tilde(Delta)"), x = parse(text = "P(tilde(Delta) > 0)")) +
+    theme(legend.position = "none")
 
 ```
-
-```{R eval = F}
-#| fig-height: 7
-summary.df <- lapply(fits, function(fit) {
-    post <- post_summarise(fit, "delta", rename = T, model_info = T,
-                           probs = c(0.025, 0.17, 0.5, 0.83, 0.975)) |>
-        rename(median = `50%`,
-               lower95 = `2.5%`,
-               upper95 = `97.5%`,
-               lower66 = `17%`,
-               upper66 = `83%`)
-}) |> bind_rows()
-
-summary.df$variable <- sub("_(prop|bin.*)$", "", summary.df$variable)
-
-ggplot(summary.df, aes(treatment_aggregation, median, color = outcome)) +
-    geom_point(size = 1.5, position = position_dodge(width = 0.25)) +
-    geom_errorbar(aes(ymin = lower95, ymax = upper95), position = position_dodge(0.25), width = 0) +
-    geom_errorbar(aes(ymin = lower66, ymax = upper66), position = position_dodge(0.25), linewidth = 1.25, width = 0) +
-    geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.2) +
-    facet_grid(variable ~ episodes)
-
-```
-
-```{R eval = F}
-regtable <- function(fits, ...) {
-    names(fits) <- NULL
-    modelsummary(fits, pars = c("delta", "beta", "alpha"), statistic = "conf.int", ...)
-}
-
-m <- c("frozen-bin-all", "strict_frozen-bin-all",
-       "frozen-bin-high_intensity", "strict_frozen-bin-high_intensity")
-regtable(fits[m], title = "Binary Treatments")
-```
-
-```{R eval = F}
-m <- c("frozen-bin_5y-all", "strict_frozen-bin_5y-all",
-       "frozen-bin_5y-high_intensity", "strict_frozen-bin_5y-high_intensity")
-regtable(fits[m], title = "Binary Treatments - Last 5 Years")
-```
-
-```{R  eval = F}
-m <- c("frozen-prop-all", "strict_frozen-prop-all",
-       "frozen-prop-high_intensity", "strict_frozen-prop-high_intensity")
-regtable(fits[m], title = "Proportion Treatments")
-```

### 2024-01-29 R/scratch.R
Remove scratch file

@@ -1,13 +1,0 @@
-#!/usr/bin/env Rscript
-
-library(dplyr)
-library(readxl)
-
-ucdp <- read_xlsx("./data/Dyadic_v23_1.xlsx")
-term <- read_xlsx("./data/ucdp-term-acd-3-2021.xlsx")
-
-filter(ucdp, gwno_a == 850) |>
-    select(conflict_id, dyad_id, year, side_a, side_b, gwno_a, gwno_b) |>
-    View()
-
-x <- filter(term, type_of_conflict %in% 3:4)

### 2024-01-29 paper.qmd
Add sbc histograms

@@
 ```{r}
+
 filter(dataset, frozen == 1) |>
@@
 ## Simulation Based Calibration
+
+```{R}
+sbc <- readRDS("./posteriors/sbc.rds") |>
+    select(-.chain, -.iteration, -.draw) |>
+    pivot_longer(everything())
+
+ggplot(sbc, aes(x = value)) +
+    geom_histogram(bins = 10) +
+    facet_wrap(~ name, scales = "free") +
+    theme(axis.title.y = element_blank(),
+          axis.title.x = element_blank())
+```
 

### 2024-02-01 R/probit.R
Identify model runs from probit script

@@ -7,6 +7,8 @@ library(tools)
 
 options(mc.cores = 1)
 
+design <- readRDS("./data/models/design.rds")
+
 input <- commandArgs(trailingOnly = T)
 default <- "data/models/data_0001.RData"
 model_data <- if (!exists("input") || length(input) == 0) default else input
@@ -16,7 +18,10 @@ stopifnot(file.exists(model_data))
 load(model_data)
 model_name <- basename(model_data) |> file_path_sans_ext()
 
-info("Running model: %s", model_name)
+with(filter(design, id == model_name),
+     info("Model %s => outcome: %s, treatment: %s, episodes: %s, censored: %s",
+          id, outcome, treatment, episodes, censored))
+
 sprintf("posteriors/%s", model_name) |> dir.create(showWarnings = F, recursive = T)
 
 stan_data <- list(N = sum(cases),
@@ -33,7 +38,7 @@ stan_data <- list(N = sum(cases),
 str(stan_data)
 
 mod <- cmdstan_model("./stan/hierarchical_probit.stan")
-fit <- mod$sample(data = stan_data, sig_figs = 3, adapt_delta = 0.99)
+fit <- mod$sample(data = stan_data, refresh = 0, sig_figs = 3, adapt_delta = 0.99)
 
 # Treatment coefficients
 fit$summary("delta")

### 2024-02-05 R/merge.R
Add conflict censor variable

@@ -16,11 +16,20 @@ term <- read_excel("./data/raw/ucdp-term-acd-3-2021.xlsx") |>
     select(conflict_id, conflictep_id, year, gwno_a = gwno_loc, side_a, side_b,
            intensity_level, incompatibility, recur) |>
     group_by(conflictep_id) |>
-    mutate(censored = ifelse(min(year) < 1975, 1, 0),
+    mutate(episode_censored = ifelse(min(year) < 1975, 1, 0),
            recur = max(recur),
            gwno_a = as.numeric(gwno_a)) |>
     ungroup()
 
+# Censored indicates if the conflict began before 1975
+censored <- group_by(term, conflict_id, conflictep_id) |>
+    summarise(year = last(year)) |>
+    mutate(censored = any(year < 1975) |> as.integer()) |>
+    select(-year)
+
+term <- left_join(term, censored, by = c("conflict_id", "conflictep_id"))
+
+# UCDP - Conflict Data
 ucdp <- readRDS("./data/raw/UcdpPrioConflict_v23_1.rds")
 
 # Total number of ongoing high intensity interstate and intrastate
@@ -118,15 +127,20 @@ final.df <- filter(full.df, !is.na(ext_sup_s_state)) |>
               year = last(year),
               side_a = first(side_a),
               side_b = first(side_b),
+
               censored = max(censored),
+              episode_censored = max(episode_censored),
               episode_duration = n(),
               frozen_duration = last(duration),
               recur = max(recur),
+
               incompatibility = max(incompatibility == 1),
               cold_war = ifelse(max(year) < 1989, 1, 0),
+
               ongoing_interstate = max(ongoing_interstate) > 0,
               ongoing_intrastate = max(ongoing_intrastate) > 0,
               ongoing_conflict = ongoing_interstate | ongoing_intrastate,
+
               cumulative_intensity = max(cumulative_intensity),
               high_intensity = max(intensity_level) == 2,
 

### 2024-02-05 R/models.R

@@ -10,8 +10,8 @@ df <- readRDS("./data/merged_data.rds") |> ungroup()
 # Create a design matrix with all possible analysis combinations
 treatments <- c("bin", "bin_5y", "prop")
 outcomes <- c("frozen", "strict_frozen")
-units <- c("all", "high_intensity")
-censored <- c("no_censored", "no_recurring_censored", "with_censored")
+units <- c("all", "cumulative_intensity")
+censored <- c("no_censored", "no_episode_censored", "with_censored")
 
 design <- expand.grid("outcome" = outcomes, "treatment" = treatments,
                       "episodes" = units, "censored" = censored) |>
@@ -24,7 +24,7 @@ for (row in 1:nrow(design)) {
     info("Model %s", design[row, "id"])
 
     # Start off by setting the unit of analysis
-    sub.df <- if (design[row, "episodes"] == "high_intensity")
+    sub.df <- if (design[row, "episodes"] == "cumulative_intensity")
         filter(df, cumulative_intensity == 1)
     else
         df
@@ -33,18 +33,18 @@ for (row in 1:nrow(design)) {
     # beginning before 1975
     if (design[row, "censored"] == "no_censored") {
         sub.df <- filter(sub.df, censored == 0)
-    } else if (design[row, "censored"] == "no_recurring_censored") {
-        sub.df <- filter(sub.df, !(censored == 1 | recur == 1))
+    } else if (design[row, "censored"] == "no_episode_censored") {
+        sub.df <- filter(sub.df, episode_censored == 0)
     }
 
     # Select control variables - this is mostly static across models
-    X <- select(sub.df, censored, episode_duration, recur, high_intensity, incompatibility,
-                cold_war, ongoing_intrastate, ongoing_interstate)
+    X <- select(sub.df, episode_censored, episode_duration, recur, cumulative_intensity,
+                incompatibility, cold_war, ongoing_intrastate, ongoing_interstate)
 
     # If we've dropped censored episodes, no need for binary control
     # var
     if (design[row, "censored"] != "with_censored")
-        X <- select(X, -censored)
+        X <- select(X, -episode_censored)
 
     X <- mutate(X, episode_duration = log(episode_duration) |> scale()) |>
         data.matrix()

### 2024-02-05 paper.qmd
Add id to df for aggregated plot

@@
 ```{r}
-dataset <- readRDS("./data/model_data.rds")
+dataset <- readRDS("./data/merged_data.rds")
 n_frozen <- filter(dataset, frozen == 1) |> nrow()
@@
 
-model_ames <- lapply(fits, function(fit) {
-    fit$summary("ame", pzero, posterior::quantile2, median)
-}) |> bind_rows()
+model_ames <- lapply(names(fits), function(model) {
+    fits[[model]]$summary("ame", pzero, posterior::quantile2, median) |>
+                    mutate(id = sub("data_", "", model))
+}) |>
+    bind_rows() |>
+    left_join(design, by = "id")
 
+# Assign the correct variable names/groups
 ll <- extract_digits(model_ames$variable)

### 2024-02-05 paper.qmd
Add initial text describing probit model

@@
 execute:
+    cache: true
     echo: false
@@
 
-## Skeleton
-
-Bayesian probit model
-
-Let $y_i \in \{ 0, 1 \}$ be abinary indicator
-
-- Unit of analysis - civil conflict episodes
-
-  - Contiguous years of active fighting according to the UCDP definition
-
-  - Justification --- not interested in duration or predicting
-    __when__ a conflict terminates
-
-- X a matrix of control covariates
-
-- T a binary treatment matrix consisting of 3 variables:
-
-    1. External state support for rebels
-    2. External state support for the government
-    3. Interaction between previous two indicators to capture the
-       differential effect of both sides receiving external state
-       support.
-
-- Let $y^\star$ be a latent dependent variable such that ... regression equation ... and $P(y_i = 1) = \Phi(...)$ where $\Phi(\cdot)$ is the standard normal CDF.
-
-Then $\delta_{k_i}$ is the vector of group varying treatment effects. $\gamma_{c_i}$ is a country specific intercept to capture heterogeneity
-
-I place weakly information priors with hierarchical priors on $\delta$ and $\gamma$ in order to partially pool the estimators.
-
-... prior equations ...
-
 ## Probit Model
 
-I model the onset of a frozen civil conflict using a bayesian probit regression where the unit analysis is civil conflict episodes, *i.e.*, contiguous years of active fighting within a conflict.
-The outcome variable, $y \in \{ 1, 2 \}$, is a binary indicator denoting the
-
-TODO: eventually I would like to move towards a dyadic analysis, rather than conflict level
-
-- Unit of analysis: conflict episodes, ie consecutive years of active fighting (above UCDP threshold of 25 battle deaths)
-- Binary outcome, binary treatment, while conditioning on a number of relevant covariates
+I model the onset of frozen civil conflicts using a hierarchical bayesian probit model.
+The unit of analysis is conflict episodes --- contiguous years of active fighting according  to the UCDP definition of armed conflict as a "contested incompatibility that concerns a government and/or territory where the use of armed forced between two parties, of which at least one is the government of a state, results in at least 25 battle-related deaths within a calendar year" [@UCDP].
+The outcome, $y_i \in \{0, 1\}$ for $i = 1, 2, \ldots, N$ episodes, is a binary indicator denoting whether a given episode terminates in a frozen state modeled as a function of a matrix of treatment variables, $T$, and a set of additional control covariates, $X$, such that
 
 \begin{align*}
-y_i^\star & = \alpha + X_i' \beta + T_i' \delta + \epsilon_i, \; \epsilon_i \sim \text{N}(0, \sigma^2) \\
-y_i & = \begin{cases}
-1 & \text{if } y_i^\star > 0 \\
-0 & \text{otherwise}
-\end{cases}
+y_i | \pi_i & \sim \text{Bernoulli}(\pi_i) \\
+\pi_i & = \Phi\left(\alpha + X_i' \beta + T_i' \delta_{k_i} + \gamma_{c_i}\right)
 \end{align*}
 
-- Estimate with Stan, weakly informative priors, $\delta \sim \text{Normal}(1, 1)$
+where $\Phi(\cdot)$ is the standard normal cumulative density function.
+Then, $\delta_k$ is the group specific vector of treatment effects.
+I allow the treatment effects to vary according to whether the underlying incompatibility for the given conflict episode concerns territory or control of the central government in order to account for potential differential effects, thus $\delta \in \mathbb{R}^{D \times K}$ for $D$ treatment variables.
+Finally, $\gamma_c$ represents a varying intercept for each country in order to capture unmodeled time invariant heterogeneity between countries.
+
+In the case of the global intercept and slope coefficients for the control variables, $X$, I select fairly conservative, weakly informative priors given the restricted sample size and moderate number of predictors.
+Specifically, I set $\alpha \sim \text{N}(0, 5)$ and $\beta \sim \text{N}(0, 1)$ to improve sampling efficiency by placing less prior probability density on unlikely large values on the probit scale and to provide a degree of regularization.
+
+For $\delta_k$ and $\gamma$ I use hierarchical priors that allow for varying effects while partially pooling information across groups to prevent overfitting.
+Regarding the latter, this means $\delta_{jk} \sim \text{N}(\mu_j, \tau_j)$ for $j = 1, 2, \ldots, D$ where $\mu_j \sim \text{N}(0, 1)$ and $\tau_j \sim \text{N}(0, 2.5)$.
+And for the former, $\gamma_c \sim \text{N}(0, \sigma)$ and $\sigma \sim \text{N}(0, 1)$.
+The particularly restrictive hyperprior on $\sigma$ shrinks the posterior $\gamma$ estimates towards zero and encodes a significant degree of skepticism towards country-specific deviations.[^country-intercepts]
+
+[^country-intercepts]: The risk for overfitting is especially high for the country-specific intercepts since the number of observations per country is relatively small with the median country only experiencing two episodes within the time period under study.
+
+Due to the non-linear nature of a probit regression model, in order to interpret the estimated association between the treatments and outcome, the target empirical estimands are the group-specific average marginal effects, *i.e.* the changes in the conditional expectation of the outcome variable given the various treatment states versus the baseline control averaged over the relevant subgroup sample.
+Let $\theta$ denote the vector of all unknown parameters in the regression model, then the average marginal effect (AME) for the $j^{th}$ treatment state represented by vector $w_j$ for subsample $k$ can be approximated using $m = 1, 2, \ldots, M$ posterior draws from $\theta$ such that
+
+\begin{align*}
+\text{AME}_{jk}^{(m)} = \frac{1}{N_k} \sum_{i \in S_k} P\left(y_i = 1 | T = w_j, X_i, \theta^{(m)}\right) - P\left(y_i = 1 | T = 0_{D}, X_i, \theta^{(m)}\right)
+\end{align*}
 

### 2024-02-06 paper.qmd
Move draft notes to fc.org

@@
 library(dplyr)
-library(flextable)
 library(ggplot2)
@@
 library(ggdist)
-library(modelsummary)
 library(knitr)
@@
 **Research question: to what extent does external state support to conflict actors affect the probability that a civil conflict becomes frozen?**
-
-**Motivation**
-
-Contribution: (*conceptual* and *empirical*)
-
-- *Original Data*, conceptual/descriptive claim about prevalence of frozen conflicts
-- Novel empirical analysis, provides a more nuanced analysis of *conflict termination* and contributes to our understanding of how external support can shape the dynamics of conflicts
-
-The paper will argue that there exists both a theoretical and empirical gap.
-Studies on conflict termination have generally considered military victory and peace agreements as the only outcome.
-I further disaggregate the notion of termination to cover also the situation where conflicts are not conclusively termination and instead become "frozen."
-
-In doing so, I show that frozen conflicts are more prevalent than thought and conceptually can and should cover a greater range of cases than just /de facto/ states.
-I also distinguish the concept of frozen conflicts from the related notions of stalemates and intractable rivalries.
-
-I provide a novel quantitative analysis on the likelihood of a conflict becoming frozen.
-Specifically, I examine the effect of support by external state actors to belligerents in the course of a civil conflict episode.
-This will contribute both to our understanding of different types of conflict termination as well as the ways in which external support can shape the dynamics of conflicts.
-
-- Potential shortcoming: what I am not doing is taking into account the preferences of external supporters and motivation for offering support.
-Nor am I distinguishing between different external support types.
-
-**Why should we care about frozen conflicts?**
-
-# Previous Literature
-
-There exists extensive work on the impact of external third party interventions on civil conflict outcomes [@regan_third-party_2002; @balch-lindsay_third-party_2008; @gent_going_2008].
-This literature has emphasised the importance of targeted interventions, either military or economic, in favor of the opposition as increasing the probability of either a negotiated settlement or rebel military victory.
-The opposition is generally assumed to be the weaker side at the onset of fighting such that biased interventions in favor of the opposition can result in a much larger shift in the balance of power, allowing the opposition to credibly challenge the government.
-Subsequent research has further disaggregated the types of support that are offered to conflict belligerents, arguing in contrast that fungible support in particular actually decreases the likelihood of a negotiated settlement and increases conflict duration by contributing to information uncertainty over the true strength of the opposition [@sawyer_role_2017; @karlen_escalate_2023].
-Furthermore, external support has been found to also increase the likelihood of conflict recurrence by possibly creating an expectation of renewed assistance which may cause actors to be less risk averse and willing to militarily relitigate past conflicts [@karlen_legacy_2017].
-
-The common thread across these works is a general assumption that
-conflicts terminate either with a one-sided military victory or a negotiated settlement.
-Frozen conflicts, however, represent a qualitatively different type of conflict termination.
-Unlike a one-sided military victory, there is no clear victor when a conflict freezes, rather the conflict stalemates with both sides either unwilling or unable to continue large-scale hostilities.
-And, compared to a negotiated settlement, although frozen conflicts can be preceded by a ceasefire or partial settlement, they crucially do not encompass comprehensive peace agreements which would otherwise conclusively terminate a conflict.
-Thus, frozen conflicts represent a unique outcome; one that has previously not been examined, especially in regards to international dynamics such as external state support.
-
-# Defining Frozen Conflicts
-
-I follow the definition of frozen conflicts offered by @smetana_between_2019 as a "protracted, post-war conflict process, characterized by the absence of stable peace...[where] core issues between the opposing sides remain unresolved, the dispute is in the forefront of mutual relations, and there is a looming threat of the renewal of violence."  Accompanying this definition, @smetana_between_2019 lay forth three criteria determining whether a post-conflict period qualifies as a frozen conflict.
-First, there needs to be an initial period of active conflict followed by a cessation of large-scale violent hostilities.
-Isolated, sporadic incidences of violence may continue to occur; however, conflict actors are either unwilling or unable to launch major military offensives.
-Although not necessary, this may be characterized by the establishment of a general ceasefire and/or presence of external peacekeeping troops.
-Second, the core issue underlying the conflict needs to remain salient post-conflict.
-This entails actors maintaining their claims within a dispute.
-And, finally, there must be an ongoing threat of violent re-escalation.
-In other words, post-conflict relations should be characterized by mutual hostility and an unwillingness to abandon violence as a political strategy.
-
-These criteria distinguish frozen conflicts from related concepts such as protracted stalemates or enduring rivalries.
-In the case of the former, stalemates during a conflict can potentially endure for extended periods of time and offer conditions similar to those of a frozen conflict.
-However, they do not necessitate a cessation of active fighting.
-Stalemates can still occur during high intensity periods of a conflict where relatively few gains are made by either side.
-In the case of the latter, enduring rivalries are characterized in part by repeated onsets of militarized interstate disputes [@Goertz1992; @Goertz1993]; however, this might not entail the actual outbreak of violence.
-More importantly, enduring rivalries were originally conceptualized in terms of inter-state relations, while the definition of frozen conflicts here can be extended to the intra-state case as well.
-
-# Theory
-
-Why do external states intervene in civil conflicts?
-To further their national interests.
-Interveners face constraints, strategic decisions over limited resources, as well as from other interveners (ex: Anderson 2019)
-
-**Fundamental question**: why would actors accept a frozen conflict over a comprehensive peace agreement?
-
-When a conflict freezes, actors elide the ongoing costs of fighting; however, there remains an ever-present threat of violent re-escalation.
-Thus, actors continue to bear the cost of militarization, lest there is an unfavourable shift in the balance of power enabling the opposing side to re-launch the conflict.[^recurrence] The question then is, why bear the cost of an ongoing frozen conflict, as well as the uncertainty of re-escalation which ensures that war spoils remain transient?
-A comprehensive peace agreement, especially with international guarantees, solidifies the current status quo and allows actors to translate war spoils to permanent gains.
-
-[^recurrence]: Conflict recurrence can still be a threat even in the case of a conclusive termination, *i.e.*, one-sided military victory or comprehensive peace agreement.
-However, the uncertainty around future violence in a frozen conflict is higher since there is neither a *de jure* agreement settling the underlying issue or the payoff of a total victory where the victor can dictate the post-conflict status quo.
-
-Three possible reasons for why actors may prefer to let a civil conflict freeze:
-
-1.
-**Issue Indivisibility**: Issue indivisibility is often times downplayed in bargaining theory --- it is often assumed that there exists a possible division of contested goods even in the most intractable conflicts --- however, it could play an important role in frozen conflicts.
-Particularly from the perspective of the government when a negotiated settlement may involve the loss of territory to a rebel group.
-There exists strong norms around territorial integrity as well as domestic pressure which make it difficult for states to acquiesce to secession.
-Furthermore, in highly fractionalized states, the government has a strong incentive to refuse a negotiated settlement in order to deter future sovereignty claims by other groups.
-
-2.
-**Credible Committments**: For separatist conflicts it is often difficult for the government to credibly commit to an agreement granting greater autonomy to a separatist region.
-For the rebels, *de facto* independence as a half-way measure, achieved through a frozen conflict, can be the best possible outcome given that the international community tends to be fairly conservative regarding sovereignty claims.
-Even in the case of conflicts where the core contestation is over control of the central government, there exists a significant amount of literature pointing to the difficulties of guaranteeing the integrity of the incumbent, especially in authoritarian contexts were there is a fear of retribution.
-
-3.
-**Time Horizons**: Probably the most significant mechanism concerns actors with forward-looking time horizons.
-Frozen conflicts, much like ceasefires which are often times involved in creating the conditions for a frozen conflict, can be a delaying tactic if an actor believes that future shifts in the distribution of power will be favourable to their cause.
-This can be especially important for separatist conflicts where maintaining sovereignty claims over a territory allows for an easy pretext for future escalations.
-International norms against wars of territorial conquest make relitigating a separatist war where the rebels have achieved a *de jure* independent state significantly more difficult, as compared to a conflict that remains a civil war within the eyes of the international community.
-
-How does external support affect this calculus?
-
-1.
-Balance of power.
-This mirrors the theoretical arguments made in the previous literature on external support and conflict outcomes.
-Essentially, support to a belligerent increases their fighting capacity, and especially in the case of a rebel group that is severely outmatched by a government, increases the likelihood of a military stalemate.
-This does not mean that the conflict will automatically become frozen, but it is an important first step.
-Furthermore, the involvement of third party actors can complicate peace negotiations since interests may not also be aligned between the donor and recipient, or even between donor states.
-
-2.
-External support may also create an expectation that support will continue in the future such that an actor can offset the costs of an ongoing frozen conflict.
-For example, the post-Soviet separatist states are entirely reliant on Russian support, which allows them to maintain a state of militarization that deters re-escalation by the incumbent state.
 
@@
 
-- Episode Intensity
-- Episode Duration
-- Incompatibility? (ie separatist conflict over territory)
-- Recurrent episode
-- Right censored
+- Episode censored
+- Episode Duration log
+- Recur
+- Cumulative Intensity
+- Incompatibility
+- Cold War
+- Ongoing intra/interstate conflict
 
@@
 
-### Descriptive Statistics
-
-## Observational Models
-
 ### Results
 
-Figure 1 shows the parameter estimates for $\delta$.
+The actual model is implemented in Stan, a probabilistic programming language
+
+Using its builtin No-U-Turn sampler run for 2,000 iterations
 
@@
 ```
-
-[^ame]: Hello World
-
-- Parameter Plot
-- Density of predicted probabilities for true success/failure
-- Marginal effects
-- Confusion Matrix + AUR/PRC plots
-
-#### Robustness Checks
-
-- External support only within the last five years of conflict
-- Frozen conflicts with a duration of only 1 year
-- Exclude low intensity conflicts?
-
-Appendix:
-
-- Proportion of episode years with external support
-- Nr of external supporters, maybe weighted by cinc?
-- Territorial conflicts only
-- FCD data
-
-## IV Model
-
-\newpage
 

### 2024-02-06 paper.qmd
Add text describing dependent variable

@@
 I model the onset of frozen civil conflicts using a hierarchical bayesian probit model.
-The unit of analysis is conflict episodes --- contiguous years of active fighting according  to the UCDP definition of armed conflict as a "contested incompatibility that concerns a government and/or territory where the use of armed forced between two parties, of which at least one is the government of a state, results in at least 25 battle-related deaths within a calendar year" [@UCDP].
+The unit of analysis is conflict episodes --- contiguous years of active fighting according  to the UCDP definition of armed conflict as a "contested incompatibility that concerns a government and/or territory where the use of armed forced between two parties, of which at least one is the government of a state, results in at least 25 battle-related deaths within a calendar year" [@gleditsch_armed_2002; @davies_organized_2023].
 The outcome, $y_i \in \{0, 1\}$ for $i = 1, 2, \ldots, N$ episodes, is a binary indicator denoting whether a given episode terminates in a frozen state modeled as a function of a matrix of treatment variables, $T$, and a set of additional control covariates, $X$, such that
@@
 
-In order to measure the onset of frozen civil conflicts I followed the previous three criteria and coded an original dataset using dyadic intra-state conflict episodes drawn from the UCDP/PRIO Armed Conflict Dataset [@gleditsch_armed_2002; @davies_organized_2023] defined as contiguous years with at least 25 battle-related deaths in a violent contest over a government or territory between a state and non-state actor.
-To address post-conflict issue saliency, I considered an episode to be unresolved if it terminated without either a comprehensive peace treaty addressing the core underlying issues or a one-sided military victory for either the government or rebel group according to the UCDP Peace Agreement and Conflict Termination Datasets [@pettersson_organized_2019; @kreutz_how_2010].
-I also filtered out episodes where one or more actors ceased to exist within five years of the conflict termination and where the non-state actor renounces violences.
-This was done using a manual coding process with a mix of academic and news media sources for different conflicts in order to capture instances where, for example, rebel groups transition into legitimate political parties.
+The outcome variable in my regression model is a binary indicator for whether a civil conflict terminates in a frozen state.
+Following the definition of frozen conflicts offered earlier, I construct this measure using a mix of automated and manual coding of intrastate and internationalized intrastate conflict terminations taken from the UCDP Conflict Termination Dataset [@kreutz_how_2010].[^protocol] Episodes were considered terminated if the total number of battle related fatalities dropped below 25 within a year.
+I subsequently labeled a termination as a frozen conflict if three criteria were satisfied.
+First, an episode could not result in a one-sided military victory.
+This would constitute a conclusive termination allowing the victor to set the terms of the post-conflict order.
+Second, the episode could not conclude with a comprehensive peace agreement that addresses the underlying core issue according to the UCDP Peace Agreement Dataset [@pettersson_organized_2019].[^partial] To account for the possibility of a failed peace agreement I further required that the comprehensive peace agreement had to last for at least one calendar year in order to disqualify an episode termination as a frozen conflict.[^strict_frozen] Finally, an episode could not result in actor death since this is a minimum necessary condition for a conflict to become frozen.
+For a state actor this meant that although there could be a turnover in the regime, the state could not cease to exist as a sovereign entity following the conclusion of the conflict episode.
+For non-state actors this entailed a continued existence for at least five years post-conflict without renouncing political violence and/or a re-escalation of violence into a new conflict episode within the same time period.
+This was done to exclude situations where rebel groups unilaterally renounce violence as a political strategy and either disband or transition into political parties or civil society organizations.
+
+[^protocol]: For a detailed description of the codcing procedure see the protocol document in the supplemental materials.
+
+[^partial]: Note, this does not exclude tha possibility that frozen conflicts can arise following a partial peace agreement or ceasefire.
+For example, the First Nagorno-Karabkh War was terminated with the 1994 Bishkek Protocol establishing a general ceasefire; however, the agreement failed to address the status of Nagorno-Karabakh leading to a frozen conflict with periodic re-escalations until the forcible reintegration of the territory by Azerbaijan following the 2023 offensive.
+
+[^strict_frozen]: In the appendix I extend the analysis presented in this paper and run a series of robustness checks where I increase the number of years required for a compehensive peace agreement to be in place in order to better distingiush between frozen conflicts and short-term pauses in fighting.
 
@@
 
-The resulting dataset consists of `r n_frozen` frozen conflicts out
-of `r nrow(dataset)` intra-state episodes, comprising approximately
-`r signif(n_frozen / nrow(dataset), 2) * 100` percent of all observations for the period 1975 to 2019.
-Of the frozen conflicts, `r n_terr_frozen` are conflicts over territory.
-(TODO: show world map and time plot)
+In total, `r n_frozen` frozen conflicts were identified out of `r nrow(dataset)` civil conflict terminations, comprising approximately `r sprintf("%.0f", 100 * n_frozen / nrow(dataset))` percent of all observations for the period 1975 to 2017.
+Of the frozen conflicts, `r n_terr_frozen` were conflicts where the central incompatibility between conflict actors concerned territory rather than control of the central government.
+@fig-frozen shows both a world map of the countries that have experienced at least one frozen conflict within the coded time period as well as the incidence rate of frozen conflicts per year.
 
 ```{r}
+#| fig-height: 2
+#| label: fig-frozen
+#| fig-cap: |
+#|    On the left, a world map showing the geographic distribution of
+#|    countries that have experienced at least one frozen conflict
+#|    highlighted in blue. On the right, the incidence rate of new
+#|    frozen conflicts per year.
+
 frozen <- filter(dataset, frozen == 1) |>
@@
 world <- map_data("world") |>
+    filter(region != "Antarctica") |>
     left_join(sub.df, by = c("region" = "country")) |>
-    mutate(n = ifelse(is.na(n), 0, 1) )
+    mutate(n = ifelse(is.na(n), 0, 1) |> as.factor())
 
-ggplot(world, aes(long, lat, map_id = region, fill= n, group = group)) +
-    geom_map(map = world) +
-    theme(legend.position = "none") +
-    scale_color_solarized()
+p1 <- ggplot(world, aes(long, lat, map_id = region, fill = n, group = group)) +
+    geom_map(map = world, color = "grey", size = 0.25) +
+    scale_fill_manual(values = c("0" = "#CFCFC4", "1" = "#268bd2"), guide = F) +
+    theme(legend.position = "none",
+          axis.title.x = element_blank(),
+          axis.title.y = element_blank(),
+          axis.ticks = element_blank(),
+          axis.text = element_blank())
+
+p2 <- filter(dataset, frozen == 1) |>
+    ggplot(aes(year)) +
+    geom_bar() +
+    theme(axis.title.x = element_blank()) +
+    ylab("Frozen Onsets")
+
+p1 + p2
+
 ```
 
-```{r}
-filter(dataset, frozen == 1) |>
-ggplot(aes(year)) +
-    geom_bar() +
-    xlab("Year") +
-    ylab("Frozen Civil Conflict Onsets")
-```
+Compared to the only other extant coding of frozen conflicts, the Frozen Conflict Dataset (FCD) [@klosek_frozen_2021], the measure that I employ features a more expansive operationalization of the concept of frozen conflicts.
+Although the FCD covers a wider period from 1946 until 2011, only civil conflicts involving *de facto* states according to the coding by @Florea2014 are included resulting in 17 cases of frozen conflict.
+Furthermore, the termination year for a civil conflict episode in the FCD is coded as the last year in which the number of battle related deaths is above 500.
+In contrast, my measure looks beyond only *de facto* states and considers civil conflicts where the central incompatibility may center over control of the government.
+Furthermore, I follow the UCDP and employ a smaller battle related fatalities threshold thereby capturing lower intensity conflict episodes.[^insurgencies]
 
-Compared to the Frozen Conflict Dataset by @klosek_frozen_2021,
-
-Compared to FCD --- significant more cases
-  - More expansive operationalization
-    - Not only looking at de facto states
-    - Captures frozen moments in civil wars over government, ex Angola
-
-  - Timing of Frozen conflicts
-    - FCD last year in which >= 500 BRD
-    - My datasets follows UCDP last year >25
-      - Captures lower level conflicts
-
-Potential Weaknesses
-  - Low-level long running insurgencies may not be "salient"
-  - Termination year inexact
-  - Potentially capturing shorter stalemates
+[^insurgencies]: There is a risk that including low-level conflict episodes risks conflating minor, long-running insurgencies with frozen conflicts.
+It can be argued that the former may persist at such a low level of activity with only sporadic, isolated incidences of limited violence as to not represent a meaningfully salient issue to the incumbent state and would therefore not constitute a frozen conflict.
+To mitigate this risk, I also run the analysis presented in this paper on a sub-sample of conflicts with a cumulative intensity of at least 1,000 battle related fatalities since onset.
 
 ### Treatments
+
+The primary independent variables in the model are two binary treatment variables measuring whether external support was provided to the government side and whether external support was provided to the rebel side at any time during a civil conflict episode by a third party state actor, as well as the interaction in order to account for a differential effect when both sides receive support.
+Data on external state support was drawn from the UCDP External Support Dataset [@meier_external_2023] covering the period from 1975 to 2017.
 
@@
 ```
-
-Figure 2 shows the average marginal effects, *i.e.* the change in the posterior predicted probability between each treatment condition and the control condition averaged across the sample.
-
-\begin{equation}
-\text{AME}_j = \sum_{i=1}^N \Phi(\alpha + X_i' \beta + \delta_i) - \Phi(\alpha + X_i' \beta)
-\end{equation}
 

### 2024-02-06 lib/fc.utils/R/fc.utils.R
Roxygen has deprecated documenting NULL

@@ -3,6 +3,5 @@
 #' Utility functions used both in the construction of the fc-dataset and
 #' analysis of frozen conflict onsets.
 #'
-#' @docType package
-#' @name fc.utils
-NULL
+#' @keywords internal
+"_PACKAGE"

### 2024-02-06 paper.qmd
Replace geom_histogram with base hist
ggplot2 is great for everything except histograms.

@@
 library(ggdist)
+library(latex2exp)
 library(knitr)
@@
 sbc <- readRDS("./posteriors/sbc.rds") |>
-    select(-.chain, -.iteration, -.draw) |>
-    pivot_longer(everything())
+    select(-.chain, -.iteration, -.draw)
 
-ggplot(sbc, aes(x = value)) +
-    geom_histogram(bins = 10) +
-    facet_wrap(~ name, scales = "free") +
-    theme(axis.title.y = element_blank(),
-          axis.title.x = element_blank())
+# I really, really hate ggplot histograms
+par(family = "TeX Gyre Pagella", mfrow = c(5, 5),
+    mai = c(0.35, 0.2, 0.1, 0.2), mgp = c(1.75, 0.5, 0))
+for (i in seq_along(sbc)) {
+    # I also hate tibbles with every fiber of my being
+    label <- colnames(sbc)[i] |> stanvar_latex()
+    sbc[[i]] |> hist(main = "", xlab = TeX(label), ylab = "")
+}
 ```
@@
 ```{R}
-pzero <- function(x) mean(x > 0)
-
 models <- list.dirs("posteriors", recursive = F)

### 2024-02-06 lib/fc.utils/R/utils.R

@@ -1,10 +1,3 @@
-#' @export
-extract_digits <- function(x) {
-    m <- gregexpr("\\d", x)
-    regmatches(x, m)
-}
-
-
 #' Find consecutive elements
 #'
 #' Given a NumericVector, return the grouping indices for consecutive
@@ -24,4 +17,21 @@ consecutive.numeric <- function(x) {
         warning("NA's in given vector")
 
     cumsum(c(T, diff(x) != 1))
+}
+
+#' @export
+extract_digits <- function(x) {
+    m <- gregexpr("\\d", x)
+    regmatches(x, m)
+}
+
+#' @export
+pzero <- function(x) mean(x > 0)
+
+#' @export
+stanvar_latex <- function(s) {
+    if (grepl("\\[", s))
+        sub("(\\S+)\\[(\\d(,\\d)?)\\]", "$\\\\\\1_{\\2}$", s)
+    else
+        sprintf("$\\%s$", s)
 }

### 2024-02-06 paper.qmd
Add Florea 2014 reference

@@
 Compared to the only other extant coding of frozen conflicts, the Frozen Conflict Dataset (FCD) [@klosek_frozen_2021], the measure that I employ features a more expansive operationalization of the concept of frozen conflicts.
-Although the FCD covers a wider period from 1946 until 2011, only civil conflicts involving *de facto* states according to the coding by @Florea2014 are included resulting in 17 cases of frozen conflict.
+Although the FCD covers a wider period from 1946 until 2011, only civil conflicts involving *de facto* states according to the coding by @florea_facto_2014 are included resulting in 17 cases of frozen conflict.
 Furthermore, the termination year for a civil conflict episode in the FCD is coded as the last year in which the number of battle related deaths is above 500.

### 2024-02-07 paper.qmd
Draft description of treatments

@@
 
-The primary independent variables in the model are two binary treatment variables measuring whether external support was provided to the government side and whether external support was provided to the rebel side at any time during a civil conflict episode by a third party state actor, as well as the interaction in order to account for a differential effect when both sides receive support.
-Data on external state support was drawn from the UCDP External Support Dataset [@meier_external_2023] covering the period from 1975 to 2017.
+The primary independent variables in the model are two binary treatment variables measuring respectively whether external support was provided at least once to either the rebel side or the government side at any time during a civil conflict episode by a third party state actor, as well as the interaction in order to account for a potential differential effect when both sides receive support.
+Data on external state support was from the UCDP External Support Dataset (ESD) [@meier_external_2023] covering the period from 1975 to 2017 and included both fungible and non-fungible support across all support categories coded by the ESD.
 
-UCDP external support dataset, binary variable for whether an external
-state actor offered either fungible or non-fungible support at least
-once to either the rebels or government side in a conflict episode.
-
-- Include interaction term, three treatment conditions:
-
-1. No support (control group)
-2. Gov. support only
-3. Reb. support only
-4. Both support
+@fig-prop shows the total counts as well as proportion of observations that terminated as frozen conflicts for each corresponding treatment category: external support to only the government, external support to only the rebels, external support to both, and no external support received by either side.
+In general, a higher proportion of civil conflicts centered on a territorial incompatibility terminate as a frozen conflict when support is either given only to the rebels or to both sides.
 
 ```{r}
+#| fig-height: 3
+#| label: fig-prop
+#| fig-cap: |
+#|     Total counts and proportion of observations that terminated as
+#|     frozen conflict per treatment category.
+
 conditions <- dataset |>
-    mutate(treatment = case_when(ext_sup_s_rebel_bin == 1 & ext_sup_s_state_bin == 1 ~ "Ext. Sup. to Both",
-                                 ext_sup_s_rebel_bin == 1 & ext_sup_s_state_bin == 0 ~ "Ext. Sup. Rebel Only",
-                                 ext_sup_s_rebel_bin == 0 & ext_sup_s_state_bin == 1 ~ "Ext. Sup. Gov Only",
-                                 ext_sup_s_rebel_bin == 0 & ext_sup_s_state_bin == 0 ~ "Ext. Sup. to Neither"))
+    mutate(treatment = case_when(ext_sup_s_rebel_bin == 1 & ext_sup_s_state_bin == 1 ~ "Both",
+                                 ext_sup_s_rebel_bin == 1 & ext_sup_s_state_bin == 0 ~ "Rebel",
+                                 ext_sup_s_rebel_bin == 0 & ext_sup_s_state_bin == 1 ~ "Gov.",
+                                 ext_sup_s_rebel_bin == 0 & ext_sup_s_state_bin == 0 ~ "None"),
+           treatment = factor(treatment, levels = c("Gov.", "Rebel", "Both", "None")),
+           Incompatibility = ifelse(incompatibility == 0, "Government",
+                                    "Territory"))
 
-m <- table(conditions$frozen, conditions$treatment)
-rownames(m) <- c("Not Frozen", "Frozen")
-knitr::kable(m)
+p1 <- ggplot(conditions, aes(treatment, fill = Incompatibility)) +
+    geom_bar(position = "dodge") +
+    ylab("Total Count") +
+    scale_fill_solarized()
+
+condition_props <- group_by(conditions, treatment, Incompatibility) |>
+    summarise(prop = mean(frozen == 1))
+
+p2 <- ggplot(condition_props, aes(treatment, prop, fill = Incompatibility)) +
+    geom_col(position = "dodge") +
+    ylim(c(0, 1)) +
+    ylab("Proportion Frozen") +
+    scale_fill_solarized()
+
+p1 + p2 + plot_layout(axis_title = "collect", guides = "collect") & xlab("External Support")
+
 ```

### 2024-02-09 paper.qmd
Draft edits

@@
 The primary independent variables in the model are two binary treatment variables measuring respectively whether external support was provided at least once to either the rebel side or the government side at any time during a civil conflict episode by a third party state actor, as well as the interaction in order to account for a potential differential effect when both sides receive support.
-Data on external state support was from the UCDP External Support Dataset (ESD) [@meier_external_2023] covering the period from 1975 to 2017 and included both fungible and non-fungible support across all support categories coded by the ESD.
+Data on external state support was drawn from the UCDP External Support Dataset (ESD) [@meier_external_2023] covering the period from 1975 to 2017 and included both fungible and non-fungible support across all support categories coded by the ESD.
 
@@
 
+In addition to the treatment variables I also include a set of covariates that could plausibly confound the relationship between external state support and frozen conflicts.
+This includes variables related to the following conflict dynamics: conflict episode duration, episode reoccurence within a conflict, and cumulative intensity.
+Conflicts which last longer may affect the willing of third party actors to offer support to conflict actors.
+Simulatenously, long-running conflicts may also be at a greater risk of becoming frozen since the protracted nature of the conflict indicates an inability for either side to achieve a one-sided victory or be willing to reach a compremise through a peace agreement.
+We could expect the costs to increase such that active fighting ceases; however, the underlying conflict remains unresolved and the conflict transitions to being frozen.
+Thus, I include the log of the duration for each episode.
+A similar logic may also apply to episode reoccurence.
+Recurring episodes can also indicate an intractable long-running conflict that may affect both external support and the likelihood for a conflict to become frozen.
+Therefore, I include a binary indicator variable for whether an episode is is a recurring episode within a larger conflict as identified by the UCDP.
+Finally,
+
+This includes three variables related to conflict dynamics: conflict episode duration, episode reoccurence within a conflict, and cumulative intensity.
+Conflicts which last longer may affect the willingness of third party actors to provide support to conflict actors.
+Either
+
 - Episode censored
@@
 Using its builtin No-U-Turn sampler run for 2,000 iterations
+
+In other words, the crucial predictor for whether an intra-state conflict terminates as a frozen conflict is whether the opposition received external support by a third party state during the conflict episode
+
+. Meanwhile, external support directed to the government is fairly uninformative for the outcome.
 

### 2024-02-12 R/models.R
Switch model input datasets directory

@@ -3,7 +3,7 @@
 library(dplyr)
 library(fc.utils)
 
-dir.create("data/models", showWarnings = F)
+dir.create("data/model_inputs", showWarnings = F)
 df <- readRDS("./data/merged_data.rds") |> ungroup()
 
 ###
@@ -11,14 +11,14 @@ df <- readRDS("./data/merged_data.rds") |> ungroup()
 treatments <- c("bin", "bin_5y", "prop")
 outcomes <- c("frozen", "strict_frozen")
 units <- c("all", "cumulative_intensity")
-censored <- c("no_censored", "no_episode_censored", "with_censored")
+censored <- c("with_censored", "no_censored", "no_episode_censored")
 
 design <- expand.grid("outcome" = outcomes, "treatment" = treatments,
                       "episodes" = units, "censored" = censored) |>
     mutate(id = sprintf("%04d", row_number()))
 
 # Save the design matrix
-saveRDS(design, "./data/models/design.rds")
+saveRDS(design, "./data/model_inputs/design.rds")
 
 for (row in 1:nrow(design)) {
     info("Model %s", design[row, "id"])
@@ -31,20 +31,21 @@ for (row in 1:nrow(design)) {
 
     # Filter out censored observations if necessary, ie conflicts
     # beginning before 1975
-    if (design[row, "censored"] == "no_censored") {
+    if (design[row, "censored"] == "no_censored")
         sub.df <- filter(sub.df, censored == 0)
-    } else if (design[row, "censored"] == "no_episode_censored") {
+    else if (design[row, "censored"] == "no_episode_censored")
         sub.df <- filter(sub.df, episode_censored == 0)
-    }
 
     # Select control variables - this is mostly static across models
     X <- select(sub.df, episode_censored, episode_duration, recur, cumulative_intensity,
                 incompatibility, cold_war, ongoing_intrastate, ongoing_interstate)
 
     # If we've dropped censored episodes, no need for binary control
-    # var
+    # var similar for cumulative intensity episodes
     if (design[row, "censored"] != "with_censored")
         X <- select(X, -episode_censored)
+    else if (design[row, "episodes"] == "cumulative_intensity")
+        X <- select(X, -cumulative_intensity)
 
     X <- mutate(X, episode_duration = log(episode_duration) |> scale()) |>
         data.matrix()
@@ -62,6 +63,6 @@ for (row in 1:nrow(design)) {
[…]

### 2024-02-12 R/probit.R

@@ -7,10 +7,10 @@ library(tools)
 
 options(mc.cores = 1)
 
-design <- readRDS("./data/models/design.rds")
+design <- readRDS("./data/model_inputs/design.rds")
 
 input <- commandArgs(trailingOnly = T)
-default <- "data/models/data_0001.RData"
+default <- "data/model_inputs/0001.RData"
 model_data <- if (!exists("input") || length(input) == 0) default else input
 
 stopifnot(file.exists(model_data))

### 2024-02-12 R/merge.R
Ensure that we drop right censored episodes
ESD only covers from 1975 until 2017.

@@ -11,11 +11,16 @@ library(tidyr)
 frozen <- readRDS("./data/raw/frozen_conflicts.rds") |>
     select(conflict_id, year, duration, frozen)
 
+# There is a bug in the UCDP termination dataset where 2020
+# observations for two conflicts have been erroneously given a
+# recycled conflictep_id (conflict_id 265 and 353). Our data doesn't
+# extend beyond 2017 anyways, so just drop those censored episodes.
 term <- read_excel("./data/raw/ucdp-term-acd-3-2021.xlsx") |>
-    filter(type_of_conflict %in% 3:4) |>
+    filter(type_of_conflict %in% 3:4, !(conflict_id %in% c(265, 353) & year == 2020)) |>
     select(conflict_id, conflictep_id, year, gwno_a = gwno_loc, side_a, side_b,
            intensity_level, incompatibility, recur) |>
     group_by(conflictep_id) |>
+    filter(max(year) <= 2017) |>
     mutate(episode_censored = ifelse(min(year) < 1975, 1, 0),
            recur = max(recur),
            gwno_a = as.numeric(gwno_a)) |>
@@ -49,7 +54,7 @@ ongoing <- filter(ucdp, intensity_level == 2) |>
 # Cumulative intensity per conflict
 intensity <- select(ucdp, conflict_id, year, cumulative_intensity)
 
-ep <- filter(term, between(year, 1975, 2019)) |>
+ep <- filter(term, year >= 1975) |>
     full_join(frozen, by = c("conflict_id", "year")) |>
     left_join(intensity, by = c("conflict_id", "year")) |>
     left_join(ongoing, by = c("gwno_a" = "gwno", "year")) |>

### 2024-02-12 R/merge.R
Drop conflict episodes not coded by ESD

@@ -85,7 +85,10 @@ state_sup <- filter(esd, actor_nonstate == 0) |>
 
 full_ep.df <- full_join(rebel_sup, state_sup, by = c("conflict_id", "year"),
                     suffix = c("_rebel", "_state")) |>
-    right_join(reduced, by = c("conflict_id", "year"))
+    right_join(reduced, by = c("conflict_id", "year")) |>
+    group_by(conflictep_id) |>
+    filter(all(!is.na(ext_sup_s_state))) |>
+    ungroup()
 
 ###
 # NMC - CINC data

### 2024-02-12 paper.qmd
Switch model directories for manuscript

@@
 author: J.L.A. Krusell
-date: "`r Sys.Date()`"
+date: today
 linestretch: 1.25
@@
 
-design <- readRDS("./data/models/design.rds")
+design <- readRDS("./data/model_inputs/design.rds")
 ```
@@
 
-fit <- sprintf("posteriors/data_%s/fit.rds", main_model) |>  readRDS()
+fit <- sprintf("posteriors/%s/fit.rds", main_model) |>  readRDS()
 

### 2024-02-12 paper.qmd
Follow bayesplot and use 90% CIs

@@
 ggplot(delta, aes(x = value, y = variable, fill = variable)) +
-    stat_halfeye(alpha = 0.8, show.legend = F) +
+    stat_halfeye(alpha = 0.8, show.legend = F, .width = c(0.5, 0.9)) +
     geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.2) +
@@
 ggplot(ame, aes(x = value, y = variable, fill = variable)) +
-    stat_eye(alpha = 0.8, show.legend = F) +
+    stat_eye(alpha = 0.8, show.legend = F, .width = c(0.5, 0.9)) +
     geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.2) +

### 2024-02-12 paper.qmd
Ignore the AME calculated for the interaction

@@
 ame <- fit$draws("ame", format = "df") |>
-    select(matches("ame")) |>
+    select(matches("ame\\[[124]")) |>
     pivot_longer(everything(), names_to = "variable")
@@
 ```{R}
+
 models <- list.dirs("posteriors", recursive = F)
@@
     fits[[model]]$summary("ame", pzero, posterior::quantile2, median) |>
-                    mutate(id = sub("data_", "", model))
+                    mutate(id = sub("data_", "", model)) |>
+                    filter(grepl("ame\\[[124]", variable))
 }) |>
@@
                                  ame_keys == "ame[2]" ~ "Ext. Sup. Rebel Only",
-                                 ame_keys == "ame[3]" ~ "Ext. Sup. to Both")
+                                 ame_keys == "ame[4]" ~ "Ext. Sup. to Both")
 model_ames$group <- ifelse(sapply(ll, `[`, 2) == 1, "Government Incompatibility",

### 2024-02-12 scripts/wordcount.sh
Include figure captions in word count
Easier to split this code into a separate script that trying to inline
everything in the Makefile.

@@ -1,0 +1,17 @@
+#!/usr/bin/env zsh
+# Approximate word count for a quarto/Rmarkdown file. Includes body
+# text count + references, abstract, and figure captions while
+# excluding appendix and code blocks.
+###
+
+abstract=$(yq -f extract '.abstract' $1 | wc -w)
+
+body=$(sed -e '/^```/,/^```/d' -e '/^#\s*Appendix/,$d' $1 | \
+           pandoc --quiet --citeproc -f markdown -t plain | \
+           wc -w)
+
+# This assumes that only fig-cap values are multiline within code
+# blocks
+figure_captions=$(grep '^#|' $1 | sed -e 's/^#|\s*//' -e '/^.*:/d' | wc -w)
+
+print "$1: $((abstract + body + figure_captions))"

### 2024-02-12 paper.qmd
Manuscript edits

@@
 Due to the non-linear nature of a probit regression model, in order to interpret the estimated association between the treatments and outcome, the target empirical estimands are the group-specific average marginal effects, *i.e.* the changes in the conditional expectation of the outcome variable given the various treatment states versus the baseline control averaged over the relevant subgroup sample.
-Let $\theta$ denote the vector of all unknown parameters in the regression model, then the average marginal effect (AME) for the $j^{th}$ treatment state represented by vector $w_j$ for subsample $k$ can be approximated using $m = 1, 2, \ldots, M$ posterior draws from $\theta$ such that
+Let $\theta$ denote the vector of all unknown parameters in the regression model, then the average marginal effect (AME) for the $j^{th}$ treatment state represented by vector $w_j$ for subsample $k$ can be approximated using $m = 1, 2, \ldots, M$ posterior draws from $p(\theta | y, X, T)$ such that
 
@@
 Data on external state support was drawn from the UCDP External Support Dataset (ESD) [@meier_external_2023] covering the period from 1975 to 2017 and included both fungible and non-fungible support across all support categories coded by the ESD.
+As detailed in the model specification, I allow the effect parameters to vary according to whether the underlying incompatibility concerns control of the government or territory resulting in two sets of treatment effects.
 
@@
 In addition to the treatment variables I also include a set of covariates that could plausibly confound the relationship between external state support and frozen conflicts.
-This includes variables related to the following conflict dynamics: conflict episode duration, episode reoccurence within a conflict, and cumulative intensity.
-Conflicts which last longer may affect the willing of third party actors to offer support to conflict actors.
-Simulatenously, long-running conflicts may also be at a greater risk of becoming frozen since the protracted nature of the conflict indicates an inability for either side to achieve a one-sided victory or be willing to reach a compremise through a peace agreement.
+This includes variables related to  conflict dynamics, specifically: conflict episode duration, episode recurrence within a conflict, and cumulative intensity.
+Conflicts which last longer may affect the willingness of third party actors to offer support to belligerents.
+Simultaneously, long-running conflicts may also be at a greater risk of becoming frozen since the protracted nature of the conflict indicates an inability for either side to achieve a one-sided victory or be willing to reach a compromise through a peace agreement.
 We could expect the costs to increase such that active fighting ceases; however, the underlying conflict remains unresolved and the conflict transitions to being frozen.
-Thus, I include the log of the duration for each episode.
-A similar logic may also apply to episode reoccurence.
+Thus, I include the number of years for which each episodes lasts until termination transformed with the natural logarithm function in order to reflect the diminishing impact of each one unit increase.
+A similar logic may also apply to episode recurrence.
 Recurring episodes can also indicate an intractable long-running conflict that may affect both external support and the likelihood for a conflict to become frozen.
-Therefore, I include a binary indicator variable for whether an episode is is a recurring episode within a larger conflict as identified by the UCDP.
-Finally,
+Therefore, I include a binary indicator variable for whether an episode is a recurring episode within a larger conflict as identified by the UCDP.
+Finally, I also include a binary indicator for whether the cumulative number of battle related deaths has exceeded 1,000 since onset.
+High intensity conflicts would be indicative of a greater relative strength for the opposition and may be more likely to attract outside intervention both because of the potential destabilizing effects to international/regional security as well as the greater opportunity for third party state rivals to secure a favourable post-war outcome.
+Furthermore, a stronger opposition may also raise the probability that the conflict becomes frozen since the incumbent government may be less likely to win an outright victory.
 
-This includes three variables related to conflict dynamics: conflict episode duration, episode reoccurence within a conflict, and cumulative intensity.
-Conflicts which last longer may affect the willingness of third party actors to provide support to conflict actors.
-Either
+In complement to the partially pooled varying treatment effects, I also include a binary indicator for whether the conflict incompatibility concerns territory in order to account for a potential differing baseline probability for termination as a frozen conflict and the possibility that the conflict incompatibility affects the likelihood of external support.
+Further, I include variables to control for the Cold War and ongoing inter/intra-state conflicts within the host country.
+The former is a binary indicator for whether the termination year falls before 1989 to capture temporal dynamics related to the propensity for different types of conflict terminations.
+For example, international norms concerning external intervention in civil conflicts may be shaped in part by the state of the international system influencing both the decision by third party actors to intervene in a civil conflict and the types of conflict terminations that external interveners seek [@howard2017civil].
+The latter are binary indicators as well measuring whether an interstate or intrastate conflict occurred at any time during the conflict episode.
+Third party states involved in one conflict may be more likely to also intervene in another ongoing conflict within the same country.
+Furthermore, multiple ongoing conflicts could also indicate a weakened incumbent government forced to face multiple opponents, which may in turn affect the probability that a conflict freezes.
+Lastly, I include a binary variable for whether a conflict is left censored, *i.e.*, that the onset of the episode occurred before the coding start for the dataset, since data on external support from the ESD is only available from 1975 until 2017.[^robust_censor]
 
-- Episode censored
-- Episode Duration log
-- Recur
-- Cumulative Intensity
-- Incompatibility
-- Cold War
-- Ongoing intra/interstate conflict
+[^robust_censor]: I also run multiple robustness checks in the appendix where I alternatively drop episodes with an episode onset year before the coding start for the dataset and/or recurrent episodes within conflicts that also began prior to 1975.
 
-Descriptive statistics split between positive and negative outcomes
+## Results
 
-### Results
+@fig-delta shows the posterior densities for the three treatment parameters varied by conflict incompatibility.
 
@@
 ```{r}
-#| fig.cap: Posterior estimates for \delta from a single variate probit model
+#| label: fig-delta
+#| fig-cap: |
+#|     Posterior densities as well as 50% and 90% credible intervals
+#|     for the treatment parameters in the hierarchical probit model
+#|     predicting whether a civil conflict episodes terminates as a
+#|     frozen conflict. Point estimates are the posterior medians.
+
 main_model <- filter(design, outcome == "frozen", treatment == "bin", episodes == "all",
@@
 ```{r}
-#| fig.cap: Average marginal effects for the three treatment conditions on the predicted probability of a frozen conflict onset.[^ame]
+#| label: fig-ame
+#| fig-cap: |
+#|     Posterior densities for the average marginal effects for the
+#|     three treatment conditions on the predicted probability of a
+#|     frozen conflict onset with 50% and 90% credible
+#|     intervals. Point estimates are the posterior medians.
+
 ame <- fit$draws("ame", format = "df") |>
@@
     geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.2) +
-    theme(axis.title.x = element_blank()) +
-    labs(x = parse(text = "bold('AME '*Delta)")) +
-    scale_y_discrete(labels = c("ame[1]" = "Ext. Sup. Gov Only",
-                                "ame[2]" = "Ext. Sup. Rebel Only",
-                                "ame[3]" = "Ext. Sup. to Both")) +
+    labs(x = parse(text = "bold('AME '*Delta)"),
+         y = "External Support") +
+    scale_y_discrete(labels = c("ame[1]" = "Gov. Only",
+                                "ame[2]" = "Rebel Only",
+                                "ame[4]" = "Both")) +
     coord_flip() +

### 2024-02-12 paper.qmd
Draft results section

@@
 
-Due to the non-linear nature of a probit regression model, in order to interpret the estimated association between the treatments and outcome, the target empirical estimands are the group-specific average marginal effects, *i.e.* the changes in the conditional expectation of the outcome variable given the various treatment states versus the baseline control averaged over the relevant subgroup sample.
-Let $\theta$ denote the vector of all unknown parameters in the regression model, then the average marginal effect (AME) for the $j^{th}$ treatment state represented by vector $w_j$ for subsample $k$ can be approximated using $m = 1, 2, \ldots, M$ posterior draws from $p(\theta | y, X, T)$ such that
+Due to the non-linear nature of a probit regression model, in order to interpret the estimated association between the treatments and outcome, the target empirical estimands are the group-specific average marginal effects, *i.e.* the changes in the conditional expectation of the outcome variable given the various treatment states versus the baseline control averaged over the relevant subgroup sample.[^mem] Let $\theta$ denote the vector of all unknown parameters in the regression model, then the average marginal effect (AME) for the $j^{th}$ treatment state represented by vector $w_j$ for subsample $k$ can be approximated using $m = 1, 2, \ldots, M$ posterior draws from $p(\theta | y, X, T)$ such that
 
@@
 \end{align*}
+
+[^mem]: An alternative approach would be to calculate the differences in the predicted probability for each treatment effect while holding the remaining covariates constant at their mean and/or mode.
+However, this would not necessarily be representative of the "typical case" in the population [@hanmer_behind_2013].
+Although there is a minimal amount of non-stochastic missingness in the data I employ due to differing coding coverage between datasets resulting in dropped conflict episodes, I would argue that the sample averaged marginal effects are still more representative of each treatment would affect the typical civil conflict.
 
@@
 The former is a binary indicator for whether the termination year falls before 1989 to capture temporal dynamics related to the propensity for different types of conflict terminations.
-For example, international norms concerning external intervention in civil conflicts may be shaped in part by the state of the international system influencing both the decision by third party actors to intervene in a civil conflict and the types of conflict terminations that external interveners seek [@howard2017civil].
+For example, international norms concerning external intervention in civil conflicts may be shaped in part by the state of the international system influencing both the decision by third party actors to intervene in a civil conflict and the types of conflict terminations that external interveners seek [@howard_how_2017].
 The latter are binary indicators as well measuring whether an interstate or intrastate conflict occurred at any time during the conflict episode.
@@
 
-@fig-delta shows the posterior densities for the three treatment parameters varied by conflict incompatibility.
+@fig-ame plots the average marginal effects from estimating the hierarchical probit model predicting frozen civil conflicts given the observed data on directed external support by third party states and the set of control variables detailed in the data section.[^stan] The average marginal effects capture the change in the predicted probability for the outcome given the three treatment conditions --- external support given only to the government, external support given only to the rebel(s), and external support given to both sides --- versus the baseline condition of no external support given to any conflict actor.
+The plots also depict the 50% and 90% credible intervals formed from the corresponding quantiles of the posterior draws with the point estimates representing the posterior medians.
 
-The actual model is implemented in Stan, a probabilistic programming language
-
-Using its builtin No-U-Turn sampler run for 2,000 iterations
-
-In other words, the crucial predictor for whether an intra-state conflict terminates as a frozen conflict is whether the opposition received external support by a third party state during the conflict episode
-
-. Meanwhile, external support directed to the government is fairly uninformative for the outcome.
+[^stan]: The model is implemented in Stan, a probabilistic programming language, and estimated using Stan's No-U-Turn sampler via the `CmdStanR` interface [@stan_development_team_stan_2024; @gabry_cmdstanr_2024].
+It was run with 4 parallel chains at 2,000 iterations each with a warmup of 1,000 with convergence assessed using the standard Stan diagnostics.
+All pre/post-processing was done using R `r sprintf("%s.%s", R.version$major, R.version$minor)` [@r_core_team_r_2023].
 
 ```{r}
-#| label: fig-delta
+#| label: fig-ame
+#| fig-height: 3
 #| fig-cap: |
-#|     Posterior densities as well as 50% and 90% credible intervals
-#|     for the treatment parameters in the hierarchical probit model
-#|     predicting whether a civil conflict episodes terminates as a
-#|     frozen conflict. Point estimates are the posterior medians.
+#|     Posterior densities for the average marginal effects for the
+#|     three treatment conditions varying by incompatibility type on
+#|     the predicted probability of a
+#|     frozen conflict onset versus the baseline with 50% and 90%
+#|     credible intervals. Point estimates are the posterior medians.
 
@@
 fit <- sprintf("posteriors/%s/fit.rds", main_model) |>  readRDS()
+
+ame <- fit$draws("ame", format = "df") |>
+    select(matches("ame\\[[124]")) |>
+    pivot_longer(everything(), names_to = "variable")
+
+ll <- extract_digits(ame$variable)
+ame$variable <- sprintf("ame[%s]", sapply(ll, `[`, 1))
+ame$group <- ifelse(sapply(ll, `[`, 2) == 1, "Government Incompatibility",
+                    "Territorial Incompatibility")
+
+ame2 <- filter(ame, variable == "ame[2]", group == "Territorial Incompatibility")
+ame4 <- filter(ame, variable == "ame[4]", group == "Territorial Incompatibility")
+
+ggplot(ame, aes(x = value, y = variable, fill = variable)) +
+    stat_eye(alpha = 0.8, show.legend = F, .width = c(0.5, 0.9)) +
+    geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.2) +
+    labs(x = parse(text = "bold('AME '*Delta)"),
+         y = "External Support") +
+    scale_y_discrete(labels = c("ame[1]" = "Gov. Only",
+                                "ame[2]" = "Rebel Only",
+                                "ame[4]" = "Both")) +
+    coord_flip() +
+    facet_grid(. ~ group) +
+    scale_fill_solarized()
+```
+
+From the results, it is clear that only in instances of territorial civil conflict where external support was provided to either both sides or to only the rebel(s) is there a substantive effect on the probability of an episode becoming a frozen conflict.
+In both cases, the effect is positive with respective probabilities `r mean(ame4$value > 0) |> signif(2)` and `r mean(ame2$value > 0) |> signif(2)`, and on average, each condition is associated with an approximately `r mean(ame2$value) |> signif(2)` increase in the predicted probability for a frozen conflict across the sample of territorial civil conflict episodes, albeit with a high degree of posterior uncertainty, $\text{CI}_{90\%}: [`r quantile(ame4$value, 0.05) |> signif(2)`, `r quantile(ame4$value, 0.95) |> signif(2)`]$ and $\text{CI}_{90\%}: [`r quantile(ame2$value, 0.05) |> signif(2)`, `r quantile(ame2$value, 0.95) |> signif(2)`]$.
+
+```{r}
+#| label: fig-delta
+#| fig-height: 3
+#| fig-cap: |
+#|     Posterior densities as well as 50% and 90% credible intervals
+#|     for the treatment parameters varying by incompatibility type in
+#|     the hierarchical probit model predicting whether a civil
+#|     conflict episodes terminates as a frozen conflict. Point
+#|     estimates are the posterior medians.
 
@@
                       "Territorial Incompatibility")
+
+delta2 <- filter(delta, variable == "delta[2]", group == "Territorial Incompatibility")
 
@@
 
-```{r}
-#| label: fig-ame
-#| fig-cap: |
-#|     Posterior densities for the average marginal effects for the
-#|     three treatment conditions on the predicted probability of a
-#|     frozen conflict onset with 50% and 90% credible
-#|     intervals. Point estimates are the posterior medians.
+From the parameter plots in @fig-delta, the main driver for this substantive results is the estimate for external support given to the rebel(s) in the context of a territorial civil conflict.
+In this case, the probability for the treatment parameter being greater than zero is `r mean(delta2$value > 0) |> signif(2)` with $\text{CI}_{90\%}: [`r quantile(delta2$value, 0.05) |> signif(2)`, `r quantile(delta2$value, 0.95) |> signif(2)`]$.
+Taken together, these results indicate that the crucial predictor for whether an intra-state conflict terminates as a frozen conflict is whether the opposition received external support by a third party state during a conflict episode centered around a territorial incompatibility.
+Meanwhile, external support directed to the government is fairly uninformative as are all treatment effects under the condition of a government incompatibility.
+However, the posterior uncertainty remains high, which limits the degree to which the treatment effect can be substantively interpreted.
 
-ame <- fit$draws("ame", format = "df") |>
-    select(matches("ame\\[[124]")) |>
-    pivot_longer(everything(), names_to = "variable")
-
-ll <- extract_digits(ame$variable)
-ame$variable <- sprintf("ame[%s]", sapply(ll, `[`, 1))
-ame$group <- ifelse(sapply(ll, `[`, 2) == 1, "Government Incompatibility",
-                    "Territorial Incompatibility")
-
-ggplot(ame, aes(x = value, y = variable, fill = variable)) +
-    stat_eye(alpha = 0.8, show.legend = F, .width = c(0.5, 0.9)) +
-    geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.2) +
-    labs(x = parse(text = "bold('AME '*Delta)"),
-         y = "External Support") +
-    scale_y_discrete(labels = c("ame[1]" = "Gov. Only",
-                                "ame[2]" = "Rebel Only",
-                                "ame[4]" = "Both")) +
-    coord_flip() +
-    facet_grid(. ~ group) +
-    scale_fill_solarized()
-```
+\newpage
 

### 2024-02-16 paper.qmd
Draft edits

@@
 **Research question: to what extent does external state support to conflict actors affect the probability that a civil conflict becomes frozen?**
+
+# Literature Review
+
+## External Support and Civil Conflict
+
+In contrast to the paucity of literature concerning frozen conflicts, there exists numerous studies examining the links between third party involvement on both the duration and outcome of civil conflicts [@shelton2013we].
+Previous work has argued that external interventions by third party states increase the duration of civil conflicts [@regan2002third].
+However, biased military interventions may be the exception and in some cases actually increase the probability of a favourable outcome for the supported conflict actor [@balch2008third].
+Although others have put forth empirical arguments that this only holds in the case of rebel biased interventions.
+More specifically, third party military interventions may decrease the duration of a conflict and increase the probability of either negotiated settlement or rebel military victory in part because interveners may only involve themselves in ongoing civil conflicts in favor of the government when the government is facing a relatively strong rebel group with a high level of military capabilities [@gent2008going].
+The probability for a government military victory would on average be lower in these situations, thereby masking the effects of a government biased intervention.
+Indeed, when accounting for the differential effect of intervention given the strength of the opposition, there have been empirical claims that government biased interventions decrease the probability of a rebel victory and increase the probability of a government military victory [@sullivan2015conditional].
+
+Additional work has sought to disaggregate the types of external interventions that third party states may provide to conflict actors.
+This includes looking beyond military interventions with some scholars arguing that military aid and other forms of support, excluding direct military involvement, increase the likelihood of either a rebel victory or negotiated settlement [@jones2017altering].
+Others argue that fungible support, *i.e.* financing and weapons, has an opposite impact by actually extending the duration of conflicts [@sawyer2017role].
 
@@
 
+# Conclusion
+
+The measurement of frozen conflicts provided herewithin is crude. Future work should focus on refining this measurement.
+
 \newpage
@@
 
+\newpage
+
 ## Simulation Based Calibration
@@
 ```
+
+\newpage
 

### 2024-02-20 paper.qmd
Draft edits

--- 2024-02-12 paper.qmd
@@
 
-**Research question: to what extent does external state support to conflict actors affect the probability that a civil conflict becomes frozen?**
+Frozen conflicts are among some of the most pernicious, difficult to
+resolve conflicts. Surprisingly, little scholarly attention has been
+paid to the factors influencing their emergence. This paper addresses
+
+Frozen conflicts -> pernicious difficult to solve conflicts; long-running costs
+
+- relatively little academic literature on why some conflicts
+
+"Why do some conflicts become frozen?"
+
+- focus on the class of intra-state frozen conflicts
+
+- wide variety of potential factors
+  -  I decide to focus on international dimensions, specifically external support
+  - Justify by saying there exists a broad literatur eon the importance of external support on ocnlfict outcomes
+
+- Skeleton for remaining article
+  - First, I provide a brief overview of prev. literature
+  - I then offer a theory of why external support may be linked to the onset of frozen civil conflicts. "In brief, etc etc"
+  - Provide empirical evidence with original coding showing that support is positively linked with prob. of frozen conflict
+    - H/o: only within the context of territorial civil conflicts
+
+- Results contribute to our understanding of frozen conflitcs and conflict outcomes more broadly
+
+frozen conflict - "a protracted state of affairs in which major hostilities have ended, but the enmity between the actors persists and re-escalation remains a permanent threat."
+
+When conflicts freeze, large-scale hostilities cease, but the underlying issue(s) remain unresolved and there remains an ever-present threat of violent re-escalation.
+
+# Literature Review
+
+In contrast to the paucity of literature concerning frozen conflicts, there exists numerous studies examining the links between third party involvement on both the duration and outcome of civil conflicts [@shelton_what_2013].
+Previous work has argued that external interventions by third party states increase the duration of civil conflicts [@regan_third-party_2002].
+However, biased military interventions may be the exception and in some cases actually increase the probability of a favourable outcome for the supported conflict actor [@balch-lindsay_third-party_2008].
+Although others have put forth empirical arguments that this only holds in the case of rebel biased interventions.
+More specifically, third party military interventions may decrease the duration of a conflict and increase the probability of either negotiated settlement or rebel military victory in part because interveners may only involve themselves in ongoing civil conflicts in favor of the government when the government is facing a relatively strong rebel group with a high level of military capabilities [@gent_going_2008].
+The probability for a government military victory would on average be lower in these situations, thereby masking the effects of a government biased intervention.
+Indeed, when accounting for the differential effect of intervention given the strength of the opposition, there have been empirical claims that government biased interventions decrease the probability of a rebel victory and increase the probability of a government military victory [@sullivan_conditional_2015].
+
+Additional work has sought to look at the relationships and preferences of supporting actors as well as the disaggregated types of external support.
+Third party states intervening or providing material support to opposing sides in a civil conflict for the purpose of balancing may increase conflict duration [@aydin_networks_2012; @anderson_competitive_2019].
+
+disaggregate the types of external interventions that third party states may provide to conflict actors.
+This includes looking beyond military interventions with some scholars arguing that military aid and other forms of support, excluding direct military involvement, increase the likelihood of either a rebel victory or negotiated settlement [@jones_altering_2017].
+Others argue that fungible support, *i.e.* financing and weapons, has an opposite impact by actually extending the duration of conflicts [@sawyer_role_2017].
+
+# Theory
 
@@
 
+# Conclusion
+
+The measurement of frozen conflicts provided herewithin is crude. Future work should focus on refining this measurement.
+
+Potential shortcoming: what I am not doing is taking into account the preferences of external supporters and motivation for offering support.
+Nor am I distinguishing between different external support types.
+
 \newpage

### 2024-02-20 paper.qmd
Draft edits

@@
 Frozen conflicts are among some of the most pernicious, difficult to
-resolve conflicts. Surprisingly, little scholarly attention has been
+resolve conflicts.
+When conflicts freeze large-scale hostilities cease, but the underlying issue(s) remain unresolved with an ever present threat of recurrent violence.
+This state of unstable peace can be costly for actors as a certain degree of military preparedness needs to be maintained to secure transitory war gains lest the conflict reignites leading to an outright military defeat.
+
+Suprisingly, little scholarly attention has been paid to the factors influencing the emergence of frozen conflicts.
+Previous empirical work tends to consist of qualitative single-case studies [@klosek_frozen_2021].
+
+Surprisingly, little scholarly attention has been
 paid to the factors influencing their emergence. This paper addresses
-
-Frozen conflicts -> pernicious difficult to solve conflicts; long-running costs
 

### 2024-02-29 scripts/wordcount.sh
Add error checks to the wordcount script

@@ -1,8 +1,19 @@
 #!/usr/bin/env zsh
+#
 # Approximate word count for a quarto/Rmarkdown file. Includes body
 # text count + references, abstract, and figure captions while
 # excluding appendix and code blocks.
 ###
+
+setopt err_exit
+
+if [[ $# -ne 1 ]]; then
+    print -u 2 "Missing file argument"
+    print -u 2 "Usage: $0 <file.[qmd|Rmd|md]>"
+    exit 1
+fi
+
+[[ ! -f $1 ]] && { print -u 2 "Invalid file: $1"; exit 1 }
 
 abstract=$(yq -f extract '.abstract' $1 | wc -w)
 

### 2024-02-29 paper.qmd
Draft edits

--- 2024-02-20 paper.qmd
@@
 
-Frozen conflicts are among some of the most pernicious, difficult to
-resolve conflicts. Surprisingly, little scholarly attention has been
+Frozen conflicts are among some of the most pernicious, difficult to resolve conflicts.
+They are geographically and temporally dispersed, and although they are characterized by a cessation of large-scale hostilities, the underlying disputes remain unresolved.
+This puts actors into a state of uncertainty with an ever present threat of violent re-escalations and can lead to long-running accumulating costs.
+
+Surprisingly, little scholarly attention has been paid to the factors influencing the emergence of frozen conflicts [@klosek_frozen_2021].
+
+This paper seeks to address this gap by asking why do some conflicts become frozen?
+
+Surprisingly, little scholarly attention has been
 paid to the factors influencing their emergence. This paper addresses
-
-Frozen conflicts -> pernicious difficult to solve conflicts; long-running costs
 
@@
 # Theory
+
+## Defining Frozen Conflict
+
+I start with the definition of frozen conflict offered by @smetana_between_2019:
+
+> We define frozen conflict as a protracted, post-war conflict process, characterized by the absence of stable peace...
+[where] core issues between the opposing sides remain unresolved, the dispute is in the forefront of mutual relations, and there is a looming threat of the renewal of violence.
+
+Accompanying this definition, @smetana_between_2019 lay forth four criteria determining whether a post-conflict period qualifies as a frozen conflict.
+First, the conflict must be international since it is assumed that frozen conflict are a subset of international conflict featuring state or state-like actors.
+Second, there needs to be an initial period of active conflict followed by a cessation of large-scale violent hostilities.
+Isolated, sporadic incidences of violence may continue to occur; however, conflict actors are either unwilling or unable to launch major military offensives.
+Although not necessary, this may be characterized by the establishment of a general ceasefire and/or presence of external peacekeeping troops.
+Third, the core issue underlying the conflict needs to remain salient post-conflict.
+This entails actors maintaining their claims within a dispute.
+And, finally, there must be an ongoing threat of violent re-escalation.
+In other words, post-conflict relations should be characterized by mutual hostility and an unwillingness to abandon violence as a political strategy.
+
+These are precise criteria that expand the coverage of frozen conflicts both temporally and geographically, beyond the traditional scope of the post-Soviet separatist states that emerged in the early 1990s following the collapse of the Soviet Union.
+Furthermore, they hint at the dynamic nature of frozen conflicts --- far from a static condition, when a conflict freezes it may still be characterized by low levels of violence and periodic attempts at negotiations and international mediation.
+However, I deviate from these criteria in one crucial regard.
+I relax the assumption that frozen conflicts are a subset of international conflicts and assert that frozen conflicts are a subset of conflict more broadly speaking.
+In regards to civil conflicts, the @smetana_between_2019 criteria restrict frozen conflicts to *de facto* independent states.
+This precludes the possibility that non-secessionist conflicts can exhibit the same post-conflict characteristics of unstable peace that define frozen conflicts.
+
+Consider the case of the Second Libyan Civil War.
+In the aftermath of the overthrow of Muammar Gaddafi in 2011, central political authority fragmented and two competing governments emerged: the internationally recognized and UN sanctioned Government of National Accord (GNA) based in Tripoli and the so-called Tobruk Government comprised of the Libyan House of Representatives elected in the 2014 parliamentary elections aligned with the Libyan National Army under Khalifa Haftar.
+A ceasefire was signed in 2015 leaving the GNA nominally in control of Tripoli with the Tobruk government in control of the eastern portion of Libya.
+Intense fighting continued among other factions,
+
+[^libya_ceasefire]: https://www.reuters.com/article/us-libya-security-idUSKBN0KP0VL20150116/
+
+## How do frozen conflicts emerge?
+
+As an equilibrium state
+
+crucially, the expected utility of maintaining the dispute outweighs the potential gain from a peace agreement.
 
@@
 
+```{R cache=F}
+output <- system("scripts/wordcount.sh paper.qmd", intern = T)
+wc <- gsub("[^\\d]", "", output, perl = T)
+```
+
+**Word Count**: `r wc`
+
 \newpage

### 2024-03-01 paper.qmd
Draft edits

@@
 subtitle: "The Role of External State Support"
-author: J.L.A. Krusell
+authors:
+    - name: J.L.A. Krusell
+      affiliations:
+          - name: University of Gothenburg
 date: today
@@
 bibliography: library.bib
-format: pdf
+filters:
+    - lib/luafilter/meta.lua
+format:
+    pdf:
+        fontsize: 11pt
+        papersize: A4
+        include-in-header:
+            - text: |
+                \usepackage[font={small}]{caption}
+        template-partials:
+            - templates/before-body.tex
+            - templates/title.tex
 execute:
@@
 A ceasefire was signed in 2015 leaving the GNA nominally in control of Tripoli with the Tobruk government in control of the eastern portion of Libya.
-Intense fighting continued among other factions,
+However, the core contested issue, legitimized control of the central government, remained unresolved with
+
+dyadic frozen conflict
 
@@
 
-As an equilibrium state
+Frozen conflict as an equilibrium of a situation where both actors strictly prefer a military victory but are unable to achieve a significant military advantage and are unwilling to acquiesce to a peace agreement.
+The belligerents will seek to minimize costs by de-escalating the conflict intensity and the conflict will eventually freeze.
 
-crucially, the expected utility of maintaining the dispute outweighs the potential gain from a peace agreement.
+Assumes that the expected utility of maintaining the dispute is greater than the utility of a peace agreement.
+
+-  transitory war gains outweigh peaceful division of contested good
+- time horizon, low discount factor (Stall war paper assumes that actors actually discount the future?). Anticipation of a future favourable shift in the balance of power
+- credible commitment issues undermine payoff from a peace agreement
+
+Unlike conflict onsets, assume that fighting has revealed information about actors' resolve and capabilities.
 
@@
 
-```{r}
+```{r dataset}
 dataset <- readRDS("./data/merged_data.rds")
@@
 
-```{R cache=F}
-output <- system("scripts/wordcount.sh paper.qmd", intern = T)
-wc <- gsub("[^\\d]", "", output, perl = T)
-```
-
-**Word Count**: `r wc`
-
 \newpage
@@
 ```{r}
+#| label: tbl-frozen
 
@@
 ```{R}
+#| label: fig-sbc
+#| fig-cap: |
+#|     Simulation based calibration results
+
 sbc <- readRDS("./posteriors/sbc.rds") |>
@@
 ```{R}
+#| label: fig-robust
+#| fig-cap: Robustness checks
 

### 2024-03-01 scripts/wordcount.sh
Latex partial templates

@@ -25,4 +25,4 @@ body=$(sed -e '/^```/,/^```/d' -e '/^#\s*Appendix/,$d' $1 | \
 # blocks
 figure_captions=$(grep '^#|' $1 | sed -e 's/^#|\s*//' -e '/^.*:/d' | wc -w)
 
-print "$1: $((abstract + body + figure_captions))"
+print "$((abstract + body + figure_captions))"

### 2024-03-01 paper.qmd
Update Latex/Pandoc partial templates

@@
   affiliations:
-          - name: University of Gothenburg
+      - "University of Gothenburg"
 date: today
+date-format: long
 linestretch: 1.25
@@
     pdf:
-        fontsize: 11pt
-        papersize: A4
-        include-in-header:
-            - text: |
-                \usepackage[font={small}]{caption}
         template-partials:
@@
 ```
-
-# Introduction
 

### 2024-03-04 paper.qmd
Draft edits

@@
 - name: J.L.A. Krusell
+  email: "joshua.krusell@gu.se"
   affiliations:
@@
 format:
+    html:
+        toc: true
     pdf:
@@
 Accompanying this definition, @smetana_between_2019 lay forth four criteria determining whether a post-conflict period qualifies as a frozen conflict.
-First, the conflict must be international since it is assumed that frozen conflict are a subset of international conflict featuring state or state-like actors.
-Second, there needs to be an initial period of active conflict followed by a cessation of large-scale violent hostilities.
+First, the conflict must be international since it is assumed that frozen conflict is a subset of international conflict featuring state or state-like actors.
+Second, there needs to be an initial period of active, armed conflict followed by a cessation of large-scale violent hostilities.
 Isolated, sporadic incidences of violence may continue to occur; however, conflict actors are either unwilling or unable to launch major military offensives.
@@
 However, I deviate from these criteria in one crucial regard.
-I relax the assumption that frozen conflicts are a subset of international conflicts and assert that frozen conflicts are a subset of conflict more broadly speaking.
+I relax the assumption that frozen conflict is a subset of international conflict and assert that frozen conflict is a subset of conflict more broadly speaking.
 In regards to civil conflicts, the @smetana_between_2019 criteria restrict frozen conflicts to *de facto* independent states.
@@
 Consider the case of the Second Libyan Civil War.
-In the aftermath of the overthrow of Muammar Gaddafi in 2011, central political authority fragmented and two competing governments emerged: the internationally recognized and UN sanctioned Government of National Accord (GNA) based in Tripoli and the so-called Tobruk Government comprised of the Libyan House of Representatives elected in the 2014 parliamentary elections aligned with the Libyan National Army under Khalifa Haftar.
-A ceasefire was signed in 2015 leaving the GNA nominally in control of Tripoli with the Tobruk government in control of the eastern portion of Libya.
-However, the core contested issue, legitimized control of the central government, remained unresolved with
+In the aftermath of the overthrow of Muammar Gaddafi in 2011, central political authority fragmented and eventually within the midst of a renewed civil war following the contested June 2014 parliamentary elections two competing governments emerged: the internationally recognized Libyan House of Representatives (HoR) elected in 2014 based in Tubruk and in control of much of the Eastern portion of the country with the support of the Libyan National Army under Khalifa Haftar, and the Tripoli-based General National Congress (GNC) originally elected in 2012 and unilaterally self-reinstated in 2014 [@noauthor_libya_nodate-1].
+A ceasefire was signed in 2015 followed by the UN brokered Libyan Political Agreement in December that established the Government of National Accord (GNA) with the intention of uniting both factions [@noauthor_libyan_2016].
+The agreement was short-lived, by August the HoR had voted to reject the GNA, and cooperation rapidly deteriorated [@noauthor_libya_nodate].
+Dyadic hostilities, however, remained at a low-level with only intermittent escalations while focus shifted towards the various Islamist insurgent groups also active in the conflict.
+Meanwhile, the core contested issue, legitimized control of the central government, remained unresolved with a strong potential for a renewal in high-intensity fighting, which occurred in 2019 with Khalifa Haftar's offensive against Tripoli and the UN-backed GNA [@hibrawi_can_2019].
 
-dyadic frozen conflict
+I contend that the intervening years between the initial ceasefire agreement and the 2019 offensive are best characterized as a dyadic frozen conflict between non-separatist entities within a larger civil war.
+Although significantly shorter than the typical examples of frozen conflicts involving *de facto* independent separatist states, it satisfies the criteria of an initial period of violence followed by a cessation of large-scale hostilities, albeit with intermittent fighting, and the core issue clearly remained highly salient whilst central government authority was seriously contested.
 
-[^libya_ceasefire]: https://www.reuters.com/article/us-libya-security-idUSKBN0KP0VL20150116/
+Finally, it is worth distinguishing frozen conflict from several related concepts, specifically protracted stalemates and enduring internal rivalries.
+In the case of the former, stalemates during a conflict can potentially endure for extended periods of time and offer conditions similar to those of a frozen conflict.
+However, they do not necessitate a cessation of active fighting.
+Instead, they may develop into wars of attrition where actors are unable to achieve an outright military advantage, but continue fighting in the hope of outlasting and "grinding down" the enemy, *i.e.* being better able to endure the prolonged costs of conflict.
+In the case of the latter, enduring internal rivalries are defined by @derouen_enduring_2008 as "...internal conflicts between a government and an insurgency with at least 10 years of [non-consecutive] armed conflict in which there are at least 25 deaths", an extension of the interstate concept of enduring rivalries [@goertz_enduring_1993] characterized by repeated onsets of militarized disputes.
+Compared to frozen conflicts, enduring internal rivalries include an explicit temporal requirement and, similar to stalemates, are not restricted to conflicts where fighting has largely stopped.
+The concept of frozen concepts uniquely captures the post-war condition of unstable peace that characterizes certain conflicts where there is no conclusive termination to the initial episode of fighting.
 

### 2024-03-07 paper.qmd
Draft edits

@@
 In the case of the former, stalemates during a conflict can potentially endure for extended periods of time and offer conditions similar to those of a frozen conflict.
-However, they do not necessitate a cessation of active fighting.
+However, they do not necessitate stop active fighting.
 Instead, they may develop into wars of attrition where actors are unable to achieve an outright military advantage, but continue fighting in the hope of outlasting and "grinding down" the enemy, *i.e.* being better able to endure the prolonged costs of conflict.
@@
 
-## How do frozen conflicts emerge?
+## The Emergence of Frozen Conflicts
 
-Frozen conflict as an equilibrium of a situation where both actors strictly prefer a military victory but are unable to achieve a significant military advantage and are unwilling to acquiesce to a peace agreement.
-The belligerents will seek to minimize costs by de-escalating the conflict intensity and the conflict will eventually freeze.
+I view the emergence of frozen conflicts as an equilibrium state of a situation where actors strictly prefer a military victory, but are unable to achieve a significant military advantage and are unwilling to acquiesce to a comprehensive peace agreement.
+Faced with a stalemate, belligerents will seek to minimize costs by de-escalating the conflict intensity whilst maintaining a minimum degree of hostility.
+This may be done informally or formally --- the latter entailing ceasefires or partial peace agreements --- resulting in a cessation of large-scale active fighting following which the conflict will freeze.
+
+To achieve this equilibrium state requires firstly a convergence in beliefs over the utility of additional fighting.
+Actors must believe that there is a high probability that additional fighting will simply maintain the status quo, which will decrease the incentive towards continued violence.[^convergence] From the previous definition, frozen conflict is a *post-conflict* condition that follows after the termination of the initial violent conflict.
+Thus, a stalemate can be considered a minimum necessary condition for the emergence of a frozen conflict if it incentivizes a conflict termination.[^stalemates]
+
+[^stalemates]: Notably, I do not define stalemates strictly in terms of power parity since there are a number of reasons for which an actor may be unable to achieve a military advantage during the course of a conflict.
+This can, of course, include a balance in the material distribution of power, but also constraints arising from, for example, domestic audience costs and political interests of external, third party actors.
+
+[^convergence]: This is akin to the "Principle of Converge" laid forth by @slantchev_principle_2003 whereby the usefulness of fighting decreases when it ceases to convey information on an opponent's capabilities and resolve
 
 Assumes that the expected utility of maintaining the dispute is greater than the utility of a peace agreement.
-
--  transitory war gains outweigh peaceful division of contested good
-- time horizon, low discount factor (Stall war paper assumes that actors actually discount the future?). Anticipation of a future favourable shift in the balance of power
-- credible commitment issues undermine payoff from a peace agreement
-
-Unlike conflict onsets, assume that fighting has revealed information about actors' resolve and capabilities.
 

### 2024-03-08 paper.qmd
First draft

@@
     motivated by an application of bargaining theory to the domain of
-    frozen conflicts. Results indicate that directed support to rebels in
-    the absence of similar support to the government side increases the
-    probability that a civil conflict episode terminates as a frozen
-    conflict. All other treatment conditions --- specifically, directed
+    frozen conflicts. Results indicate that directed support to rebels increases
+    the probability that a civil conflict episode terminates as a frozen
+    conflict within the context of territorial civil conflicts. All other
+    treatment conditions --- specifically, directed
     support provided only to the government and directed supported given
@@
 Frozen conflicts are among some of the most pernicious, difficult to resolve conflicts.
-They are geographically and temporally dispersed, and although they are characterized by a cessation of large-scale hostilities, the underlying disputes remain unresolved.
-This puts actors into a state of uncertainty with an ever present threat of violent re-escalations and can lead to long-running accumulating costs.
+They are unique among conflict outcomes in that they emerge from an inconclusive conflict termination, and leave conflict actors in an indeterminate state of unstable peace with an ever present threat of violent re-escalation.
 
-Surprisingly, little scholarly attention has been paid to the factors influencing the emergence of frozen conflicts [@klosek_frozen_2021].
+Surprisingly, despite the geographic and temporal dispersion of frozen conflicts, little scholarly attention has been paid to the factors influencing their emergence [@klosek_frozen_2021].
+The frozen conflict specific literature mostly consists of qualitative case studies of specific conflicts, while the broader academic work on conflict termination has so far not incorporated frozen conflict as a unique conflict outcome.
 
-This paper seeks to address this gap by asking why do some conflicts become frozen?
+This paper seeks to address this gap by asking, why do some civil conflicts become frozen?
+There are a wide variety of potential factors which could plausibly be correlated with the emergence of frozen conflicts; however, I decide to focus on the international dimension, specifically the role of external support provided by third party state actors to conflict belligerents.
+There exists extensive work on the impact of external support on conflict onset, duration, and termination [@hartzell_civil_2016].
+Although there may be disagreements on the exact relationship between external support and various aspects of civil conflicts, the aggregate finding is that external support is an important factor.
 
-Surprisingly, little scholarly attention has been
-paid to the factors influencing their emergence. This paper addresses
+In what follows, I begin by providing an overview of the previous literature, both in terms of the frozen conflict specific work as well as the broader literature on civil conflict termination.
+I then offer a theory for why certain civil conflicts become frozen and link it to the role played by external support.
+In brief, conflicts freeze when actors seek to minimize the costs of fighting by de-escalating the conflict intensity when unable to either achieve a military advantage during fighting or agree to a negotiated settlement.
+External support is thought to be positively correlated with the likelihood that a civil conflict becomes frozen, primarily by shifting the balance of power under certain conditions towards a stalemate situation.
 
-- relatively little academic literature on why some conflicts
-
-"Why do some conflicts become frozen?"
-
-- focus on the class of intra-state frozen conflicts
-
-- wide variety of potential factors
-  -  I decide to focus on international dimensions, specifically external support
-  - Justify by saying there exists a broad literatur eon the importance of external support on ocnlfict outcomes
-
-- Skeleton for remaining article
-  - First, I provide a brief overview of prev. literature
-  - I then offer a theory of why external support may be linked to the onset of frozen civil conflicts. "In brief, etc etc"
-  - Provide empirical evidence with original coding showing that support is positively linked with prob. of frozen conflict
-    - H/o: only within the context of territorial civil conflicts
-
-- Results contribute to our understanding of frozen conflitcs and conflict outcomes more broadly
-
-frozen conflict - "a protracted state of affairs in which major hostilities have ended, but the enmity between the actors persists and re-escalation remains a permanent threat."
-
-When conflicts freeze, large-scale hostilities cease, but the underlying issue(s) remain unresolved and there remains an ever-present threat of violent re-escalation.
+To empirically investigate this theory I provide an exploratory analysis using a series of bayesian hierarchical probit models with original coded data on frozen civil conflict onsets.
+The findings show that external state support does increase the probability that a civil conflict terminates in a frozen state, but only in the case of directed support to rebel groups in civil conflicts where the central incompatibility centers around territory.
+These results contribute to our understanding of frozen conflicts, as well as provide a more nuanced understanding of conflict termination more broadly.
 
 # Literature Review
+
+Previous empirical research into frozen conflicts is fairly limited.
+The work that does exist largely consists of qualitative, single-case studies [*e.g.*, @blank_russia_2008; @welt_thawing_2010; @pokalova_conflict_2015; @tudoroiu_unfreezing_2016].
+While this literature provides richly detailed and illuminating analyses of specific frozen conflicts, it does not easily provide a comparative perspective to investigate commonalities and differences across cases.
+Thus, we know relatively little about the general correlates of frozen civil conflict emergence.
 
@@
 Third party states intervening or providing material support to opposing sides in a civil conflict for the purpose of balancing may increase conflict duration [@aydin_networks_2012; @anderson_competitive_2019].
+When looking at disaggregated types of external interventions, some have argued that military aid and other forms of support are the key to either a rebel victory or negotiated settlement, as opposed to direct military involvement [@jones_altering_2017].
+Others argue that fungible support, *i.e.* financing and weapons, has an opposite impact by actually extending the duration of conflicts [@sawyer_role_2017].
 
-disaggregate the types of external interventions that third party states may provide to conflict actors.
-This includes looking beyond military interventions with some scholars arguing that military aid and other forms of support, excluding direct military involvement, increase the likelihood of either a rebel victory or negotiated settlement [@jones_altering_2017].
-Others argue that fungible support, *i.e.* financing and weapons, has an opposite impact by actually extending the duration of conflicts [@sawyer_role_2017].
+Taken together, this literature points to the importance of external support as a factor on both conflict duration and termination outcomes.
+Most studies tend to find positive support for the supposition that external support primarily benefits the rebel side, both in terms of increasing the chances of a military victory or a negotiated settlement.
+However, this may not hold when examining more fine grained measures for specific types of support offered to belligerents.
+Nonetheless, the common thread across these works is a general assumption that conflicts terminate either with a one-sided military victory or a negotiated settlement.
+Frozen conflicts, however, represent a qualitatively different type of conflict termination.
+Unlike a one-sided military victory, there is no clear victor whena conflict freezes, rather the conflict stales with both sides either unwilling or unable to continue large-scale hostilities, And, compared to a negotiated settlement, they crucially do not encompass comprehensive peace agreements which would otherwise conclusively terminate a conflict.
+Thus frozen conflicts represent a unique outcome; one that has previously not been examined quantitatively, especially in regards to international dynamics such as external state support.
 
@@
 In the case of the former, stalemates during a conflict can potentially endure for extended periods of time and offer conditions similar to those of a frozen conflict.
-However, they do not necessitate stop active fighting.
+However, they do not necessitate a stop in active fighting.
 Instead, they may develop into wars of attrition where actors are unable to achieve an outright military advantage, but continue fighting in the hope of outlasting and "grinding down" the enemy, *i.e.* being better able to endure the prolonged costs of conflict.
@@
 I view the emergence of frozen conflicts as an equilibrium state of a situation where actors strictly prefer a military victory, but are unable to achieve a significant military advantage and are unwilling to acquiesce to a comprehensive peace agreement.
-Faced with a stalemate, belligerents will seek to minimize costs by de-escalating the conflict intensity whilst maintaining a minimum degree of hostility.
+Faced with a stalemate, belligerents will seek to minimize costs by de-escalating conflict intensity whilst maintaining a minimum degree of hostility.
 This may be done informally or formally --- the latter entailing ceasefires or partial peace agreements --- resulting in a cessation of large-scale active fighting following which the conflict will freeze.
@@
 
-[^convergence]: This is akin to the "Principle of Converge" laid forth by @slantchev_principle_2003 whereby the usefulness of fighting decreases when it ceases to convey information on an opponent's capabilities and resolve
+[^convergence]: This is akin to the "Principle of Converge" laid forth by @slantchev_principle_2003 whereby the usefulness of fighting decreases when it ceases to convey information on an opponent's capabilities and resolve.
+However, rather than emphasising the opportunity for bargaining as @slantchev_principle_2003, I instead link the convergence of beliefs to also the possibility for a conflict to become frozen.
 
-Assumes that the expected utility of maintaining the dispute is greater than the utility of a peace agreement.
+Secondly, there is an implicit assumption that conflict costs are dynamic over time.
+Rather than conceiving of the costs of fighting as a static quantity that accumulates with each time period, I view costs as a function of the strategic choices taken by actors that influence conflict intensity.
+It is often assumed that one of the primary reasons for conflict termination is when the costs of fighting exceeds the expected utility of victory [@t_david_mason_how_1996].[^lit] When viewed in dynamic terms, actors have varying degrees of agency in setting costs such that a violent dispute can be maintained at a minimum level for extended periods of time.
+In other words, even if the expected payoff from maintaining a violent dispute is negative, actors may still avoid a settlement conditional on there being a positive future value associated with winning the dispute and no one belligerent has such a dominant military position during the conflict to effectively set the costs imposed on all other actors.
+
+[^lit]: This assumption is often carried on in more recent literature as well. See for example the overviews provided by @reiter_exploring_2003 and @walter_bargaining_2009.
+
+Finally, for a frozen conflict to emerge as a equilibrium state the underlying issue has to remain salient.
+This was previously discussed in the definition for frozen conflict, but it is worth reiterating in order to emphasize that a stalemate situation is not sufficient to guarantee that a frozen conflict will emerge.
+For example, in the case where a conflict terminates in an indeterminate state, but one of the belligerents unilaterally renounces violence then the underlying incompatibility as an impetus for armed conflict would no longer be pertinent.
+Such a scenario could hardly be described as an ongoing conflict that has frozen.
+
+Taken together, convergent beliefs, dynamic costs, and issue salience represent minimum requirements for a conflict to become frozen, but there is still the open question, why would an actor in a stalemate situation not prefer a bargained solution terminating the conflict?[^mirror] Actors may be able to minimize costs by de-escalating the intensity; however, because the conflict is not conclusively terminated there is an ever present threat of violent re-escalation meaning that actors need to maintain some degree of military preparedness, lest the balance of power swings too far in favor of the opposing side.
+More formally, the expected utility of maintaining the conflict in a frozen state, with the potential to violently relitigate the issue at some future time, has to outweigh the costs associated with a frozen conflict.
+This means that an actor either expects to be able to gain the military advantage they were unable to achieve in the first round of fighting, or believes that they can achieve a more favourable bargain than what is currently available.
+To this end, I propose three different mechanisms that can plausibly incentivize actors to forgo a peaceful settlement under the condition of being unable to achieve a one-sided military victory.
+
+[^mirror]: This mirrors the foundational question that is often the starting point for exploring the causes for the onset of war, "why fight?" If fighting is costly, then there should always exist an *ex ante* bargain dividing the contested good without incurring the cost of conflict [@fearon_rationalist_1995].
+A theoretical mechanism explaining the onset of war needs to be able to account for bargaining failure.
+In the context of frozen conflicts, although fighting has already occurred, a complete theory needs to similarly account for why actors prefer not to seek a bargained solution with a conclusive termination instead of perpetuating the dispute.
+
+#### Credible Commitments
+
+That certain actors are at times unable to credibly commit to a negotiated settlement has long been identified as a potential hindrance to civil conflict termination [@walter_critical_1997; @sawyer_role_2017].
+If an actor believes that an opponent will eventually renege on an agreement, they may be incentivized to fight even if they prefer to avoid war.
+A similar mechanism can be applied to frozen conflicts.
+Distrust and uncertainty can make it difficult to establish a bargaining space between actors.
+If neither actor is willing to continue fighting, then it may be preferable to let the conflict freeze if the costs of a frozen conflict are outweighed by the risk of the opponent reneging.
+
+#### Transitory War Gains
+
+A second important mechanism relates to the wartime distribution of disputed goods.
+During the course of a conflict, actors achieve transitory gains, such as for example territory or access to specific resources, which may not carry over into the post-conflict order under a bargained solution.
+If those transitory gains outweigh the benefits of a settlement, an actor who discounts the future will seek to maintain the *intrabellum* distribution and forestall any negotiations [@spaniel_stall_2020].
+This is particularly applicable to frozen conflicts since by definition the belligerents are unable to agree on a bargained solution to the conflict.
+When the conflict freezes this also freezes the current status quo such that actors are able to continue deriving utility from their transitory war gains.
+Therefore, a frozen conflict may arise when there is a mismatch between the wartime distribution and the expected distribution under a post-conflict settlement.
+
+#### Time Horizons
+
+Lastly, actors may also prefer the utility gained from a frozen conflict over a negotiated settlement due to long time horizons.
+The mechanism for transitory war gains posited that actors that discount the future and value the present distribution of goods may be in favour of a frozen conflict.
+Although seemingly contradictory, the opposite relationship may also be present whereby an actor values the future under the expectation that they will eventually achieve a military advantage, while discounting the present costs associated with a frozen conflict.
+More specifically, an actor may allow a conflict to freeze to bide their time rebuilding their forces and re-strategizing in anticipation of a future violent showdown.
+Thus, a frozen conflict becomes a waiting device for actors unable to achieve an immediate military victory, but hoping the there will be a future favourable shift in the balance of power.[^ceasefires]
+
+[^ceasefires]: This argument mirrors some of the work on the role of ceasefires in bargaining [@clayton_logic_2021; @sticher_bargaining_2021; @sticher_healing_2022].
+Actors may accept a ceasefires according to a variety of logics, for example, they may see it as the first step towards a peace agreement, or an opportunity to gather their strength before pursuing once again a military victory.
+
+### Distinguishing Between Civil Conflict Types
+
+All three of these mechanisms may be present in different conflicts that have become frozen; their logics are not necessarily mutually exclusive and in some ways can be complementary.
+The degree to which each mechanism is relevant to any specific conflict depends in part on the conflict type.
+It is useful here to distinguish between civil conflicts fought over control of the state and civil conflicts fought over autonomy and/or independence claims.
+In the case of the former, the ultimate goal is to seize power from an incumbent regime.
+The prime impediment to a settlement is the risk that the post-conflict incumbent will violate the agreement, especially within authoritarian contexts where there is a fear of retribution.
+Thus, the preference for a frozen conflict, assuming that neither side can achieve a military advantage, may primarily stem from credible commitment problems.
+
+In the case of the latter, civil conflicts fought over autonomy and/or independence claims, can also involve credible commitment issues as an incentive towards frozen conflict.
+Specifically, it is difficult for the government to credible commit to an agreement granting greater autonomy, in part because it may galvanize additional claims from other groups [@walter_information_2006; @toft_self-determination_2012].
+Furthermore, as some scholars have argued, the international community tends to be fairly conservative regarding new sovereignty claims [@huddleston_foulweather_2021; @siroky_unilateral_2021].
+Since rebel groups are typically outmatched by the government [@brandt_when_2008] such that it is unlikely that they will win in an absolute war and successfully disarm the regime, a frozen conflict may be the optimal outcome, especially when the opposition is able to seize and consolidate territory.
+If the rebel group is unable to trust the government to credibly bargain, then they will highly value the present transitory gains and seek to maintain the status quo.
+Meanwhile, the government may also face an incentive towards a frozen conflict in these scenarios.
+There is a strong international norm in favor territorial integrity.[^IR] If the government is unable to outright defeat a rebel group, especially one that has established a *de facto* independent state, then the government may be seeking a long-term plan for forcible reintegration.
+A frozen conflict allows the government to ease the burden of active fighting while maintaining their sovereignty claims on breakaway territory.
+Crucially, this latter point allows the government to frame the frozen conflict as an ongoing civil war if and when there is another violent re-escalation, as opposed to an interstate war of territorial conquest against a newly independent state.
+
+[^IR]: I need to check the IR literature to provide a citation for this claim.
+
+### External State Support
+
+Where does external support by third party states to belligerents fit into the previously mentioned mechanisms and how does external support shape the utility calculation of actors either in favor or against a conflict freezing?
+I see external state support affecting the propensity for a frozen conflict in several different ways.
+Principally, support to a conflict actor can increase their military capabilities and shift the balance of power.
+As some scholars have argued, this is especially true in the case of directed support to rebel groups [@balch-lindsay_third-party_2008; @gent_going_2008], providing them an opportunity to launch a credible challenge to the government.
+This does not mean that the conflict will automatically become frozen, but it may raise the likelihood of a stalemate which is an important first step.
+Furthermore, external support also helps subsidize the potential cost of maintaining a frozen conflict, which may allow actors a longer time horizon such that they will reject an immediate settlement if they are optimistic about the future utility of renewed fighting.
+
+Scholars have also argued that the involvement of additional third party actors can complicate the bargaining process for an ongoing civil conflict [@cunningham_veto_2006; @cunningham_blocking_2010].
+External supporters may have conflicting outcome preferences to the actors they are supporting.
+As the number of actors involved in a bargaining process grows, coordinating the disparate interests becomes increasingly difficult, another disincentive against achieving a negotiated settlement.
+
+In sum, external support should increase the likelihood of a stalemate and incentivize actors to maintain the transitory status quo in the form of a frozen conflict.
+My empirical expectation is that the effect should be strongest when directed support is only provided to the rebels.
+When support is either provided to the government, or to both sides, then it is less likely that a balance of power will emerge and I assume the situation will on average favor the government.
+Furthermore, I also expect this correlation to be strongest for civil conflicts concerning independence or autonomy.
+Given how difficult it is to secure autonomy rights from a government or achieve international sovereignty recognition for separatist rebels, as well as the general desire for governments to forcibly reintegrate territories where they maintain a sovereignty claim, there exists a stronger incentive from both sides to favour a frozen outcome.
+External support is especially important then to provide the capacity to rebel groups to secure territory and consolidate power.
 
@@
 To empirically assess whether external state support for conflict actors affects the probability of a civil conflict episode terminating as a frozen conflict a series of bayesian probit models were run using original data on frozen civil conflicts.
-A mixed identification strategy was employed by both conditioning on relevant covariates and instrumenting external support with UN roll call votes.
-
-In what follows, I will first present the model specification and results from a probit regression predicting the onset of a frozen conflict.
-I will then decrease the dependent and independent variables, including operationalizations and data sources.
+In what follows, I will first present the model specification followed by a discussion of the included dependent and independent variables, including operationalizations and data sources.
+Finally, I will present the results and discuss the implications for understanding the emergence of frozen conflicts.
 
@@
 For $\delta_k$ and $\gamma$ I use hierarchical priors that allow for varying effects while partially pooling information across groups to prevent overfitting.
-Regarding the latter, this means $\delta_{jk} \sim \text{N}(\mu_j, \tau_j)$ for $j = 1, 2, \ldots, D$ where $\mu_j \sim \text{N}(0, 1)$ and $\tau_j \sim \text{N}(0, 2.5)$.
-And for the former, $\gamma_c \sim \text{N}(0, \sigma)$ and $\sigma \sim \text{N}(0, 1)$.
+Regarding the former, this means $\delta_{jk} \sim \text{N}(\mu_j, \tau_j)$ for $j = 1, 2, \ldots, D$ where $\mu_j \sim \text{N}(0, 1)$ and $\tau_j \sim \text{N}(0, 2.5)$.
+And for the latter, $\gamma_c \sim \text{N}(0, \sigma)$ and $\sigma \sim \text{N}(0, 1)$.
 The particularly restrictive hyperprior on $\sigma$ shrinks the posterior $\gamma$ estimates towards zero and encodes a significant degree of skepticism towards country-specific deviations.[^country-intercepts]
@@
 
-[^mem]: An alternative approach would be to calculate the differences in the predicted probability for each treatment effect while holding the remaining covariates constant at their mean and/or mode.
+[^mem]: An alternative approach would be to calculate the differences in the predicted probability for each treatment effect while holding the remaining covariates constant at their mean or mode.
 However, this would not necessarily be representative of the "typical case" in the population [@hanmer_behind_2013].
-Although there is a minimal amount of non-stochastic missingness in the data I employ due to differing coding coverage between datasets resulting in dropped conflict episodes, I would argue that the sample averaged marginal effects are still more representative of each treatment would affect the typical civil conflict.
+Although there is a minimal amount of non-stochastic missingness in the data I employ due to differing coding coverage between datasets resulting in dropped conflict episodes, I would argue that the sample averaged marginal effects are still more representative of how each treatment would affect the typical civil conflict.
 
@@
 
-The measurement of frozen conflicts provided herewithin is crude. Future work should focus on refining this measurement.
+This paper sought to investigate a potential link between external support offered by third party states and the likelihood that a civil conflict transitions into a frozen conflict.
+A theory was offered emphasizing the expected utility of actors
+and how external support can shape the cost-benefit calculation in favour of a frozen outcome.
+This was then investigated empirically with a bayesian hierarchical probit model showing correlationary evidence that directed support to rebel groups does increase the probability that a civil conflict freezes, but only within the context of territorial civil conflicts.
 
-Potential shortcoming: what I am not doing is taking into account the preferences of external supporters and motivation for offering support.
-Nor am I distinguishing between different external support types.
+One potential shortcoming is that I did not take into account the preferences of external supporters and motivation for offering support.
+Nor did I disaggregate between different types of external support.
+In line with other empirical work on the disaggregated effects of external support and conflict duration, there may be a differential effect depending on the type and amount of support offered to conflict belligerents.
+
+Furthermore, the measurement offered for frozen conflicts was necessarily crude.
+It operationalized frozen conflict as a discrete event, "frozeness" maybe is best conceptualized as lying on a continuous dimension that allows conflicts to fluctuate  between "more" and "less" frozen.
+Future work should focus on refining this measure and to hopefully further expand on our understanding of both frozen conflicts and civil conflict termination.
 

### 2024-03-08 lib/fc.utils/tests/testthat.R
Bugfix unit tests

@@ -7,6 +7,6 @@
 # * https://testthat.r-lib.org/articles/special-files.html
 
 library(testthat)
-library(fc)
+library(fc.utils)
 
-test_check("fc")
+test_check("fc.utils")

### 2024-03-09 R/probit.R
Fix cmdstan_version for precompiled models
This is solely to make the generated project image smaller by only using
the Cmdstan image to pre-compile the Stan models. Once we generate our
model binaries we no longer need Cmdstan; however, `cmdstanr` still
requires a valid path to do version checks so override with a hack.

@@ -7,6 +7,10 @@ library(tools)
 
 options(mc.cores = 1)
 
+# This is only for building the project image in order to avoid
+# including CmdStan by precompiling our model
+assignInNamespace("cmdstan_version", function(...) "2.34.1", ns = "cmdstanr")
+
 design <- readRDS("./data/model_inputs/design.rds")
 
 input <- commandArgs(trailingOnly = T)
@@ -37,7 +41,7 @@ stan_data <- list(N = sum(cases),
                   y = y[cases])
 str(stan_data)
 
-mod <- cmdstan_model("./stan/hierarchical_probit.stan")
+mod <- cmdstan_model(exe_file = "stan/hierarchical_probit")
 fit <- mod$sample(data = stan_data, refresh = 0, sig_figs = 3, adapt_delta = 0.99)
 
 # Treatment coefficients

### 2024-03-09 R/sbc.R
Add Containerfile

@@ -10,6 +10,10 @@ library(parallel)
 
 options(mc.cores = parallel::detectCores() / 2)
 
+# This is only for building the project image in order to avoid
+# including CmdStan by precompiling our model
+assignInNamespace("cmdstan_version", function(...) "2.34.1", ns = "cmdstanr")
+
 rank_statistic <- function(draws, true_value) {
     sum(draws < true_value)
 }

### 2024-03-11 R/sbc.R
BUGFIX: use precompiled sbc model

@@ -47,7 +47,7 @@ data <- list(N = n,
              n_contest_types = n_conflict_types,
              contest_id = conflict_id)
 
-sim <- cmdstan_model("./stan/sim.stan")
+sim <- cmdstan_model(exe_file = "stan/sim")
 sim_data <- sim$sample(data = data, fixed_param = TRUE, chains = 1)
 
 # Simulated parameter values

### 2024-03-11 R/sbc.R
BUGFIX: use precompiled hierarchical_probit model for sbc

@@ -8,7 +8,7 @@ library(dplyr)
 library(MASS, include.only = "mvrnorm")
 library(parallel)
 
-options(mc.cores = parallel::detectCores() / 2)
+options(mc.cores = parallel::detectCores())
 
 # This is only for building the project image in order to avoid
 # including CmdStan by precompiling our model
@@ -60,14 +60,14 @@ y_sim <- sim_data$draws("y_sim", format = "matrix")
 # For each simulated outcome, run the model and compute the rank
 # statistic for each parameter between the posterior draws and the
 # true (ie simulated) value
-mod <- cmdstan_model("./stan/hierarchical_probit.stan")
+mod <- cmdstan_model(exe_file = "./stan/hierarchical_probit")
 ranks <- mclapply(1:nrow(y_sim), function(i) {
     stan_data <- data
     stan_data$y <- as.vector(y_sim[i, ])
     stan_data$interaction_id <- 1
 
-    fit <- mod$sample(data = stan_data, sig_figs = 3, iter_sampling = 500, chains = 2,
-                      parallel_chains = 2, adapt_delta = 0.95, thin = 2, refresh = 0)
+    fit <- mod$sample(data = stan_data, sig_figs = 3, iter_sampling = 500, chains = 1,
+                      adapt_delta = 0.95, thin = 2, refresh = 0)
     mapply(rank_statistic, fit$draws(parameters, format = "data.frame"), pv[i, ])
 })
 

### 2024-03-11 paper.qmd
Add env variable to skip sbc

@@
 
-```{R}
+```{R eval = Sys.getenv("NO_SBC") == "" && file.exists("posteriors/sbc.rds")}
 #| label: fig-sbc
@@
 #|     Simulation based calibration results
-
 sbc <- readRDS("./posteriors/sbc.rds") |>

### 2024-03-11 scripts/wordcount.sh
Switch to bash for wordcount script
I hate bash, but it's already pre-installed on debian.

@@ -1,19 +1,22 @@
-#!/usr/bin/env zsh
+#!/usr/bin/env bash
 #
 # Approximate word count for a quarto/Rmarkdown file. Includes body
 # text count + references, abstract, and figure captions while
 # excluding appendix and code blocks.
 ###
 
-setopt err_exit
+set -eo pipefail
 
 if [[ $# -ne 1 ]]; then
-    print -u 2 "Missing file argument"
-    print -u 2 "Usage: $0 <file.[qmd|Rmd|md]>"
+    printf "Missing file argument\n" >&2
+    printf "Usage: $0 <file.[qmd|Rmd|md]>\n" >&2
     exit 1
 fi
 
-[[ ! -f $1 ]] && { print -u 2 "Invalid file: $1"; exit 1 }
+if [[ ! -f "$1" ]]; then
+    printf "Invalid file: $1\n" >&2
+    exit 1
+fi
 
 abstract=$(yq -f extract '.abstract' $1 | wc -w)
 
@@ -25,4 +28,4 @@ body=$(sed -e '/^```/,/^```/d' -e '/^#\s*Appendix/,$d' $1 | \
 # blocks
 figure_captions=$(grep '^#|' $1 | sed -e 's/^#|\s*//' -e '/^.*:/d' | wc -w)
 
-print "$((abstract + body + figure_captions))"
+printf "%d\n" "$((abstract + body + figure_captions))"

### 2024-05-09 R/sbc.R
Increase number of iterations for sbc

@@ -5,27 +5,22 @@
 
 library(cmdstanr)
 library(dplyr)
+library(fc.utils)
 library(MASS, include.only = "mvrnorm")
 library(parallel)
 
 options(mc.cores = parallel::detectCores())
 
-# This is only for building the project image in order to avoid
-# including CmdStan by precompiling our model
-assignInNamespace("cmdstan_version", function(...) "2.34.1", ns = "cmdstanr")
-
-rank_statistic <- function(draws, true_value) {
-    sum(draws < true_value)
-}
-
+iter <- 5000
 n <- 248
 m <- 8
 k <- 3
 
+info("Running simulation based calibration for %d iterations and %d threads",
+     iter, parallel::detectCores())
+
 parameters <- c("alpha", "beta", "gamma", "sigma", "delta", "mu", "tau")
 
-#X <- mvrnorm(n, rep(0, m), diag(rep(2, m), nrow = m, ncol = m)) |>
-#    apply(2, scale)
 X <- sample(0:1, n * m, replace = T) |> matrix(nrow = n, ncol = m)
 
 n_countries <- 3
@@ -48,7 +43,7 @@ data <- list(N = n,
              contest_id = conflict_id)
 
 sim <- cmdstan_model(exe_file = "stan/sim")
-sim_data <- sim$sample(data = data, fixed_param = TRUE, chains = 1)
+sim_data <- sim$sample(data = data, fixed_param = TRUE, chains = 1, iter_sampling = iter)
 
 # Simulated parameter values
 pv <- sim_data$draws(parameters, format = "data.frame")
@@ -60,7 +55,7 @@ y_sim <- sim_data$draws("y_sim", format = "matrix")
 # For each simulated outcome, run the model and compute the rank
 # statistic for each parameter between the posterior draws and the
 # true (ie simulated) value
-mod <- cmdstan_model(exe_file = "./stan/hierarchical_probit")
+mod <- cmdstan_model(exe_file = "stan/hierarchical_probit")
 ranks <- mclapply(1:nrow(y_sim), function(i) {
     stan_data <- data
     stan_data$y <- as.vector(y_sim[i, ])
@@ -72,4 +67,6 @@ ranks <- mclapply(1:nrow(y_sim), function(i) {
 })
 
 df <- bind_rows(ranks)
+
+dir.create("posteriors", showWarnings = F)
 saveRDS(df, "posteriors/sbc.rds")

### 2024-05-09 lib/fc.utils/R/utils.R
Move rank_statistic function to fc.utils

@@ -35,3 +35,8 @@ stanvar_latex <- function(s) {
     else
         sprintf("$\\%s$", s)
 }
+
+#' @export
+rank_statistic <- function(draws, true_value) {
+    sum(draws < true_value)
+}

### 2024-05-09 lib/fc.utils/R/zzz.R
Try to override cmdstan_version in fc.utils on load

@@ -1,0 +1,5 @@
+.onLoad <- function(libname, pkgname) {
+    # This is only for building the project image in order to avoid
+    # including CmdStan by precompiling our model
+    assignInNamespace("cmdstan_version", function(...) "2.34.1", ns = "cmdstanr")
+}

### 2024-05-09 lib/fc.utils/R/zzz.R
BUGFIX: find assignInNamespace function

@@ -1,5 +1,6 @@
 .onLoad <- function(libname, pkgname) {
     # This is only for building the project image in order to avoid
-    # including CmdStan by precompiling our model
-    assignInNamespace("cmdstan_version", function(...) "2.34.1", ns = "cmdstanr")
+    # including CmdStan by precompiling our model. Don't every do
+    # something like this in a real package, please.
+    utils::assignInNamespace("cmdstan_version", function(...) "2.34.1", ns = "cmdstanr")
 }

### 2024-05-09 R/sbc.R
Reject sbc model fits failing diagnostics

@@ -9,7 +9,7 @@ library(fc.utils)
 library(MASS, include.only = "mvrnorm")
 library(parallel)
 
-options(mc.cores = parallel::detectCores())
+options(mc.cores = parallel::detectCores() / 2)
 
 iter <- 5000
 n <- 248
@@ -61,8 +61,15 @@ ranks <- mclapply(1:nrow(y_sim), function(i) {
     stan_data$y <- as.vector(y_sim[i, ])
     stan_data$interaction_id <- 1
 
-    fit <- mod$sample(data = stan_data, sig_figs = 3, iter_sampling = 500, chains = 1,
-                      adapt_delta = 0.95, thin = 2, refresh = 0)
+    fit <- mod$sample(data = stan_data, chains = 2, iter_warmup = 1000,
+                      iter_sampling = 1000, adapt_delta = 0.95, refresh = 0)
+
+    diagnostics <- fit$diagnostic_summary()
+    if (sum(diagnostics$num_divergent) > 0 | sum(diagnostics$num_max_treedepth) > 0) {
+        warning("Divergent transitions or max treedepth exceeded")
+        return(NULL)
+    }
+
     mapply(rank_statistic, fit$draws(parameters, format = "data.frame"), pv[i, ])
 })
 

### 2024-05-10 paper.qmd
Paper edits

--- 2024-03-08 paper.qmd
@@
 execute:
-    cache: true
+    cache: false
     echo: false
@@
 The outcome variable in my regression model is a binary indicator for whether a civil conflict terminates in a frozen state.
-Following the definition of frozen conflicts offered earlier, I construct this measure using a mix of automated and manual coding of intrastate and internationalized intrastate conflict terminations taken from the UCDP Conflict Termination Dataset [@kreutz_how_2010].[^protocol] Episodes were considered terminated if the total number of battle related fatalities dropped below 25 within a year.
-I subsequently labeled a termination as a frozen conflict if three criteria were satisfied.
+Following the definition of frozen conflicts offered earlier, I construct this measure using a mix of automated and manual coding of intrastate and internationalized intrastate conflict terminations taken from the UCDP Conflict Termination Dataset [@kreutz_how_2010].[^protocol] Episodes were considered terminated if the total number of battle related fatalities dropped below 25 within a year.[^strict_frozen] I subsequently labeled a termination as a frozen conflict if three criteria were satisfied.
 First, an episode could not result in a one-sided military victory.
 This would constitute a conclusive termination allowing the victor to set the terms of the post-conflict order.
-Second, the episode could not conclude with a comprehensive peace agreement that addresses the underlying core issue according to the UCDP Peace Agreement Dataset [@pettersson_organized_2019].[^partial] To account for the possibility of a failed peace agreement I further required that the comprehensive peace agreement had to last for at least one calendar year in order to disqualify an episode termination as a frozen conflict.[^strict_frozen] Finally, an episode could not result in actor death since this is a minimum necessary condition for a conflict to become frozen.
+Second, the episode could not conclude with a comprehensive peace agreement that addresses the underlying core issue according to the UCDP Peace Agreement Dataset [@pettersson_organized_2019].[^partial] To account for the possibility of a failed peace agreement I further required that the comprehensive peace agreement had to last for at least one calendar year in order to disqualify an episode termination as a frozen conflict.
+Finally, an episode could not result in actor death since this is a minimum necessary condition for a conflict to become frozen.
 For a state actor this meant that although there could be a turnover in the regime, the state could not cease to exist as a sovereign entity following the conclusion of the conflict episode.
@@
 
-[^protocol]: For a detailed description of the codcing procedure see the protocol document in the supplemental materials.
+[^protocol]: For a detailed description of the coding procedure see the [coding protocol][Coding Protocol] in the appendix.
 
@@
 
-[^strict_frozen]: In the appendix I extend the analysis presented in this paper and run a series of robustness checks where I increase the number of years required for a compehensive peace agreement to be in place in order to better distingiush between frozen conflicts and short-term pauses in fighting.
+[^strict_frozen]: In the [appendix][robustness checks] I extend the analysis presented in this paper and run a series of robustness checks where I adopt a more strict definition of conflict termination requiring that frozen conflicts have more than 2 years below the UCDP threshold for an active conflict year in order to better distinguish between frozen conflicts and short-term pauses in fighting.
+TODO: what are the results?
 
@@
 In total, `r n_frozen` frozen conflicts were identified out of `r nrow(dataset)` civil conflict terminations, comprising approximately `r sprintf("%.0f", 100 * n_frozen / nrow(dataset))` percent of all observations for the period 1975 to 2017.
-Of the frozen conflicts, `r n_terr_frozen` were conflicts where the central incompatibility between conflict actors concerned territory rather than control of the central government.
-@fig-frozen shows both a world map of the countries that have experienced at least one frozen conflict within the coded time period as well as the incidence rate of frozen conflicts per year.
+Of the frozen conflicts, `r n_terr_frozen` were conflicts where the central incompatibility between conflict actors concerned territory rather than control of the central government.[^list] @fig-frozen shows both a world map of the countries that have experienced at least one frozen conflict within the coded time period as well as the incidence rate of frozen conflicts per year.
+
+[^list]: The full list of all frozen civil conflict onsets is included in @tbl-frozen in the appendix.
 
@@
 It can be argued that the former may persist at such a low level of activity with only sporadic, isolated incidences of limited violence as to not represent a meaningfully salient issue to the incumbent state and would therefore not constitute a frozen conflict.
-To mitigate this risk, I also run the analysis presented in this paper on a sub-sample of conflicts with a cumulative intensity of at least 1,000 battle related fatalities since onset.
+To mitigate this risk, I also run the analysis presented in this paper on a sub-sample of conflicts with a cumulative intensity of at least 1,000 battle related fatalities since onset in the [appendix][robustness checks].
+TODO: results
 
@@
 
-[^robust_censor]: I also run multiple robustness checks in the appendix where I alternatively drop episodes with an episode onset year before the coding start for the dataset and/or recurrent episodes within conflicts that also began prior to 1975.
+[^robust_censor]: I also run multiple robustness checks in the [appendix][robustness checks] where I alternatively drop episodes with an episode onset year before the coding start for the dataset and/or recurrent episodes within conflicts that also began prior to 1975.
 
@@
 It was run with 4 parallel chains at 2,000 iterations each with a warmup of 1,000 with convergence assessed using the standard Stan diagnostics.
+Validation of the implementation code was done through simulation based calibration as detailed in the [appendix][simulation based calibration].
 All pre/post-processing was done using R `r sprintf("%s.%s", R.version$major, R.version$minor)` [@r_core_team_r_2023].
@@
 
+## Coding Protocol
+
+The frozen conflict data was created using a mix of programmatic and manual coding in a two-stage process.
+The first stage filtered out dyadic civil conflict episode outcomes according to a preset list of rules listed in detail below.
+This candidate list was then manually corrected in the second stage to form the final list of frozen conflict onsets.
+
+### Procedure
+
+#### Stage 1
+
+Dyadic conflict episode outcomes were drawn from the UCDP Conflict Termination dataset v21.1.
+This means that the definition of a conflict episode and its corresponding termination followed the coding conventions of the UCDP dataset.
+Namely, dyad-years are coded as active conflicts when the number of battle related fatalities rises above a threshold of 25.
+Consecutive conflict years, or episodes, are terminated the first year when this number drops below the threshold.
+
+Episode outcomes were selected as potential onsets of frozen civil conflicts when satisfying the following predicates:
+
+- Intrastate and internationalized intrastate conflicts --- *i.e.*, when the UCDP variable $\text{type\_of\_conflict}_i \in \{ 3, 4 \}$
+
+- Outcomes were not coded as either a one-sided military victory or actor death --- *i.e.*, when the UCDP variable $\text{outcome}_i \notin \{ 3, 4, 6 \}$
+
+- No comprehensive peace agreement lasting at least one calendar year signed either in the termination year or the subsequent year.
+This corresponds to when the UCDP variable $\text{pa\_type} = 1$ while duration was calculated based on the date of the last signature.
+
+The last predicate relied on data drawn from the UCDP Peace Agreement dataset v22.1.
+Due to the smaller time series in the Peace Agreement dataset, the final list was restricted to conflict outcomes between 1975 and 2019.
+
+```{r echo = F}
+candidates <- read.csv("data/conflict_candidates.csv")
+episodes <- read.csv("data/conflict_episodes.csv")
+```
+
+In total, `r nrow(candidates)` candidates were identified, comprising approximately
+`r sprintf("%.2f", 100 * nrow(candidates) / sum(!is.na(episodes$outcome)))` percent of all outcomes.
+
+#### Stage 2
+
+The resulting list was then manually coded to filter out episode outcomes where the non-state actor ceased to exist as a military actor within five years of termination.
+This includes situations where, for example, the rebel group unilaterally renounces violence as a political strategy and transitions to becoming a political party.
+Or, when a rebel group becomes so weakened and/or fragmented that they effectively dissolve.
+
+Evaluations were primarily done using a mix of UCDP's own description of conflicts and conflict actors, as well as a number of different academic sources.
+
 ## List of Frozen Conflicts
@@
 #| label: tbl-frozen
+#| tbl-cap: Frozen Civil Conflicts (1975 - 2017)
 
@@
 
-```{R}
+The model specification is implemented using the probabilistic programming language, Stan.
+I use simulation-based calibration to validate the correctness of the implementation code according to the following procedure [@gelman_bayesian_2020].
+
+I first fix the covariate matrix $X$ and the treatment matrix $T$ by $X_i, T_i \sim \text{Bern}(p) \; \forall i \in \{1, 2, \ldots, N \}$ where $N$ is set to the number of observations in the actual dataset.
+
+Next, letting $\theta$ be the vector of unknown parameters in the regression equation, I simulate parameter values from the prior, $\tilde{\theta}^{(m)} \sim p(\theta)$, followed by data from the likelihood, $\tilde{y}^{(m)} \sim p(y | \tilde{\theta}, X, T)$ where $m = 1, 2, \ldots, M$.
+
+For each simulated tuple, $(\tilde{\theta}^{(m)}, \tilde{y}^{(m)} )$, I fit the full inference model and draw $L$ values from the posterior, $\{ \theta_1, \theta_2, \ldots, \theta_L \} \sim p(\theta | \tilde{y}^{(m)}, X, T)$.
+
+Finally, I calculate the rank statistic for each parameter $k$, $r_{m, k} = \sum_{l = 1}^L \mathbb{I} \{ \theta_k^{(l)} < \tilde{\theta}_k^{(m)} \}$.
+If the model is correctly specified and the inference procedure is unbiased, then $r_k \sim \text{Unif}(0, L)$ as $M \to \infty$.
+[^proof]
+
+@fig-sbc plots the histograms from the results of this procedure showing that the rank statistics for each parameter is approximately uniform.
+
+[^proof]: For a full proof see Appendix B in @talts_validating_2020.
+
+```{R eval = Sys.getenv("NO_SBC") == "" && file.exists("posteriors/sbc.rds")}
 #| label: fig-sbc
 #| fig-cap: |
-#|     Simulation based calibration results
-
+#|     Simulation based calibration results for $M = 5000$ and $L =
+#|     2000$. Posteriors were not thinned since the effective sample
+#|     sizes showed anticorrelation among the MCMC draws.
 sbc <- readRDS("./posteriors/sbc.rds") |>
@@
 
-Different combinations of the following...
+There are several concerns that may be raised with the main results presented in this paper examining the potential link between directed external support by a third party state to a conflict actor and the probability for a civil conflict to have terminated as a frozen conflict.
+It what follows I detail each potential issue and present additional and/or alternative measures of frozen conflicts, conflict episodes, and the treatment variables.
+Finally, I re-run the model for all the resulting combinations and present the results.
 
-- Aggregation of the treatments: binary treatment, binary treatment in the last 5 years of the conflict episode, and proportional treatment.
-The latter is calculated as the proportion of conflict-years in an episode where one side received external support
+The first potential issue is that the original definition for a conflict termination as taken from the UCDP Conflict Termination Dataset requires that a conflict drops below a threshold of 25 battle related fatalities for only a single year.
+This risks including brief pauses in fighting as potential terminations to be coded as frozen conflicts.
+Although the conceptual definition of frozen conflict that I employ does not include an explicit temporal dimension as a requirement for a conflict to be considered frozen, it could be nonetheless argued that such pauses should not be considered candidates for being classified as frozen since actors may view the conflict as active and ongoing.
+Therefore, I also construct a stricter measure for frozen conflict whereby a conflict episode has to be followed by more than two years below the UCDP battle related threshold to be considered eligible to being classified as frozen, conditional on the remaining the coding criteria.
 
-- All conflict episodes and only high intensity episodes --- try to distinguish between low-level insurgencies and actual civil wars
+The second issue also relates to the underlying population of conflict episodes used to construct the frozen conflict measure.
+In the original coding for frozen conflict, I included all conflict episodes from the UCDP within the time period from 1975 to 2017, including low-level conflicts with relatively few battle related fatalities.
+It can be argued that minor, long-running insurgencies should not be considered as frozen conflicts since the underlying issue may no longer be particularly salient and there is a relatively low risk for a major escalation in violence.
+Such conflicts may experience sporadic, isolated incidences of violence, but would otherwise are minor factors in the domestic political agenda.
+To this end, I subset the list of conflict episodes to only include civil conflict episodes that have cumulatively experienced at least 1,000 battle related fatalities since onset in the hopes that major conflicts maintain a greater degree of salience post-termination.
 
-- Normal definition for frozen conflicts and a strict definition that requires a duration of at least 2 years in order to exclude temporary one-year pauses in fighting that may be more the result of UCDP's coding decision rather than representing actual frozen conflict.
+Third, due to the truncated nature of the dataset, some of the civil conflict episodes are censored with missing data before 1975.
+In the main results I included a binary variable if the onset of an episode occurred before the start of the dataset.
+As an alternative, in the following robustness checks I also drop all censored episodes as well as censored conflicts.
+
+Finally, the primary treatment variables used in the main text are aggregated binary measures for directed support, plus the interaction, at any time during a civil conflict episode.
+Although this aids in the interpretation of the treatment effect, these variables do not capture the extent to which support is provided during a conflict episode.
+Thus, I also replace the treatment variables with measures capturing the proportion of conflict-years within an episode that each actor received external support.
+
+In total, this leads to `r list.files("posteriors", "\\d{4}") |> length()` different model configurations for the original vs strict definition of frozen conflict, inclusion and exclusion of low-intensity conflict episodes, inclusion and exclusion of censored episodes and/or conflicts, and finally the proportional aggregation of the treatment variables.
+@fig-robust
 
@@
 
-ggplot(model_ames, aes(pzero, median)) +
-    geom_point(alpha = 0.2) +
-    geom_errorbar(aes(ymin = q5, ymax = q95), alpha = 0.25) +
-    geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.2) +
-    facet_grid(variable ~ group) +
-    labs(y = parse(text = "tilde(Delta)"), x = parse(text = "P(tilde(Delta) > 0)")) +
-    theme(legend.position = "none")
+rebel <- filter(model_ames, variable == "Ext. Sup. Rebel Only", censored == "with_censored")
+ggplot(rebel, aes(pzero, median, colour = treatment)) +
+    geom_point(aes(shape = outcome), alpha = 0.8) +
+    geom_errorbar(aes(ymin = q5, ymax = q95), alpha = 0.1) +
+    geom_vline(xintercept = 0.9, linetype = "dashed", alpha = 0.2) +
+    facet_grid(episodes ~ group)
+
+#filter(model_ames, variable == "Ext. Sup. Rebel Only") |>
+#ggplot(aes(pzero, median, colour = treatment)) +
+#    geom_point(alpha = 0.2) +
+#    geom_errorbar(aes(ymin = q5, ymax = q95), alpha = 0.25) +
+#    geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.2) +
+#    geom_vline(xintercept = 0.9, linetype = "dashed", alpha = 0.2) +
+#    facet_grid(variable ~ group) +
+#    labs(y = parse(text = "tilde(Delta)"), x = parse(text = "P(tilde(Delta) > 0)")) +
+#    theme(legend.position = "none")
 

### 2024-05-10 R/conflict_candidates.R

@@ -41,6 +41,8 @@ full.df <- filter(pce_collapsed, conflict_id %in% term$conflict_id) |>
 
 info("Finished with %d conflict terminations", sum(!is.na(full.df$outcome)))
 
+write.csv(full.df, "data/conflict_episodes.csv", row.names = F)
+
 ###
 # Filter out candidates for being frozen conflicts
 outcomes <- group_by(full.df, conflict_id) |>

### 2024-05-10 R/models.R

@@ -8,7 +8,7 @@ df <- readRDS("./data/merged_data.rds") |> ungroup()
 
 ###
 # Create a design matrix with all possible analysis combinations
-treatments <- c("bin", "bin_5y", "prop")
+treatments <- c("bin", "prop")
 outcomes <- c("frozen", "strict_frozen")
 units <- c("all", "cumulative_intensity")
 censored <- c("with_censored", "no_censored", "no_episode_censored")

### 2024-05-23 R/merge.R
Add pko control and cubic polynomials for duration

@@ -119,11 +119,24 @@ ctable <- read.csv2("refs/ucdp_countries.csv") |>
 vdem <- select(ctable, -country_name) |>
     right_join(vdem, by = "gwid")
 
+###
+# PKO Data
+pko <- read_xls("./data/raw/Third-Party-PKMs-version-3.5.xls") |>
+    select(idx = OBSNUM, COWcode = CCODE1, start = STARTYR, end = ENDYR) |>
+    mutate(end = ifelse(is.na(end), 2020, end)) |>
+    reframe(COWcode = first(COWcode), year = start:end, .by = idx) |>
+    distinct(COWcode, year) |>
+    mutate(pko = 1)
+
 full.df <- left_join(full_ep.df, vdem, by = c("gwno_a" = "gwid", "year")) |>
-    left_join(nmc, by = c("year", "COWcode" = "ccode"))
+    left_join(nmc, by = c("year", "COWcode" = "ccode")) |>
+    left_join(pko, by = c("COWcode", "year")) |>
+    mutate(pko = ifelse(is.na(pko), 0, 1))
 
 info("Missing V-Dem observation(s): %d", is.na(full.df$v2x_polyarchy) |> sum())
 info("Missing CINC observation(s): %d", is.na(full.df$cinc) |> sum())
+
+saveRDS(full.df, "data/ts.rds")
 
 ###
 # Collapse into an episode level dataset
@@ -144,6 +157,7 @@ final.df <- filter(full.df, !is.na(ext_sup_s_state)) |>
 
               incompatibility = max(incompatibility == 1),
               cold_war = ifelse(max(year) < 1989, 1, 0),
+              pko = any(pko == 1),
 
               ongoing_interstate = max(ongoing_interstate) > 0,
               ongoing_intrastate = max(ongoing_intrastate) > 0,

### 2024-05-23 R/models.R

@@ -11,10 +11,8 @@ df <- readRDS("./data/merged_data.rds") |> ungroup()
 treatments <- c("bin", "prop")
 outcomes <- c("frozen", "strict_frozen")
 units <- c("all", "cumulative_intensity")
-censored <- c("with_censored", "no_censored", "no_episode_censored")
 
-design <- expand.grid("outcome" = outcomes, "treatment" = treatments,
-                      "episodes" = units, "censored" = censored) |>
+design <- expand.grid("outcome" = outcomes, "treatment" = treatments, "episodes" = units) |>
     mutate(id = sprintf("%04d", row_number()))
 
 # Save the design matrix
@@ -29,25 +27,18 @@ for (row in 1:nrow(design)) {
     else
         df
 
-    # Filter out censored observations if necessary, ie conflicts
-    # beginning before 1975
-    if (design[row, "censored"] == "no_censored")
-        sub.df <- filter(sub.df, censored == 0)
-    else if (design[row, "censored"] == "no_episode_censored")
-        sub.df <- filter(sub.df, episode_censored == 0)
-
     # Select control variables - this is mostly static across models
-    X <- select(sub.df, episode_censored, episode_duration, recur, cumulative_intensity,
+    X <- select(sub.df, episode_censored, episode_duration, recur, pko,
+                cumulative_intensity,
                 incompatibility, cold_war, ongoing_intrastate, ongoing_interstate)
 
-    # If we've dropped censored episodes, no need for binary control
-    # var similar for cumulative intensity episodes
-    if (design[row, "censored"] != "with_censored")
-        X <- select(X, -episode_censored)
-    else if (design[row, "episodes"] == "cumulative_intensity")
+    if (design[row, "episodes"] == "cumulative_intensity")
         X <- select(X, -cumulative_intensity)
 
-    X <- mutate(X, episode_duration = log(episode_duration) |> scale()) |>
+    X <- mutate(X, episode_duration = scale(episode_duration),
+                episode_duration2 = scale(episode_duration^2),
+                episode_duration3 = scale(episode_duration^3)) |>
+        select(everything(), episode_duration, episode_duration2, episode_duration3) |>
         data.matrix()
 
     cases <- rowSums(is.na(X)) == 0

### 2024-05-23 R/probit.R

@@ -19,8 +19,8 @@ load(model_data)
 model_name <- basename(model_data) |> file_path_sans_ext()
 
 with(filter(design, id == model_name),
-     info("Model %s => outcome: %s, treatment: %s, episodes: %s, censored: %s",
-          id, outcome, treatment, episodes, censored))
+     info("Model %s => outcome: %s, treatment: %s, episodes: %s",
+          id, outcome, treatment, episodes))
 
 sprintf("posteriors/%s", model_name) |> dir.create(showWarnings = F, recursive = T)
 

### 2024-05-23 paper.qmd
Draft edits

@@
 execute:
-    cache: false
+    cache: true
     echo: false
@@
     conflict actors on the probability of an ongoing civil conflict
-    becoming frozen. I use an original dataset coding frozen conflicts as
+    becoming frozen. Specifically, I investigate how external state support shapes
+    the conditions under which actors may prefer a sub-optimal solution, allowing a
+    conflict to transition into a state of unstable peace with an ever present
+    threat of violent re-escalation, rather than a conclusive termination. In terms
+    of empirics, I use an original dataset coding frozen conflicts as
     an outcome among UCDP civil conflict episodes from 1975 to 2019 in
-    order to construct a series of bayesian, hierarchical probit models
+    order to construct a  bayesian, hierarchical probit model
     motivated by an application of bargaining theory to the domain of
@@
     conclude through a negotiated settlement or a one-sided military
-    victory. Furthermore, this paper also helps shed light on the
-    conditions under which actors may prefer a sub-optimal solution,
-    allowing a conflict to transition into a state of unstable peace with
-    an ever present threat of violent re-escalation, rather than a
-    conclusive termination.
+    victory.
 ---
@@
 design <- readRDS("./data/model_inputs/design.rds")
+labels <- read.csv("./refs/labels.csv")
 ```
@@
 They are unique among conflict outcomes in that they emerge from an inconclusive conflict termination, and leave conflict actors in an indeterminate state of unstable peace with an ever present threat of violent re-escalation.
+This state can be perpetuated for decades with alternations between mediation attempts and sporadic incidences of violence, and in some cases punctuated by full blown re-ignition of the initial war.
 
-Surprisingly, despite the geographic and temporal dispersion of frozen conflicts, little scholarly attention has been paid to the factors influencing their emergence [@klosek_frozen_2021].
-The frozen conflict specific literature mostly consists of qualitative case studies of specific conflicts, while the broader academic work on conflict termination has so far not incorporated frozen conflict as a unique conflict outcome.
+It is also notable the extent to which frozen conflicts are geograpically and temporally dispersed.
+The concept of frozen conflicts has long been associated with the separatist conflicts that emerged following the breakdown of the Soviet Union and Yugoslavia; however, numerous regions across the world have at different times been host to frozen conflicts including, for example, the self-declared Sahrawi Arab Democratic Republic in Western Sahara, Northern Cyprus, and Aceh in Indonesia.
+These conflicts have emerged at different time periods and also exhibit significant variation in the duration in which they can be classified as frozen.
+
+Surprisingly, despite the risk for future violence posed by frozen conflicts, little scholarly attention has been paid to the factors influencing their emergence [@klosek_frozen_2021].
+The limited frozen conflict specific literature mostly consists of qualitative case studies of specific conflicts, while the broader academic work on conflict termination has so far not incorporated frozen conflict as a unique conflict outcome.
 
@@
 There are a wide variety of potential factors which could plausibly be correlated with the emergence of frozen conflicts; however, I decide to focus on the international dimension, specifically the role of external support provided by third party state actors to conflict belligerents.
-There exists extensive work on the impact of external support on conflict onset, duration, and termination [@hartzell_civil_2016].
-Although there may be disagreements on the exact relationship between external support and various aspects of civil conflicts, the aggregate finding is that external support is an important factor.
+The reason is twofold.
+First, several of the most notable examples of frozen conflicts have featured the involvement of third party states, which opens the question for how common international involvement is within the larger universe of frozen conflicts.
+For example, Russia has played an out-sized role both as a supporter and mediator in several of the post-Soviet separatist conflicts that have since become frozen [@lynch_separatist_2002].
+Even in the case of Western Sahara, bordering states such as Algeria have involved themselves indirectly in the conflict as a supporter for the Polisario Front, and continue to allow the government-in-exile of the Sahrawi Arab Democratic Republic to reside in Tindouf, Algeria [@zunes_western_2022].
+Second, within the broader conflict literature there exists extensive work on the impact of external support on conflict onset, duration, and termination [@hartzell_civil_2016].
+Although there may be disagreements on the exact relationship between external support and various aspects of civil conflicts, the aggregate finding is that external support is an important factor for understanding conflict dynamics.
 
@@
 
-To empirically investigate this theory I provide an exploratory analysis using a series of bayesian hierarchical probit models with original coded data on frozen civil conflict onsets.
-The findings show that external state support does increase the probability that a civil conflict terminates in a frozen state, but only in the case of directed support to rebel groups in civil conflicts where the central incompatibility centers around territory.
-These results contribute to our understanding of frozen conflicts, as well as provide a more nuanced understanding of conflict termination more broadly.
+To empirically investigate this theory I provide an exploratory analysis using a bayesian hierarchical probit model with original coded data on frozen civil conflict onsets.
+The findings show that external state support does increase the conditional probability that a civil conflict terminates in a frozen state, but only in the case of directed support to rebel groups in civil conflicts where the underlying incompatibility centers around territory.
+These results contribute to our understanding of frozen conflicts, as well as provide a more nuanced understanding of conflict termination more broadly by identifying conflict episodes that terminate, but are nonetheless perpetuated as a frozen conflict.
+This latter point has implications for the study of recurrent conflicts since a frozen state is an important link between what may be otherwise perceived as discrete episodes of violence.
 
@@
 This was previously discussed in the definition for frozen conflict, but it is worth reiterating in order to emphasize that a stalemate situation is not sufficient to guarantee that a frozen conflict will emerge.
-For example, in the case where a conflict terminates in an indeterminate state, but one of the belligerents unilaterally renounces violence then the underlying incompatibility as an impetus for armed conflict would no longer be pertinent.
+For example, in the case where a conflict terminates in an indeterminate state, but one of the belligerents unilaterally renounces violence then the underlying incompatibility as an impetus for armed conflict would no longer be relevant.
+Such was the case with the Basque separatist group, Euskadi Ta Askatasuna (ETA), which unilaterally declared a cessation of armed activity in 2011 followed by complete disbandement in 2018.
 Such a scenario could hardly be described as an ongoing conflict that has frozen.
@@
 If an actor believes that an opponent will eventually renege on an agreement, they may be incentivized to fight even if they prefer to avoid war.
-A similar mechanism can be applied to frozen conflicts.
+This belief arises when there is uncertainty over potential future shifts in the distribution of relative capabilities absent guarantees that enforce the agreement, and can occur even in ideal scenarios where actors have complete information on the opponent's strength and resolve.[^democracy] This logic can be applied directly to frozen conflicts.
 Distrust and uncertainty can make it difficult to establish a bargaining space between actors.
-If neither actor is willing to continue fighting, then it may be preferable to let the conflict freeze if the costs of a frozen conflict are outweighed by the risk of the opponent reneging.
+If neither actor is willing or able to continue fighting, then it may be preferable to let the conflict freeze if the costs of a frozen conflict are outweighed by the risk of the opponent reneging.
+
+[^democracy]: For example, within the context of civil conflicts,
+    democratic political institutions are often considered to be one
+    viable pathway through which commitment problems can be overcome
+    by translating the transitory *de facto* power of the opposition
+    into *de jure* guarantees for the regular, peaceful contest of
+    office [@acemoglu_economic_2005; hegre_democracy_2014].
 
@@
 Therefore, a frozen conflict may arise when there is a mismatch between the wartime distribution and the expected distribution under a post-conflict settlement.
+This opens up for the possibility that a frozen conflict can provide a positive utility to actors, despite the accumulating costs of maintaining hostilities in a frozen state.
 
@@
 The degree to which each mechanism is relevant to any specific conflict depends in part on the conflict type.
-It is useful here to distinguish between civil conflicts fought over control of the state and civil conflicts fought over autonomy and/or independence claims.
+It is useful here to distinguish between civil conflicts fought over control of the state and civil conflicts fought over autonomy or independence claims.
 In the case of the former, the ultimate goal is to seize power from an incumbent regime.
-The prime impediment to a settlement is the risk that the post-conflict incumbent will violate the agreement, especially within authoritarian contexts where there is a fear of retribution.
-Thus, the preference for a frozen conflict, assuming that neither side can achieve a military advantage, may primarily stem from credible commitment problems.
+Absent a clear military victory, a bargained solution will typically imply some type of power-sharing agreement.
+In such a scenario, the prime impediment to a settlement is the risk that the post-conflict incumbent will violate the agreement, especially within an authoritarian context where there is a fear of retribution.
+Thus, the preference for a frozen conflict, again assuming that neither side can achieve a military advantage, may primarily stem from credible commitment problems.
 
-In the case of the latter, civil conflicts fought over autonomy and/or independence claims, can also involve credible commitment issues as an incentive towards frozen conflict.
-Specifically, it is difficult for the government to credible commit to an agreement granting greater autonomy, in part because it may galvanize additional claims from other groups [@walter_information_2006; @toft_self-determination_2012].
+In the case of the latter, civil conflicts fought over autonomy and/or independence claims, can also involve credible commitment issues as an incentive towards frozen conflict for the opposition.
+Specifically, it is difficult for the government to credible commit to an agreement granting greater autonomy, in part because it may create a precedent that galvanizes additional claims from other groups [@walter_information_2006; @toft_self-determination_2012].
+Thus, even if the government is willing to negotiate it must balance between the immediate demands from one group versus potential future claims by other groups which may threaten territorial stability.
+This creates a dilemma that undermines the government's credibility when bargaining.
 Furthermore, as some scholars have argued, the international community tends to be fairly conservative regarding new sovereignty claims [@huddleston_foulweather_2021; @siroky_unilateral_2021].
@@
 If the rebel group is unable to trust the government to credibly bargain, then they will highly value the present transitory gains and seek to maintain the status quo.
+
 Meanwhile, the government may also face an incentive towards a frozen conflict in these scenarios.
-There is a strong international norm in favor territorial integrity.[^IR] If the government is unable to outright defeat a rebel group, especially one that has established a *de facto* independent state, then the government may be seeking a long-term plan for forcible reintegration.
+Some scholars have argued that there is a strong international norm in favor territorial integrity [@zacher_territorial_2001].[^IR] If the government is unable to outright defeat a rebel group, especially one that has established a *de facto* independent state, then the government may be seeking a long-term plan for forcible reintegration.
 A frozen conflict allows the government to ease the burden of active fighting while maintaining their sovereignty claims on breakaway territory.
-Crucially, this latter point allows the government to frame the frozen conflict as an ongoing civil war if and when there is another violent re-escalation, as opposed to an interstate war of territorial conquest against a newly independent state.
+Crucially, this latter point permits the government to frame the frozen conflict as an ongoing civil war if and when there is another violent re-escalation, as opposed to an interstate war of territorial conquest against a newly independent state.
 
-[^IR]: I need to check the IR literature to provide a citation for this claim.
+[^IR]: It is worth noting that this claim as it relates to the prevalance of territorial conquest has been contested, see for example @altman_evolution_2020 and @altman_why_2022 where it is argued that small-scale territorial conquest has persisted in the post-WW2 era.
+However, complete conquest of an opposing state has not occurred, rendering it difficult for a state to forcibly reintegrate former territory that has achieved internationally recognized *de jure* independence.
 
@@
 This does not mean that the conflict will automatically become frozen, but it may raise the likelihood of a stalemate which is an important first step.
-Furthermore, external support also helps subsidize the potential cost of maintaining a frozen conflict, which may allow actors a longer time horizon such that they will reject an immediate settlement if they are optimistic about the future utility of renewed fighting.
+As a second step, external support may also create uncertainty over the *future* balance of power.
+Recall the credible commitment mechanism, whereby actors may prefer war over a bargained solution when there exists uncertainty over the future distribution of power, absent guarantees that the opposing side will not renege.
+Previous instances of external support may create an expectation of future support, affecting the ability of an actor to credibly commit to bargaining in the present [@karlen_legacy_2017].
+Faced with a stalemate situation and a credible commitment dilemma, actors may prefer a frozen conflict.
 
-Scholars have also argued that the involvement of additional third party actors can complicate the bargaining process for an ongoing civil conflict [@cunningham_veto_2006; @cunningham_blocking_2010].
+In addition, external support may also help subsidize the potential cost of maintaining a frozen conflict.
+This is crucial for actors that value the present status quo.
+For example, offsetting the cost of a frozen conflict allows actors to maximize the utility gained from transitory war gains, which may be particularly important in the case of territorial civil conflicts, as discussed previously.
+It may also allow actors to extend their time horizons such that they will reject an immediate settlement if they are optimistic about the future utility of renewed fighting.
+For governments aiming to forcibly reintegrate a breakaway territory, external support aids and subsidizes the waiting time during which the government aims to improve their military forces in order to relitigate the conflict in the future.
+
+Finally, scholars have also argued that the involvement of additional third party actors can complicate the bargaining process for an ongoing civil conflict [@cunningham_veto_2006; @cunningham_blocking_2010].
 External supporters may have conflicting outcome preferences to the actors they are supporting.
 As the number of actors involved in a bargaining process grows, coordinating the disparate interests becomes increasingly difficult, another disincentive against achieving a negotiated settlement.
+In the face of competitive interventions, third party actors may also be unwilling to fully invest the resources necessary for a military victory if they are not confident that there will be a favourable post-conflict order.
+In such a scenario, in may be in the external supporters interest to perpetuate the conflict in a frozen state.
 
-In sum, external support should increase the likelihood of a stalemate and incentivize actors to maintain the transitory status quo in the form of a frozen conflict.
-My empirical expectation is that the effect should be strongest when directed support is only provided to the rebels.
+In sum, external support should increase the likelihood of a stalemate and incentivize actors to maintain the transitory status quo in the form of a frozen conflict through any of three aforementioned mechanisms.
+Therefore, my empirical expectation is that there should be a positive correlation between external third party support and the probability of a civil conflict becoming frozen.
+On the basis of previous empirical findings highlighting the importance of support to the opposition, I expect the effect should be strongested when directed support is only provided to the rebels.
 When support is either provided to the government, or to both sides, then it is less likely that a balance of power will emerge and I assume the situation will on average favor the government.
@@
 
-To empirically assess whether external state support for conflict actors affects the probability of a civil conflict episode terminating as a frozen conflict a series of bayesian probit models were run using original data on frozen civil conflicts.
+To empirically assess whether external state support for conflict actors affects the probability of a civil conflict episode terminating as a frozen conflict a bayesian probit model was run using original data on frozen civil conflicts.
 In what follows, I will first present the model specification followed by a discussion of the included dependent and independent variables, including operationalizations and data sources.
@@
 
-For $\delta_k$ and $\gamma$ I use hierarchical priors that allow for varying effects while partially pooling information across groups to prevent overfitting.
+For $\delta_k$ and $\gamma$ I use hierarchical priors that allow for varying effects while partially pooling information across groups to guard against overfitting.
 Regarding the former, this means $\delta_{jk} \sim \text{N}(\mu_j, \tau_j)$ for $j = 1, 2, \ldots, D$ where $\mu_j \sim \text{N}(0, 1)$ and $\tau_j \sim \text{N}(0, 2.5)$.
@@
 [^country-intercepts]: The risk for overfitting is especially high for the country-specific intercepts since the number of observations per country is relatively small with the median country only experiencing two episodes within the time period under study.
+The practical effect is that the intercepts for countries with few observations will be pulled towards zero with an increase in the posterior uncertainty intervals.
 
@@
 [^strict_frozen]: In the [appendix][robustness checks] I extend the analysis presented in this paper and run a series of robustness checks where I adopt a more strict definition of conflict termination requiring that frozen conflicts have more than 2 years below the UCDP threshold for an active conflict year in order to better distinguish between frozen conflicts and short-term pauses in fighting.
-TODO: what are the results?
 
@@
 Compared to the only other extant coding of frozen conflicts, the Frozen Conflict Dataset (FCD) [@klosek_frozen_2021], the measure that I employ features a more expansive operationalization of the concept of frozen conflicts.
-Although the FCD covers a wider period from 1946 until 2011, only civil conflicts involving *de facto* states according to the coding by @florea_facto_2014 are included resulting in 17 cases of frozen conflict.
+Although the FCD covers a wider period from 1946 until 2011, only civil conflicts involving *de facto* states according to the coding by @florea_facto_2014 are included resulting in 17 cases of frozen civil conflicts.
 Furthermore, the termination year for a civil conflict episode in the FCD is coded as the last year in which the number of battle related deaths is above 500.
@@
 To mitigate this risk, I also run the analysis presented in this paper on a sub-sample of conflicts with a cumulative intensity of at least 1,000 battle related fatalities since onset in the [appendix][robustness checks].
-TODO: results
 
@@
 
-In addition to the treatment variables I also include a set of covariates that could plausibly confound the relationship between external state support and frozen conflicts.
-This includes variables related to  conflict dynamics, specifically: conflict episode duration, episode recurrence within a conflict, and cumulative intensity.
+In addition to the treatment variables I also include a set of covariates that could plausibly confound the relationship between external state support and frozen conflicts.[^descriptive] This includes variables related to conflict dynamics, specifically: conflict episode duration, episode recurrence within a conflict, and cumulative intensity.
 Conflicts which last longer may affect the willingness of third party actors to offer support to belligerents.
@@
 We could expect the costs to increase such that active fighting ceases; however, the underlying conflict remains unresolved and the conflict transitions to being frozen.
-Thus, I include the number of years for which each episodes lasts until termination transformed with the natural logarithm function in order to reflect the diminishing impact of each one unit increase.
+Thus, I include the number of years each episodes lasts until termination as cubic polynomial terms in order to capture a potential non-linear effect [@carter_back_2010].
 A similar logic may also apply to episode recurrence.
@@
 
+[^descriptive]: Descriptive statistics for the control variables as well as the outcome and treatments are provided in @tbl-stats in the [appendix][Descriptive Statistics].
+
 In complement to the partially pooled varying treatment effects, I also include a binary indicator for whether the conflict incompatibility concerns territory in order to account for a potential differing baseline probability for termination as a frozen conflict and the possibility that the conflict incompatibility affects the likelihood of external support.
-Further, I include variables to control for the Cold War and ongoing inter/intra-state conflicts within the host country.
-The former is a binary indicator for whether the termination year falls before 1989 to capture temporal dynamics related to the propensity for different types of conflict terminations.
+Moreover, I include binary variables to control for the Cold War, ongoing inter/intra-state conflicts, and the presence of peacekeeping operations within the host country.
+The first is an indicator for whether the termination year falls before 1989 to capture temporal dynamics related to the propensity for different types of conflict terminations.
 For example, international norms concerning external intervention in civil conflicts may be shaped in part by the state of the international system influencing both the decision by third party actors to intervene in a civil conflict and the types of conflict terminations that external interveners seek [@howard_how_2017].
-The latter are binary indicators as well measuring whether an interstate or intrastate conflict occurred at any time during the conflict episode.
+The second are variables measuring whether an interstate or intrastate conflict occurred at any time during the conflict episode.
 Third party states involved in one conflict may be more likely to also intervene in another ongoing conflict within the same country.
-Furthermore, multiple ongoing conflicts could also indicate a weakened incumbent government forced to face multiple opponents, which may in turn affect the probability that a conflict freezes.
-Lastly, I include a binary variable for whether a conflict is left censored, *i.e.*, that the onset of the episode occurred before the coding start for the dataset, since data on external support from the ESD is only available from 1975 until 2017.[^robust_censor]
-
-[^robust_censor]: I also run multiple robustness checks in the [appendix][robustness checks] where I alternatively drop episodes with an episode onset year before the coding start for the dataset and/or recurrent episodes within conflicts that also began prior to 1975.
+Furthermore, multiple ongoing conflict could also indicate a weakened incumbent government forced to face multiple opponents, which may in turn affect the probability that a conflict freezes.
+The third is drawn from the Third-Party Peacekeeping Missions Dataset version 3.5 [@mark_j_mullenbach_third-party_2013] which codes UN and non-UN led peacekeeping operations.
+Lastly, I include a binary variable for whether a conflict is left censored, *i.e.*, that the onset of the episode occurred before the coding start for the dataset, since data on external support from the ESD is only available from 1975 until 2017.
 
@@
 
-main_model <- filter(design, outcome == "frozen", treatment == "bin", episodes == "all",
-                     censored == "with_censored") |> pull(id)
+main_model <- filter(design, outcome == "frozen", treatment == "bin", episodes == "all") |>
+    pull(id)
 
@@
     facet_grid(. ~ group) +
+    theme(strip.text = element_text(face = "bold")) +
     scale_fill_solarized()
@@
 
-From the parameter plots in @fig-delta, the main driver for this substantive results is the estimate for external support given to the rebel(s) in the context of a territorial civil conflict.
-In this case, the probability for the treatment parameter being greater than zero is `r mean(delta2$value > 0) |> signif(2)` with $\text{CI}_{90\%}: [`r quantile(delta2$value, 0.05) |> signif(2)`, `r quantile(delta2$value, 0.95) |> signif(2)`]$.
+From the parameter plots in @fig-delta, the main driver for these substantive results is the estimate for external support given to the rebel(s) in the context of a territorial civil conflict.[^controls] In this case, the probability for the treatment parameter being greater than zero is `r mean(delta2$value > 0) |> signif(2)` with $\text{CI}_{90\%}: [`r quantile(delta2$value, 0.05) |> signif(2)`, `r quantile(delta2$value, 0.95) |> signif(2)`]$.
 Taken together, these results indicate that the crucial predictor for whether an intra-state conflict terminates as a frozen conflict is whether the opposition received external support by a third party state during a conflict episode centered around a territorial incompatibility.
@@
 
+[^controls]: In addition to the treatment parameters presented here, @fig-controls in the [appendix][Additional Model Estimates] plots the posterior 50%/90% posterior quantile intervals for the non-treatment covariate coefficients.
+
+These results are partially confirmed by the [robustness checks][Robustness Checks] in the appendix.
+Specifically, under the alternative specifications, external support to the rebel(s) is found to still have a positive effect on the posterior probability of a civil conflict becoming frozen given that it has terminated.
+However, in contrast to the main results presented above, this effect is only under the condition of no matching external support to the government.
+In other words, the average marginal effect for external support given to both sides is no longer informative for predicting whether a terminated civil conflict becomes a frozen conflict.
+
 # Conclusion
 
-This paper sought to investigate a potential link between external support offered by third party states and the likelihood that a civil conflict transitions into a frozen conflict.
-A theory was offered emphasizing the expected utility of actors
-and how external support can shape the cost-benefit calculation in favour of a frozen outcome.
-This was then investigated empirically with a bayesian hierarchical probit model showing correlationary evidence that directed support to rebel groups does increase the probability that a civil conflict freezes, but only within the context of territorial civil conflicts.
+This paper has sought to investigate the potential link between external support offer by third party states and the likelihood that a civil conflict transitions into a becoming a frozen conflict.
+A theory was offered emphasizing the expected utility of actors and how external support can shape the cost-benefit calculation in favour of a frozen outcome.
+Specifically, it was argued that there are three plausible mechanisms by which actors may come to prefer a frozen conflict over a settlement when unable to achieve a military victory: credible commitments, transitory war gains, and time horizons.
+External state support can play an important role in all three by shaping the current and future distribution of power between conflict belligerents and subsidizing the accumulating costs of maintaining a frozen conflict such that the utility of transitory war gains can be maximized and actors can extend their time horizons.
 
-One potential shortcoming is that I did not take into account the preferences of external supporters and motivation for offering support.
-Nor did I disaggregate between different types of external support.
-In line with other empirical work on the disaggregated effects of external support and conflict duration, there may be a differential effect depending on the type and amount of support offered to conflict belligerents.
+The empirical results indicate that external state support directed to the opposition within the context of territorial civil conflicts does have a positive effect on the probability that a civil conflict becomes frozen given that it has terminated.
+However, the results provide only correlational evidence and are hampered by the observational nature of the data.
+Furthermore, from the robustness checks it is unclear if external support to the opposition holds when similar support is given to the government in territorial civil conflicts.
 
-Furthermore, the measurement offered for frozen conflicts was necessarily crude.
-It operationalized frozen conflict as a discrete event, "frozeness" maybe is best conceptualized as lying on a continuous dimension that allows conflicts to fluctuate  between "more" and "less" frozen.
-Future work should focus on refining this measure and to hopefully further expand on our understanding of both frozen conflicts and civil conflict termination.
+Nonetheless, these results provide preliminary evidence that external support may be an important factor for understanding the emergence of frozen conflicts.
+They also contribute to the pre-existing literature on conflict termination by providing a more nuanced picture of conflict outcomes with original coded data that should help research distinguish between conclusively terminated conflicts, and civil conflicts that instead transition to becoming frozen.
+
+Future research should continue to investigate the potential link between external support and frozen conflicts, examining for example the disaggregated effects of different types of support as well as whether the amount of support, especially in terms of material aid, has a significant impact.
+Furthermore, the measurement offered for frozen conflicts can be a starting point for more sophisticated attempts at measuring frozen conflicts on a continuous dimensions that allows conflicts to fluctuate between "more" and "less" frozen.
+Such an approach would allow researchers to better understand how conflicts transition between states of active and frozen conflict.
 
@@
 
-\newpage
-
 ## Simulation Based Calibration
@@
 #|     2000$. Posteriors were not thinned since the effective sample
-#|     sizes showed anticorrelation among the MCMC draws.
+#|     sizes showed anticorrelation in the MCMC draws.
 sbc <- readRDS("./posteriors/sbc.rds") |>
@@
 
-\newpage
+## Descriptive Statistics
+
+```{R}
+#| label: tbl-stats
+#| tbl-cap: Descriptive statistics for both the dependent and
+#|          independent variables in the dataset. In total there are 212
+#|          conflict episodes spanning from 1975 - 2017.
+
+load("./data/model_inputs/0001.RData")
+
+descriptive <- function(x, name) {
+    list(Variable = labels$label[labels$variable == name],
+         Mean = mean(x),
+         Min = min(x),
+         Max = max(x))
+}
+
+# Note, grab duration from sub.df since episode_duration in the
+# covariate matrix X has already been normalized.
+covariates <- as.data.frame(X) |>
+    select(-matches("episode_duration")) |>
+    bind_cols(treatments, select(sub.df, frozen, episode_duration))
+
+tbl <- lapply(colnames(covariates), \(name) descriptive(covariates[[name]], name)) |>
+    bind_rows() |>
+    arrange(Variable)
+
+kable(tbl, digits = 2)
+```
+
+## Additional Model Estimates
+
+```{R}
+#| label: fig-controls
+#| fig-cap: Posterior estimates with 50% and 90% quantile intervals
+#|          for the non-treatment covariate coefficients in the probit
+#|          model presented in the main text. Not displayed are the
+#|          partially pooled, country-specific varying-intercepts.
+fit <- readRDS("./posteriors/0001/fit.rds")
+
+pars <- fit$draws(c("alpha", "beta"), format = "matrix")
+
+colnames(pars)[grepl("alpha", colnames(pars))] <- "Intercept"
+colnames(pars)[grepl("beta", colnames(pars))] <- labels$label[match(colnames(X), labels$variable)]
+
+mcmc_intervals(pars)
+```
 
@@
 Although the conceptual definition of frozen conflict that I employ does not include an explicit temporal dimension as a requirement for a conflict to be considered frozen, it could be nonetheless argued that such pauses should not be considered candidates for being classified as frozen since actors may view the conflict as active and ongoing.
-Therefore, I also construct a stricter measure for frozen conflict whereby a conflict episode has to be followed by more than two years below the UCDP battle related threshold to be considered eligible to being classified as frozen, conditional on the remaining the coding criteria.
+Therefore, I also construct a stricter measure for frozen conflict whereby a conflict episode has to be followed by more than two years below the UCDP battle related threshold to be considered eligible to being classified as frozen, conditional on the remaining coding criteria.[^strict_num]
+
+[^strict_num]: From the original dataset, `r sum(dataset$strict_frozen)` civil conflict episodes from 1975 until 2017 were coded as frozen conflicts following this stricter operationalization.
 
@@
 It can be argued that minor, long-running insurgencies should not be considered as frozen conflicts since the underlying issue may no longer be particularly salient and there is a relatively low risk for a major escalation in violence.
-Such conflicts may experience sporadic, isolated incidences of violence, but would otherwise are minor factors in the domestic political agenda.
-To this end, I subset the list of conflict episodes to only include civil conflict episodes that have cumulatively experienced at least 1,000 battle related fatalities since onset in the hopes that major conflicts maintain a greater degree of salience post-termination.
+Such conflicts may experience sporadic, isolated incidences of violence, but would otherwise be considered minor within the larger issue of domestic politics.
+To this end, I subset the list of conflict episodes to only include civil conflict episodes that have cumulatively experienced at least 1,000 battle related fatalities since onset in the hopes that major conflicts maintain a greater degree of salience post-termination.[^high_intensity]
 
-Third, due to the truncated nature of the dataset, some of the civil conflict episodes are censored with missing data before 1975.
-In the main results I included a binary variable if the onset of an episode occurred before the start of the dataset.
-As an alternative, in the following robustness checks I also drop all censored episodes as well as censored conflicts.
+[^high_intensity]: This leads to a dataset consisting of `r dplyr::filter(dataset, cumulative_intensity == 1) |> nrow()` civil conflict episodes of which `r dplyr::filter(dataset, cumulative_intensity == 1, frozen == 1) |> nrow()` satisfy the original operationalization of frozen conflict and `r dplyr::filter(dataset, cumulative_intensity == 1, strict_frozen == 1) |> nrow()` episodes satisfy the stricter, alternative operationalization.
 
 Finally, the primary treatment variables used in the main text are aggregated binary measures for directed support, plus the interaction, at any time during a civil conflict episode.
-Although this aids in the interpretation of the treatment effect, these variables do not capture the extent to which support is provided during a conflict episode.
+Although this aids in the interpretation of the associated model parameters, these variables do not capture the extent to which support is provided during a conflict episode.
 Thus, I also replace the treatment variables with measures capturing the proportion of conflict-years within an episode that each actor received external support.
 
-In total, this leads to `r list.files("posteriors", "\\d{4}") |> length()` different model configurations for the original vs strict definition of frozen conflict, inclusion and exclusion of low-intensity conflict episodes, inclusion and exclusion of censored episodes and/or conflicts, and finally the proportional aggregation of the treatment variables.
-@fig-robust
+In total, this leads to `r list.files("posteriors", "\\d{4}") |> length()` different model configurations for the original versus strict definition of frozen conflict, inclusion and exclusion of low-intensity conflict episodes, and finally the proportional aggregation of the treatment variables.
+@fig-robust-1 and @fig-robust-2 plot the posterior average marginal effect for directed external support given only to the government during the course of a civil conflict episode.
+The former graph shows the estimates for the different models using all of the conflict episodes, while the latter estimates are fitted using only high intensity conflict episodes.
+In all cases, the posterior estimates indicate that the parameters are fairly uninformative for predicting whether a terminated civil conflict episode freezes.
+The highest posterior probability that the average marginal effect is greater than zero is only slightly above 0.8 across all models.
+This mirrors the results from the model presented in the main text.
 
 ```{R}
-#| label: fig-robust
-#| fig-cap: Robustness checks
+#| label: fig-robust-1
+#| fig-cap: Posterior estimates for the average marginal effect of
+#|          external support to the government only varied by
+#|          incompatibility type across the different model
+#|          configurations for all conflict episodes.
+
+appendix_plot <- function(df) {
+    ggplot(df, aes(pzero, median, colour = treatment)) +
+        geom_point(aes(shape = outcome), alpha = 0.9, size = 2) +
+        geom_errorbar(aes(ymin = q5, ymax = q95), alpha = 0.5, width = 0) +
+    geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.2) +
+    facet_grid(. ~ group) +
+    theme(panel.border = element_rect(fill = NA, colour = "grey"),
+          strip.text = element_text(face = "bold"),
+          plot.title = element_text(face = "bold", hjust = 0.5)) +
+    xlab(TeX("$P(AME_{jk} > 0)$")) +
+    ylab(TeX("$AME_{jk}$")) +
+    scale_color_solarized()
+}
 
 models <- list.dirs("posteriors", recursive = F)
-stopifnot(length(models) != 0)
+stopifnot(length(models) > 0)
 
@@
 
-rebel <- filter(model_ames, variable == "Ext. Sup. Rebel Only", censored == "with_censored")
-ggplot(rebel, aes(pzero, median, colour = treatment)) +
-    geom_point(aes(shape = outcome), alpha = 0.8) +
-    geom_errorbar(aes(ymin = q5, ymax = q95), alpha = 0.1) +
-    geom_vline(xintercept = 0.9, linetype = "dashed", alpha = 0.2) +
-    facet_grid(episodes ~ group)
+filter(model_ames, variable == "Ext. Sup. Gov Only", episodes == "all") |>
+    appendix_plot() +
+    ggtitle("Ext. Sup. Gov Only - All Episodes")
+```
 
-#filter(model_ames, variable == "Ext. Sup. Rebel Only") |>
-#ggplot(aes(pzero, median, colour = treatment)) +
-#    geom_point(alpha = 0.2) +
-#    geom_errorbar(aes(ymin = q5, ymax = q95), alpha = 0.25) +
-#    geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.2) +
-#    geom_vline(xintercept = 0.9, linetype = "dashed", alpha = 0.2) +
-#    facet_grid(variable ~ group) +
-#    labs(y = parse(text = "tilde(Delta)"), x = parse(text = "P(tilde(Delta) > 0)")) +
-#    theme(legend.position = "none")
+```{R}
+#| label: fig-robust-2
+#| fig-cap: Posterior estimates for the average marginal effect of
+#|          external support to the government only varied by
+#|          incompatibility type across the different model
+#|          configurations for only high intensity conflict episodes.
 
+filter(model_ames, variable == "Ext. Sup. Gov Only", episodes == "cumulative_intensity") |>
+    appendix_plot() +
+    ggtitle("Ext. Sup. Gov Only - High Intensity Episodes")
 ```
+
+@fig-robust-3 and @fig-robust-4 similarly plot the estimates for the average marginal effect of external support to only the opposition for each aforementioned model.
+Again, the results are largely inline with the conclusions drawn from the model presented in the main text.
+The posterior probability that the average marginal effect for external support to the rebel side is greater than zero is greater than 0.9 across all specifications for civil conflict episodes centering on a territorial incompatibility.
+In the case of civil conflicts where the contest is over control of the central government, the average marginal effect is found to be positively associated with the probability of a terminated episode freezing only for the models fit on the high intensity episode data whilst excluding the proportional treatment for the standard operationalization of frozen conflict.
+
+```{R}
+#| label: fig-robust-3
+#| fig-cap: Posterior estimates for the average marginal effect of
+#|          external support to the rebel side only varied by
+#|          incompatibility type across the different model
+#|          configurations for all conflict episodes.
+
+filter(model_ames, variable == "Ext. Sup. Rebel Only", episodes == "all") |>
+    appendix_plot() +
+    ggtitle("Ext. Sup. Rebel Only - All Episodes")
+```
+
+```{R}
+#| label: fig-robust-4
+#| fig-cap: Posterior estimates for the average marginal effect of
+#|          external support to the rebel side only varied by
+#|          incompatibility type across the different model
+#|          configurations for only high intensity conflict episodes.
+
+filter(model_ames, variable == "Ext. Sup. Rebel Only", episodes == "cumulative_intensity") |>
+    appendix_plot() +
+    ggtitle("Ext. Sup. Rebel Only - High Intensity Episodes")
+```
+
+Finally, @fig-robust-5 and @fig-robust-6 depict the average marginal effects when both sides in a civil conflict episode receive support.
+In the main text this was found to be positively associated with the probability of a frozen conflict onset, largely being driven by the parameter for external support to the rebel side --- in other words, external support for the opposition was found to be positively associated with frozen conflict onset even in the presence of external support to the government.
+The results here indicate that this relationship is fairly non-robust across the alternative models.
+Rather, external support support for the rebel side is positively associated with the probability of frozen conflict onset, but only in the absence of support to the government.
+
+```{R}
+#| label: fig-robust-5
+#| fig-cap: Posterior estimates for the average marginal effect of
+#|          external support to both sides varied by incompatibility
+#|          type across the different model configurations for all
+#|          conflict episodes.
+
+filter(model_ames, variable == "Ext. Sup. to Both", episodes == "all") |>
+    appendix_plot() +
+    ggtitle("Ext. Sup. to Both - All Episodes")
+```
+
+```{R}
+#| label: fig-robust-6
+#| fig-cap: Posterior estimates for the average marginal effect of
+#|          external support to both sides varied by incompatibility
+#|          type across the different model configurations for only
+#|          high intensity conflict episodes.
+
+filter(model_ames, variable == "Ext. Sup. to Both", episodes == "cumulative_intensity") |>
+    appendix_plot() +
+    ggtitle("Ext. Sup. to Both - High Intensity Episodes")
+```

### 2024-05-23 paper.qmd
Draft edits

@@
 The frozen conflict data was created using a mix of programmatic and manual coding in a two-stage process.
-The first stage filtered out dyadic civil conflict episode outcomes according to a preset list of rules listed in detail below.
+The first stage filtered out civil conflict episode outcomes according to a preset list of rules listed in detail below.
 This candidate list was then manually corrected in the second stage to form the final list of frozen conflict onsets.
@@
 
-Dyadic conflict episode outcomes were drawn from the UCDP Conflict Termination dataset v21.1.
+Conflict episode outcomes were drawn from the UCDP Conflict Termination dataset v21.1.
 This means that the definition of a conflict episode and its corresponding termination followed the coding conventions of the UCDP dataset.
-Namely, dyad-years are coded as active conflicts when the number of battle related fatalities rises above a threshold of 25.
+Namely, conflict-years are coded as active conflicts when the number of battle related fatalities rises above a threshold of 25.
 Consecutive conflict years, or episodes, are terminated the first year when this number drops below the threshold.

### 2024-12-09 lib/fc.utils/R/modelsummary.R
Remove deprecated functions

@@ -1,60 +1,0 @@
-#' @export
-model_info <- function(s) UseMethod("model_info", s)
-
-#' @export
-model_info.character <- function(s) {
-    v <- strsplit(s, "-")[[1]]
-    if (length(v) != 3 || any(nchar(v) == 0))
-        stop(sprintf("Misformed model string? %s", s))
-
-    list(outcome = v[1], treatment_aggregation = v[2], episodes = v[3])
-}
-
-#' @importFrom parameters model_parameters
-#' @export
-model_parameters.CmdStanMCMC <- function(model, pars = c(), ...) {
-    # Note, modelsummary is calling `parameters` with do.call with
-    # args as `list(model = model, ...)`. If the first function
-    # argument for our method is called anything besides `model`, it
-    # will silently fail because modelsummary decided to silence all
-    # error messages...
-    draws <- model$draws(pars, format = "df")
-
-    # This is a hack, but we need to recover the original variable names
-    # in the right order
-    labels <- dirname(model$data_file()) |> file.path("labels.rds") |> readRDS()
-
-    args <- append(list(.x = draws), labels)
-    draws <- do.call(posterior::rename_variables, args)
-
-    parameters::parameters(draws, ...)
-}
-
-#' @export
-post_summarise <- function(x, ...) UseMethod("post_summarise", x)
-
-#' @export
-post_summarise.CmdStanFit <- function(x, pars, model_info = F, rename = F, probs = c(0.05, 0.5, 0.95)) {
-    draws <- x$draws(pars)
-
-    if (isTRUE(rename)) {
-        f <- dirname(x$data_file()) |> file.path("labels.rds")
-        if (!file.exists(f))
-            stop("No labels found for model")
-
-        labels <- readRDS(f)
-        args <- append(list(.x = draws), labels[labels %in% posterior::variables(draws)])
-        draws <- do.call(posterior::rename_variables, args)
-    }
-
-    df <- posterior::summarise_draws(draws, ~stats::quantile(.x, probs))
-
-    if (isTRUE(model_info)) {
-        # This is also a horrible hack
-        model <- dirname(x$data_file()) |> basename()
-        li <- model_info(model)
-        df <- cbind(df, li)
-    }
-
-    return(df)
-}

### 2024-12-09 paper.qmd
Update author info

@@
 - name: J.L.A. Krusell
-  email: "joshua.krusell@gu.se"
-  affiliations:
-      - "University of Gothenburg"
 date: today
